---
title: "Features extracted with Python - Data import, merging, and tidying"
author: "Sandy Kroh"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    fig_caption: yes
    keep_md: yes
  pdf_document:
    toc: yes
params:
  extraction_parameter: cells
  mask_parameter: feature_0
  Sample: Default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", 
                      fig.align='center', 
                      fig.width = 9, 
                      fig.height = 6,
                      dpi = 600, 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Libraries

```{r}
library(readr) 
library(tidyverse) 
library(tidyr) 
library(dplyr) 
library(here) 
library(ggpubr)
library(readxl)
```

## Parameters

```{r}

set.seed(8) 

cols_treat <- c("darkcyan", "gold", "deeppink", "slateblue")  

cols_fov <- c("slateblue1",              
              "darkcyan",                         
              "seagreen2")   

cols_exp <- c("darkcyan", "seagreen2", "lightgreen",     
              "gold", "yellow", "olivedrab1",            
              "deeppink","#FF0066", "magenta",           
              "slateblue", "blueviolet", "purple")   

cols_nat <- c("darkcyan", "seagreen2", "lightgreen",              
              "gold", "yellow", "olivedrab1",              
              "deeppink","#FF0066", "magenta",           
              "slateblue", "blueviolet", "purple",               
              "cyan", "blue", "navyblue",               
              "brown", "red", "orange", "grey")  
```

# Import data

## Define directories

```{r}

# define path to general data main directory
data_dir_main <- here("data")

# define path to input directory
input_dir <- here("data", "Feature_extraction_python")

# define output folder for saing the environment etc. 
output_env_dir <- here("1_revisions", "1_Data_loading_merging_tidying_files") 
dir.create(output_env_dir)

```

## Load csv output files from python feature extraction

Define function to automatically extract file information:

```{r}
#' get file name ----
get_file_name = function(cells_csv){
  cells <- read.csv(paste0(input_dir, "/", cells_csv), header=TRUE) # read the cells csv file
  filename <- gsub(".*/", "", cells_csv) # remove anything before the last underscore; removes "cell.csv"
  filename <- gsub("_cell_masks.csv", "", filename)
  return(filename)
  
}
```

```{r}
my_data_list <- list.files(path = input_dir, pattern = "membrane.csv") # creates a list with all the cell.csv files
my_data_list
file <- my_data_list[1]

df_raw_mean <- data.frame()
df_raw_median <- data.frame()

filename <- gsub(".csv", "", file)
# opens the nuclei csv file
df_samp <- read.csv(paste0(input_dir, "/", file), header=TRUE) 
df_samp$Experiment_ID <- gsub("_segmentation_label_cell_membrane", "", filename)

# filter data for mean
df_samp_mean <- df_samp %>%
  # delete everything after the first dot
  mutate(Label = gsub("Vimentin", "VIM", Label), 
         Label = gsub("Fibronectin", "FN", Label), 
         Label = gsub("aGFP", "GATA3eGFP", Label), 
         Label = gsub("Emcn", "EMCN", Label), 
         Label = gsub("\\..*", "", Label), 
         Loc_X = X, 
         Loc_Y = Y
         ) %>%
  select(Label, Mean, Loc_X, Loc_Y, Experiment_ID, Area) %>%
  # change long format to wide format
  pivot_wider(names_from = Label, values_from = c(Mean))

# filter data for median
df_samp_median <- df_samp %>%
  # delete everything after the first dot
  mutate(Label = gsub("Vimentin", "VIM", Label), 
         Label = gsub("Fibronectin", "FN", Label), 
         Label = gsub("aGFP", "GATA3eGFP", Label), 
         Label = gsub("Emcn", "EMCN", Label), 
         Label = gsub("\\..*", "", Label), 
         Loc_X = X, 
         Loc_Y = Y
         ) %>%
  select(Label, Median, Loc_X, Loc_Y, Experiment_ID, Area) %>%
  # change long format to wide format
  pivot_wider(names_from = Label, values_from = c(Median))

df_raw_mean <- data.table::rbindlist(list(df_raw_mean, df_samp_mean), 
                                     fill = TRUE)
df_raw_median <- data.table::rbindlist(list(df_raw_median, df_samp_median), 
                                       fill = TRUE)


# # loop for loading all files and combining them
# for (file in my_data_list) {
#   filename <- gsub(".csv", "", file)
#   # opens the nuclei csv file
#   df_samp <- read.csv(paste0(input_dir, "/", file), header=TRUE) 
#   df_samp$MELC_ID <- filename
#   
#   # filter data for mean
#   df_samp_mean <- df_samp %>%
#     # delete everything after the first dot
#     mutate(Label = gsub("ICAM1", "ICAM", Label), 
#            Label = gsub("ICAM", "ICAM1", Label), 
#            Label = gsub("\\..*", "", Label), 
#            Loc_X = X, 
#            Loc_Y = Y
#            ) %>%
#     select(Label, Mean, Loc_X, Loc_Y, MELC_ID, Area) %>%
#     # change long format to wide format
#     pivot_wider(names_from = Label, values_from = c(Mean))
#   
#   # filter data for median
#   df_samp_median <- df_samp %>%
#     # delete everything after the first dot
#     mutate(Label = gsub("ICAM1", "ICAM", Label), 
#            Label = gsub("ICAM", "ICAM1", Label), 
#            Label = gsub("\\..*", "", Label), 
#            Loc_X = X, 
#            Loc_Y = Y
#            ) %>%
#     select(Label, Median, Loc_X, Loc_Y, MELC_ID, Area) %>%
#     # change long format to wide format
#     pivot_wider(names_from = Label, values_from = c(Median))
#   
#   df_raw_mean <- data.table::rbindlist(list(df_raw_mean, df_samp_mean), 
#                                        fill = TRUE)
#   df_raw_median <- data.table::rbindlist(list(df_raw_median, df_samp_median), 
#                                          fill = TRUE)
# 
# }

```

# Tidying data

The data is one data frame but in a long format. Names are tidied and the format is changed to the wide format.

```{r}
# change order of columns, so that all markers first, followed by Loc_X and Loc_Y
df_raw_mean <- df_raw_mean %>%
  select(5:length(colnames(df_raw_mean)), 1:4)

df_raw_median <- df_raw_median %>%
  select(5:length(colnames(df_raw_median)), 1:4)

# create individual cell ID
df_raw_mean$CellID <- rownames(df_raw_mean)
df_raw_median$CellID <- rownames(df_raw_median)


```

## Visualize data

Depict nuclei in x and y:

```{r, fig.width=9, fig.height=9}

melced_samples <- unique(df_raw_mean$Experiment_ID)
x_co <- 2048
y_co <- 2048

plot_list <- lapply(melced_samples, function(sample) {
  df_samp <- df_raw_mean %>%
    filter(Experiment_ID == sample)
  
  n_cells <- nrow(df_samp)
  
  ggplot(df_samp, aes(x = Loc_X, y = Loc_Y)) +
    geom_point(size = 1, shape = 20, color = "white") +
    theme_void() +
    theme(legend.title = element_blank()) +
    theme(panel.background = element_rect(fill = 'black', color = 'black', size = 0)) +
    ggtitle(paste(sample, " Objects:", n_cells))
})

ggarrange(plotlist = plot_list, ncol = 2, nrow = 2)
```

## Load meta data

```{r}
df_meta <- read_excel(paste0(data_dir_main, "/GATA3eGFP_sample_list.xlsx"), 
    sheet = "MELC experiments")

df_meta <- df_meta %>%
  mutate(Run = Run_ID, 
         MELC_ID = paste0(Experiment_ID, "_", Condition)) %>%
  as.data.frame()

head(df_meta)
```

Combine metadata and feature data:

```{r}
df_raw_mean <- merge(df_raw_mean, df_meta, by = "Experiment_ID")
df_raw_mean <- df_raw_mean %>%
  select(-Experiment_ID, everything(), Experiment_ID) %>%
  # introduce a column to store the extracted fluorescence information
  mutate(extracted_fluorescence = "mean")

df_raw_median <- merge(df_raw_median, df_meta, by = "Experiment_ID")
df_raw_median <- df_raw_median %>%
  select(-Experiment_ID, everything(), Experiment_ID) %>%
  # introduce a column to store the extracted fluorescence information
  mutate(extracted_fluorescence = "median")

head(df_raw_mean)
head(df_raw_median)

length(unique(df_raw_mean$MELC_ID))
length(unique(df_raw_median$MELC_ID))
```

# Save data

```{r}

# export the merged data frame for mean and median after tidying
write.csv(df_raw_mean, file = paste0(output_env_dir, "/", 
                        # params$extraction_parameter, "_", 
                        # params$mask_parameter, 
                        "df_mean_merged_raw_tidied.csv"))


write.csv(df_raw_median, 
          file = paste0(output_env_dir, "/", 
                        # params$extraction_parameter, "_", 
                        # params$mask_parameter, 
                        "df_median_merged_raw_tidied.csv"))


# save the environment
save.image(paste0(output_env_dir, "/environment.RData"))

sessionInfo()
```
