---
title: "Quality control: marker selection, data transformation, data imputation"
author: "Sandy Kroh"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    fig_caption: yes
    keep_md: yes
  pdf_document:
    toc: yes
params:
  extraction_parameter: cells
  mask_parameter: feature_0
  fluorescence_parameter: median
  Sample: Default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", 
                      fig.align='center', 
                      fig.width = 9, 
                      fig.height = 6,
                      dpi = 600, 
                      set.seed(8),
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Libraries

```{r}
library(here)
library(readr)
library(dplyr)
library(ggridges)
library(ggplot2)
library(ggpubr)
library(readxl)
library(tidyr)

```

## Parameters

```{r}

set.seed(8) 

cols_treat <- c("darkcyan", "gold", "deeppink", "slateblue")  

cols_fov <- c("slateblue1",              
              "darkcyan",                         
              "seagreen2")   

cols_exp <- c("darkcyan", "seagreen2", "lightgreen",     
              "gold", "yellow", "olivedrab1",            
              "deeppink","#FF0066", "magenta",           
              "slateblue", "blueviolet", "purple")   

cols_nat <- c("darkcyan", "seagreen2", "lightgreen",              
              "gold", "yellow", "olivedrab1",              
              "deeppink","#FF0066", "magenta",           
              "slateblue", "blueviolet", "purple",               
              "cyan", "blue", "navyblue",               
              "brown", "red", "orange", "grey") 
```

# Load data

## Directories

```{r}
# define path to general data main directory
data_dir_main <- here::here("data")

# define path to input directory
input_dir <- here::here("1_revisions", "1_Data_loading_merging_tidying_files")

# define output folder for saing the environment etc. 
output_env_dir <- here::here("1_revisions", "2_feature_selection_arcsinh_transformation_median_files") 
dir.create(output_env_dir)

```

## Data

```{r}
df_raw <- read_csv(paste0(input_dir, "/", 
                        # params$extraction_parameter, "_", 
                        # params$mask_parameter, 
                        # "_", 
                        # params$fluorescence_parameter, 
                        "df_median_merged_raw_tidied.csv"), 
    col_types = cols(...1 = col_skip()))

features <- colnames(df_raw[ , c(1:(match("Loc_X", colnames(df_raw)))-1)])
features

unique(df_raw$MELC_ID)
length(unique(df_raw$MELC_ID))

```

# Marker inspection

Define arrangement of plots:

```{r}
library(tidyverse)
library(magick)
library(grid)
library(gridExtra)

# 1. Base path
base_path <- "D:/Repositories/2025_Kroh_et_al/Murine_ILC_niches_lung_SI_IL-33/data/MELC_data/Revision/MELC_images_renamed"

# 2. Extract and organize MELC_IDs
fov_info <- data.frame(MELC_ID = unique(df_raw$MELC_ID)) %>%
  mutate(Derived_Cond = sub(".*_", "", MELC_ID))

# 3. Create the specific column-wise order
ctrl_ids <- fov_info %>% filter(Derived_Cond == "CTRL") %>% pull(MELC_ID)
d3_ids   <- fov_info %>% filter(Derived_Cond == "D3")   %>% pull(MELC_ID)

target_rows <- 1
pad_list <- function(x, len) { c(x, rep(list(NULL), max(0, len - length(x)))) }

col1 <- pad_list(ctrl_ids, target_rows)
col2 <- pad_list(d3_ids[1:4], target_rows)

ordered_ids <- as.vector(t(matrix(c(col2), ncol = 1)))

```

## Single marker overviews

```{r fig.height=8, fig.width=9}

# Define features to iterate through
features_of_interest <- c("CD45", #"CD31", "EMCN", "CD146", "CD105", 
                          # "FN", "Sca1", "PDPN", "COL1", 
                          # # myeloid markers
                          # "CD11b", #pan-myeloid
                          # "SiglecF", #Eosinophils
                          # "CD11c", #DCs
                          # "Ly6G", # Neutrophils
                          # "Ly6C", # Monocytes
                          # "CCR2", # inflammatory Monocytes
                          # "CD68", "F480", # macrophages
                          # #pro-inflammatory
                          # "Mac2", "CD80", "iNOS", "MHCII", 
                          # #antiinflammatory 
                          # "CD206", "CD163", 
                          "GATA3eGFP") 



# 4. Modified loading function: Plots ID even if image is missing
load_fov_image_fixed <- function(melc_id, feature) {
  if (is.null(melc_id)) return(ggplot() + theme_void())
  
  img_path <- file.path(base_path, melc_id, paste0(feature, ".png"))
  
  if (file.exists(img_path)) {
    img <- image_read(img_path)
    img_grob <- rasterGrob(as.raster(img))
    
    ggplot() + 
      annotation_custom(img_grob) + 
      theme_void() +
      labs(title = melc_id) + 
      theme(plot.title = element_text(size = 7, hjust = 0.5, face = "bold"))
  } else {
    # Plot the MELC_ID title over a blank space if image is missing
    ggplot() + 
      theme_void() + 
      labs(title = melc_id) +
      theme(plot.title = element_text(size = 7, hjust = 0.5, face = "bold", color = "black"))
  }
}

# 5. Loop through features and generate grids
for (feat in features_of_interest) {
  
  plot_list <- map(ordered_ids, ~load_fov_image_fixed(.x, feat))
  
  grid_title <- textGrob(paste("Overview", feat), 
                         gp = gpar(fontsize = 14, fontface = "bold"))
  
  final_grid <- grid.arrange(
    grobs = plot_list, 
    ncol = 5, 
    top = grid_title,
    padding = unit(1, "line")
  )
  
  grid.draw(final_grid)
}
```

## Marker overlays

Two color overlay of EMCN and CD31 for Type H-vessels:

```{r eval=FALSE, fig.height=8, fig.width=9, include=FALSE}

# 4. Function for channel combination
load_overlay_image <- function(melc_id, feat_green = "EMCN", feat_red = "CD31") {
  if (is.null(melc_id)) return(ggplot() + theme_void())
  
  path_g <- file.path(base_path, melc_id, paste0(feat_green, ".png"))
  path_r <- file.path(base_path, melc_id, paste0(feat_red, ".png"))
  
  if (file.exists(path_g) & file.exists(path_r)) {
    img_g <- image_read(path_g) %>% image_convert(colorspace = "gray")
    img_r <- image_read(path_r) %>% image_convert(colorspace = "gray")
    
    info <- image_info(img_g)
    img_b <- image_blank(info$width, info$height, color = "black") %>% 
             image_convert(colorspace = "gray")
    
    final_img <- image_join(c(img_r, img_g, img_b)) %>% 
                 image_combine(colorspace = "sRGB")
    
    img_grob <- rasterGrob(as.raster(final_img))
    
    ggplot() + 
      annotation_custom(img_grob) + 
      theme_void() +
      labs(title = melc_id) + 
      theme(plot.title = element_text(size = 7, hjust = 0.5, face = "bold"))
    
  } else {
    ggplot() + 
      theme_void() + 
      labs(title = melc_id) +
      theme(plot.title = element_text(size = 7, hjust = 0.5, face = "bold"))
  }
}

# 5. Generate the Plot List
plot_list <- map(ordered_ids, ~load_overlay_image(.x))

# 6. Create a tightly packed, centered colored title
title_emcn <- textGrob("EMCN ", gp = gpar(fontsize = 16, fontface = "bold", col = "green4"))
title_cd31 <- textGrob("CD31", gp = gpar(fontsize = 16, fontface = "bold", col = "red"))

# Use a gtable or a simple grob tree to center them together
combined_title <- g_title <- frameGrob()
g_title <- packGrob(g_title, title_emcn, side = "left")
g_title <- packGrob(g_title, title_cd31, side = "right")

# 7. Final Grid Arrangement
final_grid <- grid.arrange(
  grobs = plot_list, 
  ncol = 5, 
  top = g_title,
  padding = unit(1, "line")
)

grid.draw(final_grid)
```

Thee color overlay of CD45, DAPI, and CD68:

```{r eval=FALSE, fig.height=8, fig.width=9, include=FALSE}

# 4. Function for 3-channel combination (RGB)
load_overlay_image <- function(melc_id, feat_green = "CD68", 
                               feat_red = "CD45", 
                               feat_blue = "DAPI") {
  if (is.null(melc_id)) return(ggplot() + theme_void())
  
  path_g <- file.path(base_path, melc_id, paste0(feat_green, ".png"))
  path_r <- file.path(base_path, melc_id, paste0(feat_red, ".png"))
  path_b <- file.path(base_path, melc_id, paste0(feat_blue, ".png"))
  
  # Check if all 3 files exist
  if (all(file.exists(c(path_g, path_r, path_b)))) {
    img_g <- image_read(path_g) %>% image_convert(colorspace = "gray")
    img_r <- image_read(path_r) %>% image_convert(colorspace = "gray")
    img_b <- image_read(path_b) %>% image_convert(colorspace = "gray")
    
    # Combine: Red = CD31, Green = EMCN, Blue = DAPI
    final_img <- image_join(c(img_r, img_g, img_b)) %>% 
                 image_combine(colorspace = "sRGB")
    
    img_grob <- rasterGrob(as.raster(final_img))
    
    ggplot() + 
      annotation_custom(img_grob) + 
      theme_void() +
      labs(title = melc_id) + 
      theme(plot.title = element_text(size = 7, hjust = 0.5, face = "bold"))
    
  } else {
    ggplot() + 
      theme_void() + 
      labs(title = melc_id) +
      theme(plot.title = element_text(size = 7, hjust = 0.5, face = "bold"))
  }
}

# 5. Generate the Plot List
plot_list <- map(ordered_ids, ~load_overlay_image(.x))

# 6. Create a tightly packed, centered color-coded title
title_emcn <- textGrob("CD68 ", gp = gpar(fontsize = 16, fontface = "bold", col = "green4"))
title_cd31 <- textGrob("CD45 ", gp = gpar(fontsize = 16, fontface = "bold", col = "red"))
title_dapi <- textGrob("DAPI",  gp = gpar(fontsize = 16, fontface = "bold", col = "blue"))

# Pack them into a single centered unit
g_title <- frameGrob()
g_title <- packGrob(g_title, title_emcn, side = "left")
g_title <- packGrob(g_title, title_cd31, side = "left") # Pack sequentially to the left
g_title <- packGrob(g_title, title_dapi, side = "left")

# 7. Final Grid Arrangement
final_grid <- grid.arrange(
  grobs = plot_list, 
  ncol = 5, 
  top = g_title,
  padding = unit(1, "line")
)

grid.draw(final_grid)
```

```{r eval=FALSE, fig.height=8, fig.width=9, include=FALSE}

# 4. Function for 3-channel combination (RGB)
load_overlay_image <- function(melc_id, feat_green = "FN", 
                               feat_red = "EMCN", 
                               feat_blue = "CD45") {
  if (is.null(melc_id)) return(ggplot() + theme_void())
  
  path_g <- file.path(base_path, melc_id, paste0(feat_green, ".png"))
  path_r <- file.path(base_path, melc_id, paste0(feat_red, ".png"))
  path_b <- file.path(base_path, melc_id, paste0(feat_blue, ".png"))
  
  # Check if all 3 files exist
  if (all(file.exists(c(path_g, path_r, path_b)))) {
    img_g <- image_read(path_g) %>% image_convert(colorspace = "gray")
    img_r <- image_read(path_r) %>% image_convert(colorspace = "gray")
    img_b <- image_read(path_b) %>% image_convert(colorspace = "gray")
    
    # Combine: Red = CD31, Green = EMCN, Blue = DAPI
    final_img <- image_join(c(img_r, img_g, img_b)) %>% 
                 image_combine(colorspace = "sRGB")
    
    img_grob <- rasterGrob(as.raster(final_img))
    
    ggplot() + 
      annotation_custom(img_grob) + 
      theme_void() +
      labs(title = melc_id) + 
      theme(plot.title = element_text(size = 7, hjust = 0.5, face = "bold"))
    
  } else {
    ggplot() + 
      theme_void() + 
      labs(title = melc_id) +
      theme(plot.title = element_text(size = 7, hjust = 0.5, face = "bold"))
  }
}

# 5. Generate the Plot List
plot_list <- map(ordered_ids, ~load_overlay_image(.x))

# 6. Create a tightly packed, centered color-coded title
title_emcn <- textGrob("FN ", gp = gpar(fontsize = 16, fontface = "bold", col = "green4"))
title_cd31 <- textGrob("EMCN ", gp = gpar(fontsize = 16, fontface = "bold", col = "red"))
title_dapi <- textGrob("CD45",  gp = gpar(fontsize = 16, fontface = "bold", col = "blue"))

# Pack them into a single centered unit
g_title <- frameGrob()
g_title <- packGrob(g_title, title_emcn, side = "left")
g_title <- packGrob(g_title, title_cd31, side = "left") # Pack sequentially to the left
g_title <- packGrob(g_title, title_dapi, side = "left")

# 7. Final Grid Arrangement
final_grid <- grid.arrange(
  grobs = plot_list, 
  ncol = 5, 
  top = g_title,
  padding = unit(1, "line")
)

grid.draw(final_grid)
```

Three color composite of endothelial markers:

```{r eval=FALSE, fig.height=8, fig.width=9, include=FALSE}

# 4. Function for 3-channel combination (RGB)
load_overlay_image <- function(melc_id, feat_green = "CD31", 
                               feat_red = "EMCN", 
                               feat_blue = "CD146") {
  if (is.null(melc_id)) return(ggplot() + theme_void())
  
  path_g <- file.path(base_path, melc_id, paste0(feat_green, ".png"))
  path_r <- file.path(base_path, melc_id, paste0(feat_red, ".png"))
  path_b <- file.path(base_path, melc_id, paste0(feat_blue, ".png"))
  
  # Check if all 3 files exist
  if (all(file.exists(c(path_g, path_r, path_b)))) {
    img_g <- image_read(path_g) %>% image_convert(colorspace = "gray")
    img_r <- image_read(path_r) %>% image_convert(colorspace = "gray")
    img_b <- image_read(path_b) %>% image_convert(colorspace = "gray")
    
    # Combine: Red = CD31, Green = EMCN, Blue = DAPI
    final_img <- image_join(c(img_r, img_g, img_b)) %>% 
                 image_combine(colorspace = "sRGB")
    
    img_grob <- rasterGrob(as.raster(final_img))
    
    ggplot() + 
      annotation_custom(img_grob) + 
      theme_void() +
      labs(title = melc_id) + 
      theme(plot.title = element_text(size = 7, hjust = 0.5, face = "bold"))
    
  } else {
    ggplot() + 
      theme_void() + 
      labs(title = melc_id) +
      theme(plot.title = element_text(size = 7, hjust = 0.5, face = "bold"))
  }
}

# 5. Generate the Plot List
plot_list <- map(ordered_ids, ~load_overlay_image(.x))

# 6. Create a tightly packed, centered color-coded title
title_emcn <- textGrob("CD31 ", gp = gpar(fontsize = 16, fontface = "bold", col = "green4"))
title_cd31 <- textGrob("EMCN ", gp = gpar(fontsize = 16, fontface = "bold", col = "red"))
title_dapi <- textGrob("CD146 ",  gp = gpar(fontsize = 16, fontface = "bold", col = "blue"))

# Pack them into a single centered unit
g_title <- frameGrob()
g_title <- packGrob(g_title, title_emcn, side = "left")
g_title <- packGrob(g_title, title_cd31, side = "left") # Pack sequentially to the left
g_title <- packGrob(g_title, title_dapi, side = "left")

# 7. Final Grid Arrangement
final_grid <- grid.arrange(
  grobs = plot_list, 
  ncol = 5, 
  top = g_title,
  padding = unit(1, "line")
)

grid.draw(final_grid)
```

# Feature selection

Here we check which markers will be excluded from the dataset:

```{r, fig.width=12, fig.height=6}
library(data.table)
library(dplyr)
library(ComplexHeatmap)

# Assuming df_raw is your dataframe with columns: MELC_ID (Filename) and features
df_ov <- data.table(df_raw)
df_ov <- df_ov[, lapply(.SD, function(x) sum(is.na(x))), by = list(MELC_ID)]

# Convert integer columns to numeric (if needed)
df_ov <- df_ov %>%
  mutate_if(is.integer, as.numeric)

# Extract Filenames (MELC_ID) for row names
row_names <- df_ov$MELC_ID

# Create the heatmap matrix (exclude MELC_ID column)
heatmap_data <- as.matrix(df_ov[, -"MELC_ID"])

# Assign row names to the matrix
rownames(heatmap_data) <- row_names

# Plot the heatmap with Filenames on the left
Heatmap(
  heatmap_data,
  name = "NAs in dataset",  # Title of legend
  column_title = "Available features", 
  row_title = "Filename",
  row_names_side = "left",  # Display row names on the left
  row_names_gp = gpar(fontsize = 7),  # Adjust text size
  # Optional: Cluster rows/columns if needed
  cluster_rows = FALSE,
  cluster_columns = FALSE
)
```

Exclude the features that are not available in at least 75% of the runs per condition or the nuclei markers etc. that should not be included in the downstream analysis. We will do this for each condition.

First exclude markers that are not relevant anyways and should be excluded:

```{r}
# Add your manual exclusion list
manual_excluded_markers <- c("DAPI")

df_selected <- df_raw %>%
  select(-manual_excluded_markers)
```

```{r}

# Define which columns are “metadata” (they are *not* markers)
meta_cols <- colnames(df_raw[(which(colnames(df_raw) == "Loc_X"):(length(colnames(df_raw))))])

# Get the marker names (everything else in df_raw)
marker_cols <- colnames(df_raw)[1:match("Loc_X",
       colnames(df_raw))-1]

# How many FOVs per condition?
fov_per_cond <- df_raw %>%
  distinct(Condition, MELC_ID) %>%
  group_by(Condition) %>%
  summarise(n_fov = n(), .groups = "drop")

# Count NAs per marker, per Condition/FOV
missing_counts <- df_raw %>%
  # keep only metadata + markers
  select(all_of(c(meta_cols, marker_cols))) %>%
  # for each marker, flag TRUE if NA in that FOV
  mutate(across(all_of(marker_cols), ~ is.na(.), .names = "miss_{col}")) %>%
  # collapse to one row per Condition/MELC_ID (a FOV)
  group_by(Condition, MELC_ID) %>%
  summarise(across(starts_with("miss_"), ~ any(.), .names = "miss_{col}"),
            .groups = "drop") %>%
  # now count how many FOVs miss each marker in each condition
  pivot_longer(cols = starts_with("miss_"),
               names_to = "marker",
               names_prefix = "miss_",
               values_to = "missed") %>%
  group_by(Condition, marker) %>%
  summarise(n_missing = sum(missed),
            n_fov      = n(),
            .groups    = "drop") %>%
  left_join(fov_per_cond, by = c("Condition", "n_fov")) %>%
  mutate(pct_missing = n_missing / n_fov)


# Keep markers that are present in ≥75% of the FOVs of each condition ----
threshold <- 0.25
kept_markers_by_cond <- missing_counts %>%
  group_by(Condition) %>%
  summarise(kept     = list(marker[pct_missing <= threshold]),
            discarded = list(marker[pct_missing  > threshold]),
            .groups   = "drop")


```

# Arcsinh data transformation

## Formula

The following formula[^1] is used to perform inverse hyperbolic arcsinh transformation on the feature values:

[^1]: [Arcsinh transformation](https://mathworld.wolfram.com/InverseHyperbolicSine.html)

    [Arcsinh transformation for flow](https://wiki.centenary.org.au/display/SPECTRE/Data+transformation)

$$ arcsinh(x) = ln( \frac{x}{cofactor}+ \sqrt{1+( \frac{x}{cofactor} )^2} $$

using a cofactor of 0.2.

```{r}
# function used
arcsinh_transform <- function(x, cofactor = 0.002){
  log((x/cofactor)+ sqrt(1+(x/cofactor)^2))
}

```

## **Calculate marker-specific cofactors**

### **Why Calculate Marker-Specific Cofactors?**

-   **Antibody Variability:** Different antibodies have unique binding affinities and background noise levels; a structural marker like LYVE1 may have a baseline fluorescence orders of magnitude higher than a transcription factor like TBET.

-   **Noise Floor Normalization:** A single global cofactor assumes the “noise floor” is identical for every channel, which can lead to over-squashing dim markers (losing signal) or under-squashing bright ones (leaving high background).

-   **Distance-Based Accuracy:** Clustering algorithms (e.g., PhenoGraph or Louvain) rely on Euclidean distance; marker-specific cofactors ensure that a “positive” shift in a dim marker carries the same mathematical weight as a shift in a bright marker.

-   **Dynamic Range Alignment:** It brings all markers into a comparable numeric scale (typically 0–10), preventing high-intensity markers from disproportionately influencing the clustering results.

### **Why Use the 5th Percentile?**

-   **Empirical Noise Estimation:** The bottom 5% of non-zero intensity values generally represents the “background” or “negative” population of cells for that specific staining run.

-   **Defining the Linear-to-Log Transition:** In the arcsinh formula, the cofactor determines the point where the transformation switches from linear (preserving the structure of negative cells) to logarithmic (compressing the positive dynamic range).

-   **Population Separation:** By setting the transition point at the edge of the background noise, you maximize the visual and mathematical “gap” between marker-negative and marker-positive cells.

-   **Outlier Resilience:** Using a low percentile rather than a mean or median ensures that the transformation is sensitive to the actual signal while ignoring tiny amounts of extreme background or technical artifacts.

```{r, fig.width=9, fig.height=14}
library(grid)

# 1. Define features (columns before Loc_X)
features <- colnames(df_selected)[1:(match("Loc_X", colnames(df_selected)) - 1)]

# 2. Calculate Marker-Specific Cofactors
# Start with a complete data frame of all features to ensure none are ever dropped
cofactors_per_marker <- data.frame(Marker = features) %>%
  left_join(
    df_selected %>%
      pivot_longer(cols = all_of(features), names_to = "Marker", values_to = "Value") %>%
      filter(!is.na(Value) & Value > 0) %>%
      group_by(Marker) %>%
      summarise(Cofactor = quantile(Value, 0.01, na.rm = TRUE), .groups = 'drop'),
    by = "Marker"
  ) %>%
  # 🛡️ THE FIX: If a marker (like RORgt) was completely empty, it becomes NA. 
  # Replace it with 1 so the downstream math doesn't crash.
  mutate(Cofactor = replace_na(Cofactor, 1))


# 3. Create the Marker Table and Title separately
marker_table_base <- cofactors_per_marker %>%
  mutate(Cofactor = formatC(Cofactor, format = "e", digits = 3)) %>%
  ggtexttable(
    rows = NULL, 
    theme = ttheme("light", base_size = 11),
    widths = unit(c(1.5, 1.5), "inch") 
  )

title_grob <- text_grob(
  "MELC Marker-Specific Cofactors (5th Percentile)", 
  face = "bold", 
  size = 14
)

# Arrange the title and table vertically using ggarrange
marker_table_final <- ggarrange(
  title_grob, 
  marker_table_base, 
  ncol = 1, 
  heights = c(0.1, 1) # Give the title a small portion of the vertical space
)

# Print the table
print(marker_table_final)
```

## Transformation

```{r, fig.width=9, fig.height=14}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(tibble) # Required for deframe()


# 4. Apply Arcsinh Transformation with specific cofactors
# Convert table to named vector for lookup
marker_cofactors_vec <- deframe(cofactors_per_marker)

df_trans_selected <- df_selected %>%
  mutate(across(all_of(features), 
                ~ asinh(.x / marker_cofactors_vec[[cur_column()]])))


# 5. Prepare data for "Before vs After" visualization
df_feat_before <- df_selected %>%
  select(all_of(features)) %>%
  pivot_longer(cols = everything(), names_to = "Feature", values_to = "Value")

df_trans_stacked <- df_trans_selected %>%
  select(all_of(features)) %>%
  pivot_longer(cols = everything(), names_to = "Feature", values_to = "Value")

# 6. Create Comparison Plots
before_plot <- df_feat_before %>%
  ggplot(aes(x = Value, y = Feature, fill = Feature)) +
  geom_density_ridges(na.rm = TRUE) +
  theme_ridges() +
  theme(legend.position = "none") +
  labs(title = "Raw Intensities (Linear Scale)")

after_plot <- df_trans_stacked %>%
  ggplot(aes(x = Value, y = Feature, fill = Feature)) +
  geom_density_ridges(na.rm = TRUE) +
  theme_ridges() +
  theme(legend.position = "none") +
  geom_vline(color = "deeppink", xintercept = c(4, 8), linetype = "dashed") +
  labs(title = "Arcsinh Transformed (Marker-Specific)")

# Display plots side-by-side
ggarrange(before_plot, after_plot, ncol = 2)
```

# Data imputation

# Thresholding

```{r eval=FALSE, include=FALSE}
# define threshold 
threshold <- 5

df_trans %>%
  dplyr::mutate( max_features =  apply( df_trans[, -((match("Loc_X",
         colnames(df_trans))):length(colnames(
           df_trans)))], 1, max, na.rm = TRUE ), 
         Intensity = "Intensity") %>%
  ggplot(aes(x = max_features, y = Intensity)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none") + 
    geom_vline(color = "deeppink", xintercept = c(4, 8)) + 
  theme(legend.position = "none") + 
    geom_vline(color = "gold", xintercept = c(threshold))


# filter cells that do express at least one feature
df_trans_filtered <- df_trans %>% 
  dplyr::mutate( max_features =  apply( df_trans[, -((match("Loc_X",
         colnames(df_trans))):length(colnames(
           df_trans)))], 1, max, na.rm = TRUE )) %>%
  # filter for cells that have a higher max feature value than 0
  filter(max_features > threshold) %>%
  select(-max_features)
  

# filter cells that do not express any feature
df_trans_others <- df_trans %>% 
  dplyr::mutate( max_features =  apply( df_trans[, -((match("Loc_X",
         colnames(df_trans))):length(colnames(
           df_trans)))], 1, max, na.rm = TRUE )) %>%
  # filter for cells that have a higher max feature value than 0
  filter(max_features <= threshold) %>%
  select(-max_features)

nrow(df_trans_others)
nrow(df_trans_filtered)
```

# Data overview

We will plot an overview table to see how many cells, samples and acquired FOVs per conditions are included in the dataset:

```{r fig.height=2, fig.width=9}
library(tidyverse)
library(ggpubr)

# 1. Define a helper function to extract stats from each dataframe
get_stats <- function(df, condition_name) {
  # Identify feature columns (everything before Loc_X)
  loc_x_idx <- which(colnames(df) == "Loc_X")
  n_features <- loc_x_idx - 1
  
  data.frame(
    Condition = condition_name,
    Cells = nrow(df),
    Features = n_features,
    Samples = n_distinct(df$Sample),
    FOVs = n_distinct(df$MELC_ID)
  )
}

# 2. Combine stats for all three conditions
summary_df <- bind_rows(
  get_stats(df_trans_selected, unique(df_trans_selected$Condition)))

# 3. Create the Table Grob
summary_table_base <- summary_df %>%
  ggtexttable(
    rows = NULL, 
    theme = ttheme("light", base_size = 11),
    # Adjusting widths for the 5 columns
    widths = unit(c(1.2, 1, 1, 1, 1), "inch") 
  )

# 4. Create the Title Grob
title_grob <- text_grob(
  "Proteomics Dataset Overview: Murine Femur Osteotomy", 
  face = "bold", 
  size = 14
)

# 5. Arrange the title and table vertically
summary_table_final <- ggarrange(
  title_grob, 
  summary_table_base, 
  ncol = 1, 
  heights = c(0.15, 1) 
)

# 6. Display
print(summary_table_final)
```

We will plot the markers that are common between the different conditions:

```{r eval=FALSE, fig.height=2, fig.width=9, include=FALSE}
library(tidyverse)
library(ggpubr)

# 1. Identify feature columns for each dataframe (everything before Loc_X)
features_ctrl <- colnames(df_trans_ctrl_imputed)[1:(which(colnames(df_trans_ctrl_imputed) == "Loc_X") - 1)]
features_d4   <- colnames(df_trans_d4_imputed)[1:(which(colnames(df_trans_d4_imputed) == "Loc_X") - 1)]
features_d8   <- colnames(df_trans_d8_imputed)[1:(which(colnames(df_trans_d8_imputed) == "Loc_X") - 1)]

# 2. Find the intersection (common markers)
common_features <- intersect(intersect(features_ctrl, features_d4), features_d8)

# 3. Format the list into a multi-column table for better visualization
# We'll wrap the vector into a matrix/dataframe with, say, 4 columns
n_cols <- 10
n_rows <- ceiling(length(common_features) / n_cols)

# Pad with empty strings to make it a perfect rectangle
feature_matrix <- c(common_features, rep("", (n_rows * n_cols) - length(common_features)))
feature_df <- as.data.frame(matrix(feature_matrix, ncol = n_cols, byrow = TRUE))
colnames(feature_df) <- paste0("Col_", 1:n_cols)

# 4. Create the Table Grob
feature_table_base <- feature_df %>%
  ggtexttable(
    rows = NULL, 
    cols = NULL)


# 5. Create the Title Grob
feature_title <- text_grob(
  paste0("Common Markers Across All Conditions (n = ", length(common_features), ")"), 
  face = "bold", 
  size = 14
)

# 6. Arrange and Print
common_features_plot <- ggarrange(
  feature_title, 
  feature_table_base, 
  ncol = 1, 
  heights = c(0.15, 1)
)

print(common_features_plot)
```

Depict all common markers for each FOV:

```{r eval=FALSE, fig.height=6, fig.width=9, include=FALSE}
library(tidyverse)
library(magick)
library(grid)
library(gridExtra)

# 1. Configuration
base_path <- "D:/Repositories/Osteotomy_MELC/data/images_renamed"
# Using the common_features defined in your previous steps
markers_to_plot <- common_features 

# 2. Get unique FOVs (e.g., from your df_raw)
unique_fovs <- unique(df_raw$MELC_ID)

# 3. Helper function to load single grayscale markers
load_grayscale_marker <- function(melc_id, marker) {
  img_path <- file.path(base_path, melc_id, paste0(marker, ".png"))
  
  if (file.exists(img_path)) {
    # Read and ensure it is treated as grayscale
    img <- image_read(img_path) %>% image_convert(colorspace = "gray")
    img_grob <- rasterGrob(as.raster(img))
    
    ggplot() + 
      annotation_custom(img_grob) + 
      theme_void() +
      labs(title = marker) + 
      theme(plot.title = element_text(size = 8, hjust = 0.5, face = "bold"))
  } else {
    # If file is missing, plot the ID with a blank space
    ggplot() + 
      theme_void() + 
      labs(title = marker) +
      theme(plot.title = element_text(size = 8, hjust = 0.5, face = "italic", color = "grey50"))
  }
}

# 4. Loop through each FOV to generate the full panel
for (fov in unique_fovs) {
  
  # Generate the list of plots for all common markers for THIS fov
  fov_plots <- map(markers_to_plot, ~load_grayscale_marker(fov, .x))
  
  # Calculate grid title
  grid_title <- textGrob(paste("Full Marker Panel | FOV:", fov), 
                         gp = gpar(fontsize = 15, fontface = "bold"))
  
  # Arrange into a 7-column grid
  # This will handle ~43 markers in roughly 6-7 rows
  fov_panel <- grid.arrange(
    grobs = fov_plots, 
    ncol = 7, 
    top = grid_title,
    padding = unit(0.5, "line")
  )
  
  # Draw to the active device
  grid.draw(fov_panel)
  
  # Optional: Pause between FOVs if viewing in RStudio
  # readline(prompt="Press [enter] to continue to next FOV")
}
```

Get an overview of the additional markers available per condition:

```{r eval=FALSE, fig.height=3, fig.width=9, include=FALSE}
library(tidyverse)
library(ggpubr)

# 1. Identify additional features for each condition (relative to common_features)
get_additional <- function(df, common) {
  all_feats <- colnames(df)[1:(which(colnames(df) == "Loc_X") - 1)]
  setdiff(all_feats, common)
}

add_ctrl <- get_additional(df_trans_ctrl_imputed, common_features)
add_d4   <- get_additional(df_trans_d4_imputed, common_features)
add_d8   <- get_additional(df_trans_d8_imputed, common_features)

# 2. Align the lists for the table
max_len <- max(length(add_ctrl), length(add_d4), length(add_d8))

# Function to pad with empty strings and sort for neatness
format_add <- function(x, max_l) {
  if(length(x) == 0) return(rep("", max_l))
  c(sort(x), rep("", max_l - length(x)))
}

# Use "Day 4" and "Day 8" as column names
additional_df <- data.frame(
  Control = format_add(add_ctrl, max_len),
  `Day 4` = format_add(add_d4, max_len),
  `Day 8` = format_add(add_d8, max_len),
  check.names = FALSE # Prevents R from converting spaces to dots
)

# 3. Create the Table Grob
table_base <- additional_df %>%
  ggtexttable(
    rows = NULL,
    theme = ttheme("light", base_size = 12)
  )

# 4. Create the Title Grob
title_grob <- text_grob(
  "Additional Features per Condition", 
  face = "bold", 
  size = 14
)

# 5. Arrange the title and table vertically
additional_table_final <- ggarrange(
  title_grob, 
  table_base, 
  ncol = 1, 
  heights = c(0.15, 1) 
)

# 6. Print the final formatted table
print(additional_table_final)
```

```{r eval=FALSE, fig.height=8, fig.width=9, include=FALSE}
library(tidyverse)
library(magick)
library(grid)
library(gridExtra)

# Define features to iterate through
features_of_interest <- c("CD45", "CD31", "EMCN", "CD146", "CD105", 
                          "FN", "Sca1", "PDPN", "COL1", 
                          # myeloid markers
                          "CD11b", #pan-myeloid
                          "SiglecF", #Eosinophils
                          "CD11c", #DCs
                          "Ly6G", # Neutrophils
                          "Ly6C", # Monocytes
                          "CCR2", # inflammatory Monocytes
                          "CD68", "F480", # macrophages
                          #pro-inflammatory
                          "Mac2", "CD80", "iNOS", "MHCII", 
                          #antiinflammatory 
                          "CD206", "CD163", "CX3CR1") 

# 1. Base path
base_path <- "D:/Repositories/Osteotomy_MELC/data/images_renamed"

# 2. Extract and organize MELC_IDs from df_raw
fov_info <- data.frame(MELC_ID = unique(df_raw$MELC_ID)) %>%
  mutate(
    Derived_Cond = sub(".*_", "", MELC_ID) 
  )

# 3. Create the specific column-wise order
ctrl_ids <- fov_info %>% filter(Derived_Cond == "CTRL") %>% pull(MELC_ID)
d4_ids   <- fov_info %>% filter(Derived_Cond == "D4")   %>% pull(MELC_ID)
d8_ids   <- fov_info %>% filter(Derived_Cond == "D8")   %>% pull(MELC_ID)

target_rows <- 4
pad_list <- function(x, len) { c(x, rep(list(NULL), max(0, len - length(x)))) }

col1 <- pad_list(ctrl_ids, target_rows)
col2 <- pad_list(d4_ids[1:4], target_rows)
col3 <- pad_list(d4_ids[5:8], target_rows)
col4 <- pad_list(d8_ids[1:4], target_rows)
col5 <- pad_list(d8_ids[5:8], target_rows)

ordered_ids <- as.vector(t(matrix(c(col1, col2, col3, col4, col5), ncol = 5)))

# 4. Modified loading function: Plots ID even if image is missing
load_fov_image_fixed <- function(melc_id, feature) {
  if (is.null(melc_id)) return(ggplot() + theme_void())
  
  img_path <- file.path(base_path, melc_id, paste0(feature, ".png"))
  
  if (file.exists(img_path)) {
    img <- image_read(img_path)
    img_grob <- rasterGrob(as.raster(img))
    
    ggplot() + 
      annotation_custom(img_grob) + 
      theme_void() +
      labs(title = melc_id) + 
      theme(plot.title = element_text(size = 7, hjust = 0.5, face = "bold"))
  } else {
    # Plot the MELC_ID title over a blank space if image is missing
    ggplot() + 
      theme_void() + 
      labs(title = melc_id) +
      theme(plot.title = element_text(size = 7, hjust = 0.5, face = "bold", color = "black"))
  }
}

# 5. Loop through features and generate grids
for (feat in features_of_interest) {
  
  plot_list <- map(ordered_ids, ~load_fov_image_fixed(.x, feat))
  
  grid_title <- textGrob(paste("Overview", feat), 
                         gp = gpar(fontsize = 14, fontface = "bold"))
  
  final_grid <- grid.arrange(
    grobs = plot_list, 
    ncol = 5, 
    top = grid_title,
    padding = unit(1, "line")
  )
  
  grid.draw(final_grid)
}
```

# Save data

```{r}
# selected marker data ------------------------------------------------
write.csv(df_selected, file = paste0(
  output_env_dir, "/", params$extraction_parameter, "_",
  params$fluorescence_parameter, "_",
  tolower(unique(df_selected$Condition)),
  "_selected.csv"))

# arcsinh transformed data --------------------------------------------
write.csv(df_trans_selected, file = paste0(
  output_env_dir, "/", params$extraction_parameter, "_",
  params$fluorescence_parameter, "_",
  tolower(unique(df_trans_selected$Condition)),
  "_selected_arcsinh.csv"))

```

## Session info and environment

```{r}
# save the environment
save.image(paste0(output_env_dir, "/environment.RData"))

sessionInfo()
```
