---
title: "Figure 7: ILC2s localize together with lymphatics and myeloid cells"
author: "Sandy Kroh"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    fig_caption: yes
    keep_md: yes
  pdf_document:
    toc: yes
params:
  Sample: Default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", 
                      fig.align='center', dpi = 600, 
                      set.seed(8),
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Libraries

```{r}
library(SeuratObject)
library(dplyr)
library(rstatix)

# remove.packages("rlang")
# remove.packages("dplyr")
# install.packages("rlang")
# install.packages("dplyr")
library(rlang)
library(dplyr)

if (!requireNamespace("Giotto", quietly = TRUE))
  devtools::install_github("drieslab/Giotto@suite")
if (!requireNamespace("VoltRon", quietly = TRUE))
#  devtools::install_github("Artur-man/VoltRon")
  devtools::install_github("BIMSBbioinfo/VoltRon@dev")

if (!requireNamespace("Seurat", quietly = TRUE))
  install.packages("Seurat")
library(Giotto)
library(Seurat)
library(VoltRon)
library(ggplot2)
library(ggpubr)
library(readr)
library(ggbeeswarm)
library(stringr)
library(tidyr)
library(patchwork)
library(ggplotify)
library(ComplexHeatmap)
library(circlize)

```

## Parameters

```{r}
set.seed(123)

input_dir <- here::here("1_data_tidying", "Lung_SI_all_cells_all_ALs_files")

output_dir <- here::here("2_visualizations_for_figures", "Fig_7_spatial_analysis_ILC2s_lung_files")
dir.create(output_dir)




main_markers <- c(
  "EpCAM", "EMCN", "LYVE1", "PDPN", "PDGFRa", "CD8a", "CD4",
  "CD45", "CD3", "IRF4", "Kappa", "CD11c", "CD127", "GATA3eGFP", "RORgt"
)


immune_markers <- c(
 "CD3", "CD4", "CD8a", "Kappa", "IRF4", "CD11c",
  "CD127", "CD90", "EOMES", "GATA3eGFP", "RORgt", "Ki67",  "KLRG1", "NKp46", "CD117", "Areg", "CCR6", "CD44", "MHCII", "Sca1"
)

ilc_markers <- c(
  "CD3", "CD4", "CD8a",
  "CD127", "CD90", "EOMES", "GATA3eGFP", "RORgt", "KLRG1", "NKp46", "CD117", "CCR6", "MHCII", "Ki67", "Areg", "IRF4", "Sca1", "CD44"
)


cols_nat <- c("magenta", "cyan", "blue", "purple", "green", 
                       "red", "yellow", "olivedrab1", "slateblue1", 
                       "darkcyan", "gold","indianred1", "seagreen", "deeppink", 
                       "orange", "brown", "violet",
                       "deeppink4", "pink", 
                       "grey", "black", "lightgreen", 
                       "#FF0066",  
                       "lightblue", "#FFCC99", "#CC00FF", 
                       "blueviolet",  "goldenrod4", 
                       "navy", "olivedrab", "lightcyan", "seagreen2", "darkviolet", "lightpink", "slateblue4", "olivedrab2")

colfunc <- colorRampPalette(c("darkcyan", "green", "yellow", "magenta", "purple"))

cols_ilcs_lung <- c("darkcyan", "seagreen2", "deeppink4")

cols_treat <- c("CTRL"  = "darkcyan", 
              "1"  = "gold",
              "2"  = "deeppink", 
              "3"  = "slateblue", 
              "D1"  = "gold",
              "D2"  = "deeppink", 
              "D3"  = "slateblue")


# colorblind-friendly palette for conditions
cols_con <- cols_treat
ColorsCellType_conditions <- cols_con

cols_heat <- c(
               "#648FFF", "white", "#FFB000")


order_al3 <- c(
      "NK cells/ILC1s",
    "ILC2s",
    "ILC3s",
    "T cytotox cells",
    "T helper cells",
    "B cells & Plasma cells",
    "Myeloid cells",
    "LYVE1 CD31 vessels",
    "LYVE1 CD90 Lymphatics",
    "EMCN CD31 Blood vessels",
    "Epithelia"

)

ColorsCellType <-  list(`NK cells/ILC1s` = "darkcyan", 
                        `ILC2s` = "seagreen2", 
                        `ILC3s` = "darkmagenta", 
                        `T helper cells` = "deeppink",
                        `T cytotox cells` = "slateblue", 
                        `Myeloid cells` = "burlywood", 
                        `B cells & Plasma cells` = "indianred1",
                        `LYVE1 CD31 vessels` = "darkgreen", 
                        `LYVE1 CD90 Lymphatics` = "yellow", 
                        `EMCN CD31 Blood vessels` = "red", 
                        `Epithelia` = "green")

celltypes <- c(
  "NK cells/ILC1s", 
  "ILC2s", 
  "ILC3s", 
  "T helper cells", 
  "T cytotox cells", 
  "Myeloid cells", 
  "B cells & Plasma cells", 
  "LYVE1 CD31 vessels", 
  "LYVE1 CD90 Lymphatics", 
  "EMCN CD31 Blood vessels", 
  "Epithelia"
)

```

# Load data

```{r, fig.width=9, fig.height=6}

# from import_Giotto.Rmd
gio_list <- readRDS(here::here("data", "Giotto_data_lung.rds"))

# from import_VoltRon.Rmd
vr_list <- readRDS(here::here("data", "VoltRon_data_lung.rds"))

# original data
metadatax <- read_csv(here::here("data", "MELC_data_murine_lung_CTRL_D1_D2_D3_withfolders.csv"))
metadatax <- metadatax %>%
  mutate(CellType = AL3)  %>%
  filter(Sample != "20210906_3_lu_d3")

unique(metadatax$CellType)

vr_list_names <- unique(metadatax$Sample)


cell_proximities_list <- list()
for(samp in vr_list_names){
  print(samp)
  cell_proximities_list[[samp]] <-cellProximityEnrichment(
    gobject = gio_list[[samp]],
    cluster_column = 'CellType',
    spatial_network_name = 'Delaunay_network',
    adjust_method = 'fdr',
    number_of_simulations = 1000)
  cell_proximities_list[[samp]] <- cell_proximities_list[[samp]]$enrichm_res
}

vr_merged <- merge(vr_list[[1]], vr_list[-1])
vrImageNames(vr_merged)
unique(vr_merged$CellType)

```

Calculate co-enrichment scores and plot them:

```{r, fig.height=4, fig.width=9}

# FOVs for representative overview images
fovs <- vr_list_names

set_alpha <- 0.35
set_ptsize <- 2
set_nrows <- 1
set_ncols <- 3
cell_shape <- 20


cols_con <- c("darkcyan", "gold", "deeppink", 
                "slateblue")

cols_fov <- c("darkcyan", "gold", "deeppink", 
                "slateblue")

ColorsCellTypeSingle <-  list(
  #`NK cells/ILC1s/ILC3s` = "cyan", 
  `ILC2s` = "magenta",
  #`ILC3s` = "magenta", 
  `EMCN CD31 Blood vessels` = "green")

uni_celltypes <- unique(vr_merged$CellType)
backgroundlist <- list("EpCAM","CD31","LYVE1","LYVE1","CD11c","B220","EOMES", "CD8a", "CD4", "GATA3eGFP", "RORgt")
names(backgroundlist) <- uni_celltypes
uni_celltypes <- uni_celltypes[!uni_celltypes %in% "ILC2s"]

g_master_list <- list()
for(unic in uni_celltypes){
  
    ### selected cell groups ####
  selected_celltypes <- c("ILC2s", unic)
  # interactions <- c("EMCN CD31 Blood vessels--ILC2s")
  interactions <- unique(cell_proximities_list[[vr_list_names[30]]]$unified_int)
  interactions <- interactions[grepl("ILC2s", interactions) & grepl(unic, interactions)]
  background_image <- backgroundlist[[unic]]
  if(length(interactions) > 0){
     ### get interaction results ####
    interaction_celltypes <- NULL
    for(samp in vr_list_names){
      cur_cell_proximities <- cell_proximities_list[[samp]]
      cur_cell_proximities <- cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,]
      sample <- unique(metadatax$FullInfo[metadatax$Sample==samp])
      if(nrow(cur_cell_proximities) > 0 & sample != "20210906_FOV3_D3"){
        interaction_celltypes <- rbind(interaction_celltypes,
                                       data.frame(cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,], 
                                                  experiment = strsplit(sample, split = "_")[[1]][1], fov = strsplit(sample, split = "_")[[1]][2], condition = strsplit(sample, split = "_")[[1]][3]))
      }
    }
    interaction_celltypes$p.adj <- ifelse(interaction_celltypes$enrichm > 0, interaction_celltypes$p.adj_higher, interaction_celltypes$p.adj_lower)


    
    sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("*"), ""))

    g_test <- ggplot(interaction_celltypes, aes(x = condition, y = enrichm, fill = condition)) +
      geom_bar(stat = "identity", position = position_dodge2(width=0.9, preserve = "single")) +
      facet_grid(.~condition, scales = "free_x") +
      geom_text(aes(label=sig_label), position=position_dodge2(width=0.9, preserve = "single"), angle = 90, hjust = -0.02, size = 4) +
      ylim(-2,3.5)+
      NoLegend()+
      theme_classic2()+
      scale_fill_manual(values = cols_fov, name = "") +
      theme(axis.text.x = element_text(#angle = 50,
                                       vjust = 1, size = 12, hjust = 0.5, face = "bold"
                                       ),
            axis.text.y = element_text(hjust = 0.5, size = 12),
            axis.title.x = element_blank(),
            axis.title.y = element_text(size = 12),
            plot.title = element_text(size =14, hjust = 0.5),
            plot.margin = margin(0, 0.5, 0.5, 0.5, "cm"),
            legend.position = "none",
            strip.background=element_blank(),
            strip.background.x= element_blank(),
            strip.text.x = element_text(size = 1, color = "white"),
            panel.grid.major.y = element_line())+
      NoLegend()+  
      ggtitle(gsub("LYVE1 CD90 ", "", interactions)) +
      ylab("Enrichment")
    
    
    g_master_list[[unic]] <- g_test+
        theme(plot.margin = margin(0, 0.5, 0, 0.5, "cm"))
  }
}

# ILC2s around ILC2s -----------------------------------------
interactions <- c("ILC2s--ILC2s")
unic <- "ILC2s"
background_image <- backgroundlist[["ILC2s"]]
interaction_celltypes <- NULL
for(samp in vr_list_names){
    cur_cell_proximities <- cell_proximities_list[[samp]]
    cur_cell_proximities <- cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,]
    sample <- unique(metadatax$FullInfo[metadatax$Sample==samp])
    if(nrow(cur_cell_proximities) > 0 & sample != "20210906_FOV3_D3"){
      interaction_celltypes <- rbind(interaction_celltypes,
                                     data.frame(cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,], 
                                                experiment = strsplit(sample, split = "_")[[1]][1], fov = strsplit(sample, split = "_")[[1]][2], condition = strsplit(sample, split = "_")[[1]][3]))
    }
  }
interaction_celltypes$p.adj <- ifelse(interaction_celltypes$enrichm > 0, interaction_celltypes$p.adj_higher, interaction_celltypes$p.adj_lower)
  # plot test results
  # sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("p=",round(interaction_celltypes$p.adj,3)), ""))
  sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("*"), ""))
    # print(sig_label)
g_test <- ggplot(interaction_celltypes, aes(x = condition, y = enrichm, fill = condition)) +
  geom_bar(stat = "identity", position = position_dodge2(width=0.9, preserve = "single")) +
  facet_grid(.~condition, scales = "free_x") +
  geom_text(aes(label=sig_label), position=position_dodge2(width=0.9, preserve = "single"), angle = 90, hjust = -0.02, size = 4) +
  ylim(-2,3.5)+
  NoLegend()+
  theme_classic2()+
  scale_fill_manual(values = cols_fov, name = "") +
  theme(axis.text.x = element_text(#angle = 50,
                                   vjust = 1, size = 12, hjust = 0.5, face = "bold"
                                   ),
        axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        plot.margin = margin(0, 0.5, 0.5, 0.5, "cm"),
        legend.position = "none",
        strip.background=element_blank(),
        strip.background.x= element_blank(),
        strip.text.x = element_text(size = 1, color = "white"),
        panel.grid.major.y = element_line())+
  NoLegend()+  
  ggtitle(interactions) +
  ylab("Enrichment")


g_master_list[[unic]] <- g_test+
    theme(plot.margin = margin(0, 0.5, 0, 0.5, "cm"))

```

# IF overlays

```{r fig.height=3.6, fig.width=9}
library(png)
library(grid)
library(ggplot2)

# 1. Define the path
img_path <- "D:/Repositories/2025_Kroh_et_al/Murine_ILC_niches_lung_SI_IL-33/data/images/ILC2s_in_myeloid_lymphatic_niche.png"

# 2. Read the image and get dimensions
img <- readPNG(img_path)
h <- nrow(img)
w <- ncol(img)

# 3. Create the ggplot object
# We use coord_fixed to ensure the image doesn't stretch
img_plot_if_ilc2s <- ggplot() +
  annotation_raster(img, xmin = 0, xmax = w, ymin = 0, ymax = h) +
  coord_fixed() +
  theme_void() +
  # Optional: adds a tiny margin to ensure the edges aren't clipped
  scale_x_continuous(expand = c(0, 0), limits = c(0, w)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, h)) +
  theme(plot.margin = margin(0, 0, 0, 0.7, "cm"))

# Display the plot
print(img_plot_if_ilc2s)
```

```{r fig.height=3.6, fig.width=9}
library(png)
library(grid)
library(ggplot2)

# 1. Define the path
img_path <- "D:/Repositories/2025_Kroh_et_al/Murine_ILC_niches_lung_SI_IL-33/data/images/ILC2s_in_myeloid_lymphatic_niche_ROI.png"

# 2. Read the image and get dimensions
img <- readPNG(img_path)
h <- nrow(img)
w <- ncol(img)

# 3. Create the ggplot object
# We use coord_fixed to ensure the image doesn't stretch
img_plot_if_ilc2s_roi <- ggplot() +
  annotation_raster(img, xmin = 0, xmax = w, ymin = 0, ymax = h) +
  coord_fixed() +
  theme_void() +
  # Optional: adds a tiny margin to ensure the edges aren't clipped
  scale_x_continuous(expand = c(0, 0), limits = c(0, w)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, h)) +
  theme(plot.margin = margin(0, 0, 0, 0.7, "cm"))

# Display the plot
print(img_plot_if_ilc2s_roi)
```

```{r fig.height=4.5, fig.width=9}
library(png)
library(grid)
library(ggplot2)

# 1. Define the path
img_path <- "D:/Repositories/2025_Kroh_et_al/Murine_ILC_niches_lung_SI_IL-33/data/images/ILC2s_in_myeloid_lymphatic_niche_ILC2s_myeloid_cells.png"

# 2. Read the image and get dimensions
img <- readPNG(img_path)
h <- nrow(img)
w <- ncol(img)

# 3. Create the ggplot object
# We use coord_fixed to ensure the image doesn't stretch
img_plot_if_ilc2s_my <- ggplot() +
  annotation_raster(img, xmin = 0, xmax = w, ymin = 0, ymax = h) +
  coord_fixed() +
  theme_void() +
  # Optional: adds a tiny margin to ensure the edges aren't clipped
  scale_x_continuous(expand = c(0, 0), limits = c(0, w)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, h)) +
  theme(plot.margin = margin(0, 0, 0, 0.7, "cm"))

# Display the plot
print(img_plot_if_ilc2s_my)
```

# Coenrichment analysis

```{r}
# FOVs for representative overview images
fovs <- vr_list_names

set_alpha <- 0.35
set_ptsize <- 2
set_nrows <- 1
set_ncols <- 3
cell_shape <- 20


cols_con <- c("darkcyan", "gold", "deeppink", 
                "slateblue")

cols_fov <- c("darkcyan", "gold", "deeppink", 
                "slateblue")

ColorsCellTypeSingle <-  list(
  #`NK cells/ILC1s/ILC3s` = "cyan", 
  `ILC2s` = "magenta",
  #`ILC3s` = "magenta", 
  `EMCN CD31 Blood vessels` = "green")

uni_celltypes <- unique(vr_merged$CellType)
backgroundlist <- list("EpCAM","CD31","LYVE1","LYVE1","CD11c","B220","EOMES", "CD8a", "CD4", "GATA3eGFP", "RORgt")
names(backgroundlist) <- uni_celltypes
uni_celltypes <- uni_celltypes[!uni_celltypes %in% "ILC2s"]

```

## Coenrichment heatmaps

```{r, fig.height=2.8, fig.width=9}

# --- 1. Gather all data into one master dataframe ---
all_interactions <- data.frame()

for (vr_name in vr_list_names) {
  
  cur_cell_proximities <- cell_proximities_list[[vr_name]]
  
  # Skip if this FOV has no interaction data
  if(is.null(cur_cell_proximities) || nrow(cur_cell_proximities) == 0) next
  
  # Get metadata for condition
  vr_samp <- subset(vr_merged, samples = vr_name)
  vr_meta <- Metadata(vr_samp)
  condition <- paste(unique(vr_meta$Condition), collapse = ", ")
  
  # Create a temporary dataframe for this FOV
  temp_df <- data.frame(
    Interaction = cur_cell_proximities$unified_int,
    Enrichment = cur_cell_proximities$enrichm,
    FOV = vr_name,
    Condition = condition
  )
  
  # Append to master dataframe
  all_interactions <- rbind(all_interactions, temp_df)
}

# --- 2. Reshape data into a Wide Matrix ---
wide_data <- all_interactions %>%
  filter(grepl("ILC2s", Interaction)) %>%
  select(Interaction, FOV, Enrichment) %>%
  mutate(Interaction = gsub("ILC2s--|--ILC2s", "", Interaction)) %>%
  pivot_wider(names_from = FOV, values_from = Enrichment, values_fill = 0, values_fn = mean) 

# Convert to matrix and assign rownames
enrichmat_global <- as.matrix(wide_data[, -1])
rownames(enrichmat_global) <- wide_data$Interaction

# --- 3. Create Column Annotations (Condition mapping) ---
anno_df <- unique(all_interactions[, c("FOV", "Condition")])
rownames(anno_df) <- anno_df$FOV

# Reorder the rows
anno_df <- anno_df[colnames(enrichmat_global), , drop = FALSE]

condition_colors <- cols_treat

ha_top <- HeatmapAnnotation(
  Condition = anno_df$Condition,
  col = list(Condition = condition_colors),
  annotation_name_side = "left"
)

# --- 4. Draw the Unified Master Heatmap ---
master_heatmap <- Heatmap(
  enrichmat_global,
  
  col = circlize::colorRamp2(c(-1, 0, 1), cols_heat),
  name = "Enrichment\nScore",
  heatmap_legend_param = list(direction = "vertical"),
  
  column_split = anno_df$Condition,
  cluster_column_slices = FALSE, 
  cluster_columns = FALSE,        
  cluster_rows = TRUE,           
  
  # Y-Axis Label
  row_title = "Interactions",
  row_title_gp = gpar(fontsize = 11#, fontface = "bold"
                      ),

  # Aesthetics
  show_column_names = FALSE,
  column_names_rot = 45,
  column_names_gp = gpar(fontsize = 8),
  row_names_gp = gpar(fontsize = 9),
  show_row_dend = TRUE,
  
  # Annotations and Titles
  top_annotation = ha_top,
  column_title = "Neighborhood Enrichment ILC2s",
  column_title_gp = gpar(fontsize = 12, fontface = "bold")
)

# --- 5. Convert to ggplot2 and explicitly position the X-axis ---

# First, capture the raw heatmap (with normal, minimal padding)
heat_grob <- grid::grid.grabExpr(
  draw(master_heatmap, 
       heatmap_legend_side = "right",     
       annotation_legend_side = "right",  
       merge_legend = TRUE,
       padding = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
)

# Convert to a standard ggplot object
gg_heat <- as.ggplot(heat_grob)

# Add the X-axis label and manually nudge it to center under the heatmap body
final_plot_ILC2s <- gg_heat + 
  labs(x = "FOV") + 
  theme(
    axis.title.x = element_text(
      size = 11, 
      # face = "bold", 
      hjust = 0.37 
    ), 
    plot.margin = margin(0, 0, 0.5, 0.2, "cm")
  )

# Print your beautiful, perfectly aligned plot!
print(final_plot_ILC2s)
```

```{r, fig.height=2.8, fig.width=9}

# --- 1. Gather all data into one master dataframe ---
all_interactions <- data.frame()

for (vr_name in vr_list_names) {
  
  cur_cell_proximities <- cell_proximities_list[[vr_name]]
  
  # Skip if this FOV has no interaction data
  if(is.null(cur_cell_proximities) || nrow(cur_cell_proximities) == 0) next
  
  # Get metadata for condition
  vr_samp <- subset(vr_merged, samples = vr_name)
  vr_meta <- Metadata(vr_samp)
  condition <- paste(unique(vr_meta$Condition), collapse = ", ")
  
  # Create a temporary dataframe for this FOV
  temp_df <- data.frame(
    Interaction = cur_cell_proximities$unified_int,
    Enrichment = cur_cell_proximities$enrichm,
    FOV = vr_name,
    Condition = condition
  )
  
  # Append to master dataframe
  all_interactions <- rbind(all_interactions, temp_df)
}

# --- 2. Reshape data into a Wide Matrix ---
wide_data <- all_interactions %>%
  filter(grepl("NK cells/ILC1s", Interaction)) %>%
  select(Interaction, FOV, Enrichment) %>%
  mutate(Interaction = gsub("NK cells/ILC1s--|--NK cells/ILC1s", "", Interaction)) %>%
  pivot_wider(names_from = FOV, values_from = Enrichment, values_fill = 0, values_fn = mean) 

# Convert to matrix and assign rownames
enrichmat_global <- as.matrix(wide_data[, -1])
rownames(enrichmat_global) <- wide_data$Interaction

# --- 3. Create Column Annotations (Condition mapping) ---
anno_df <- unique(all_interactions[, c("FOV", "Condition")])
rownames(anno_df) <- anno_df$FOV

# Reorder the rows
anno_df <- anno_df[colnames(enrichmat_global), , drop = FALSE]

condition_colors <- cols_treat

ha_top <- HeatmapAnnotation(
  Condition = anno_df$Condition,
  col = list(Condition = condition_colors),
  annotation_name_side = "left"
)

# --- 4. Draw the Unified Master Heatmap ---
master_heatmap <- Heatmap(
  enrichmat_global,
  
  col = circlize::colorRamp2(c(-1, 0, 1), cols_heat),
  name = "Enrichment\nScore",
  heatmap_legend_param = list(direction = "vertical"),
  
  column_split = anno_df$Condition,
  cluster_column_slices = FALSE, 
  cluster_columns = FALSE,        
  cluster_rows = TRUE,           
  
  # Y-Axis Label
  row_title = "Interactions",
  row_title_gp = gpar(fontsize = 11#, fontface = "bold"
                      ),

  # Aesthetics
  show_column_names = FALSE,
  column_names_rot = 45,
  column_names_gp = gpar(fontsize = 8),
  row_names_gp = gpar(fontsize = 9),
  show_row_dend = TRUE,
  
  # Annotations and Titles
  top_annotation = ha_top,
  column_title = "Neighborhood Enrichment NK cells/ILC1s",
  column_title_gp = gpar(fontsize = 12, fontface = "bold")
)

# --- 5. Convert to ggplot2 and explicitly position the X-axis ---

# First, capture the raw heatmap (with normal, minimal padding)
heat_grob <- grid::grid.grabExpr(
  draw(master_heatmap, 
       heatmap_legend_side = "right",     
       annotation_legend_side = "right",  
       merge_legend = TRUE,
       padding = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
)

# Convert to a standard ggplot object
gg_heat <- as.ggplot(heat_grob)

# Add the X-axis label and manually nudge it to center under the heatmap body
final_plot_nkilc1 <- gg_heat + 
  labs(x = "FOV") + 
  theme(
    axis.title.x = element_text(
      size = 11, 
      # face = "bold", 
      hjust = 0.37 
    ), 
    plot.margin = margin(0, 0, 0.5, 0.2, "cm")
  )

# Print your beautiful, perfectly aligned plot!
print(final_plot_nkilc1)
```

## Coenrichment barplots

### ILC2s

```{r, fig.height=3, fig.width=9}

# fine tune the co-enrichment plot
# ILC2s around lymphatics -----------------------------------------
interactions <- c("ILC2s--LYVE1 CD90 Lymphatics")
unic <- "LYVE1 CD90 Lymphatics"
background_image <- backgroundlist[["LYVE1 CD90 Lymphatics"]]
interaction_celltypes <- NULL
for(samp in vr_list_names){
    cur_cell_proximities <- cell_proximities_list[[samp]]
    cur_cell_proximities <- cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,]
    sample <- unique(metadatax$FullInfo[metadatax$Sample==samp])
    if(nrow(cur_cell_proximities) > 0 & sample != "20210906_FOV3_D3"){
      interaction_celltypes <- rbind(interaction_celltypes,
                                     data.frame(cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,], 
                                                experiment = strsplit(sample, split = "_")[[1]][1], fov = strsplit(sample, split = "_")[[1]][2], condition = strsplit(sample, split = "_")[[1]][3]))
    }
  }
interaction_celltypes$p.adj <- ifelse(interaction_celltypes$enrichm > 0, interaction_celltypes$p.adj_higher, interaction_celltypes$p.adj_lower)
  # plot test results
  # sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("p=",round(interaction_celltypes$p.adj,3)), ""))
  sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("*"), ""))
    # print(sig_label)
plot_coenrichment_lymp <- ggplot(interaction_celltypes, aes(x = condition, y = enrichm, fill = condition)) +
  geom_bar(stat = "identity", position = position_dodge2(width=0.9, preserve = "single")) +
  facet_grid(.~condition, scales = "free_x") +
  geom_text(aes(label=sig_label), position=position_dodge2(width=0.9, preserve = "single"), angle = 90, hjust = -0.02, size = 4) +
  ylim(-1.6,3.6)+
  NoLegend()+
  theme_classic2()+
  scale_fill_manual(values = cols_fov, name = "") +
  ggtitle(gsub("LYVE1 CD90 ", "", interactions)) +
  theme(axis.text.x = element_text(#angle = 50,
                                   vjust = 1, size = 9, hjust = 0.5
                                   ),
        axis.text.y = element_text(hjust = 0.5, size = 9),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size =11, hjust = 0.5, face = "bold"),
        plot.margin = margin(0.1, 0.1, 0.1, 0.1, "cm"),
        legend.position = "none",
        strip.background=element_blank(),
        strip.background.x= element_blank(),
        strip.text.x = element_text(size = 1, color = "white"),
        panel.grid.major.y = element_line())+
  NoLegend()+  
  ylab("Enrichment")


# fine tune the co-enrichment plot
# ILC2s around ILC2s -----------------------------------------
interactions <- c("ILC2s--ILC2s")
unic <- "ILC2s"
background_image <- backgroundlist[["ILC2s"]]
interaction_celltypes <- NULL
for(samp in vr_list_names){
    cur_cell_proximities <- cell_proximities_list[[samp]]
    cur_cell_proximities <- cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,]
    sample <- unique(metadatax$FullInfo[metadatax$Sample==samp])
    if(nrow(cur_cell_proximities) > 0 & sample != "20210906_FOV3_D3"){
      interaction_celltypes <- rbind(interaction_celltypes,
                                     data.frame(cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,], 
                                                experiment = strsplit(sample, split = "_")[[1]][1], fov = strsplit(sample, split = "_")[[1]][2], condition = strsplit(sample, split = "_")[[1]][3]))
    }
  }
interaction_celltypes$p.adj <- ifelse(interaction_celltypes$enrichm > 0, interaction_celltypes$p.adj_higher, interaction_celltypes$p.adj_lower)
  # plot test results
  # sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("p=",round(interaction_celltypes$p.adj,3)), ""))
  sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("*"), ""))
    # print(sig_label)
plot_coenrichment_ilc2 <- ggplot(interaction_celltypes, aes(x = condition, y = enrichm, fill = condition)) +
  geom_bar(stat = "identity", position = position_dodge2(width=0.9, preserve = "single")) +
  facet_grid(.~condition, scales = "free_x") +
  geom_text(aes(label=sig_label), position=position_dodge2(width=0.9, preserve = "single"), angle = 90, hjust = -0.02, size = 4) +
  ylim(-1.6,3.6)+
  NoLegend()+
  theme_classic2()+
  scale_fill_manual(values = cols_fov, name = "") +
  ggtitle(gsub("LYVE1 CD90 ", "", interactions)) +
  theme(axis.text.x = element_text(#angle = 50,
                                   vjust = 1, size = 9, hjust = 0.5
                                   ),
        axis.text.y = element_text(hjust = 0.5, size = 9),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size =11, hjust = 0.5, face = "bold"),
        plot.margin = margin(0.1, 0.1, 0.1, 0.1, "cm"),
        legend.position = "none",
        strip.background=element_blank(),
        strip.background.x= element_blank(),
        strip.text.x = element_text(size = 1, color = "white"),
        panel.grid.major.y = element_line())+
  NoLegend()+  
  ylab("Enrichment")


# fine tune the co-enrichment plot
# ILC2s around Myeloid cells -----------------------------------------
interactions <- c("ILC2s--Myeloid cells")
unic <- "Myeloid cells"
background_image <- backgroundlist[["Myeloid cells"]]
interaction_celltypes <- NULL
for(samp in vr_list_names){
    cur_cell_proximities <- cell_proximities_list[[samp]]
    cur_cell_proximities <- cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,]
    sample <- unique(metadatax$FullInfo[metadatax$Sample==samp])
    if(nrow(cur_cell_proximities) > 0 & sample != "20210906_FOV3_D3"){
      interaction_celltypes <- rbind(interaction_celltypes,
                                     data.frame(cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,], 
                                                experiment = strsplit(sample, split = "_")[[1]][1], fov = strsplit(sample, split = "_")[[1]][2], condition = strsplit(sample, split = "_")[[1]][3]))
    }
  }
interaction_celltypes$p.adj <- ifelse(interaction_celltypes$enrichm > 0, interaction_celltypes$p.adj_higher, interaction_celltypes$p.adj_lower)
  # plot test results
  # sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("p=",round(interaction_celltypes$p.adj,3)), ""))
  sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("*"), ""))
    # print(sig_label)
plot_coenrichment_my <- ggplot(interaction_celltypes, aes(x = condition, y = enrichm, fill = condition)) +
  geom_bar(stat = "identity", position = position_dodge2(width=0.9, preserve = "single")) +
  facet_grid(.~condition, scales = "free_x") +
  geom_text(aes(label=sig_label), position=position_dodge2(width=0.9, preserve = "single"), angle = 90, hjust = -0.02, size = 4) +
  ylim(-1.6,3.6)+
  NoLegend()+
  theme_classic2()+
  scale_fill_manual(values = cols_fov, name = "") +
  ggtitle(gsub("LYVE1 CD90 ", "", interactions)) +
  theme(axis.text.x = element_text(#angle = 50,
                                   vjust = 1, size = 9, hjust = 0.5
                                   ),
        axis.text.y = element_text(hjust = 0.5, size = 9),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size =11, hjust = 0.5, face = "bold"),
        plot.margin = margin(0.1, 0.1, 0.1, 0.1, "cm"),
        legend.position = "none",
        strip.background=element_blank(),
        strip.background.x= element_blank(),
        strip.text.x = element_text(size = 1, color = "white"),
        panel.grid.major.y = element_line())+
  NoLegend()+  
  ylab("Enrichment")

coenrichment_fig <- ggarrange(plot_coenrichment_ilc2, plot_coenrichment_my, plot_coenrichment_lymp, ncol = 3, nrow = 1)

coenrichment_fig_ann <- annotate_figure(coenrichment_fig, 
                left = text_grob("Coenrichment score", color = "black", rot = 90, 
                                 size = 11))

coenrichment_fig_ann
```

### NK cells/ILC1s

```{r, fig.height=3, fig.width=9}

# fine tune the co-enrichment plot
# ILC2s around lymphatics -----------------------------------------
interactions <- c("EMCN CD31 Blood vessels--NK cells/ILC1s")
unic <- "EMCN CD31 Blood vessels"
background_image <- backgroundlist[["EMCN CD31 Blood vessels"]]
interaction_celltypes <- NULL
for(samp in vr_list_names){
    cur_cell_proximities <- cell_proximities_list[[samp]]
    cur_cell_proximities <- cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,]
    sample <- unique(metadatax$FullInfo[metadatax$Sample==samp])
    if(nrow(cur_cell_proximities) > 0 & sample != "20210906_FOV3_D3"){
      interaction_celltypes <- rbind(interaction_celltypes,
                                     data.frame(cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,], 
                                                experiment = strsplit(sample, split = "_")[[1]][1], fov = strsplit(sample, split = "_")[[1]][2], condition = strsplit(sample, split = "_")[[1]][3]))
    }
  }
interaction_celltypes$p.adj <- ifelse(interaction_celltypes$enrichm > 0, interaction_celltypes$p.adj_higher, interaction_celltypes$p.adj_lower)
  # plot test results
  # sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("p=",round(interaction_celltypes$p.adj,3)), ""))
  sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("*"), ""))
    # print(sig_label)
plot_coenrichment_bec <- ggplot(interaction_celltypes, aes(x = condition, y = enrichm, fill = condition)) +
  geom_bar(stat = "identity", position = position_dodge2(width=0.9, preserve = "single")) +
  facet_grid(.~condition, scales = "free_x") +
  geom_text(aes(label=sig_label), position=position_dodge2(width=0.9, preserve = "single"), angle = 90, hjust = -0.02, size = 4) +
  ylim(-1.6,3.6)+
  NoLegend()+
  theme_classic2()+
  scale_fill_manual(values = cols_fov, name = "") +
  ggtitle(gsub("EMCN CD31 ", "", interactions)) +
  theme(axis.text.x = element_text(#angle = 50,
                                   vjust = 1, size = 9, hjust = 0.5
                                   ),
        axis.text.y = element_text(hjust = 0.5, size = 9),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size =11, hjust = 0.5, face = "bold"),
        plot.margin = margin(0.1, 0.1, 0.1, 0.1, "cm"),
        legend.position = "none",
        strip.background=element_blank(),
        strip.background.x= element_blank(),
        strip.text.x = element_text(size = 1, color = "white"),
        panel.grid.major.y = element_line())+
  NoLegend()+  
  ylab("Enrichment")


# fine tune the co-enrichment plot
# ILC2s around ILC2s -----------------------------------------
interactions <- c("NK cells/ILC1s--NK cells/ILC1s")
unic <- "NK cells/ILC1s"
background_image <- backgroundlist[["NK cells/ILC1s"]]
interaction_celltypes <- NULL
for(samp in vr_list_names){
    cur_cell_proximities <- cell_proximities_list[[samp]]
    cur_cell_proximities <- cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,]
    sample <- unique(metadatax$FullInfo[metadatax$Sample==samp])
    if(nrow(cur_cell_proximities) > 0 & sample != "20210906_FOV3_D3"){
      interaction_celltypes <- rbind(interaction_celltypes,
                                     data.frame(cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,], 
                                                experiment = strsplit(sample, split = "_")[[1]][1], fov = strsplit(sample, split = "_")[[1]][2], condition = strsplit(sample, split = "_")[[1]][3]))
    }
  }
interaction_celltypes$p.adj <- ifelse(interaction_celltypes$enrichm > 0, interaction_celltypes$p.adj_higher, interaction_celltypes$p.adj_lower)
  # plot test results
  # sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("p=",round(interaction_celltypes$p.adj,3)), ""))
  sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("*"), ""))
    # print(sig_label)
plot_coenrichment_nkilc1 <- ggplot(interaction_celltypes, aes(x = condition, y = enrichm, fill = condition)) +
  geom_bar(stat = "identity", position = position_dodge2(width=0.9, preserve = "single")) +
  facet_grid(.~condition, scales = "free_x") +
  geom_text(aes(label=sig_label), position=position_dodge2(width=0.9, preserve = "single"), angle = 90, hjust = -0.02, size = 4) +
  ylim(-1.6,3.6)+
  NoLegend()+
  theme_classic2()+
  scale_fill_manual(values = cols_fov, name = "") +
  ggtitle(gsub("LYVE1 CD90 ", "", interactions)) +
  theme(axis.text.x = element_text(#angle = 50,
                                   vjust = 1, size = 9, hjust = 0.5
                                   ),
        axis.text.y = element_text(hjust = 0.5, size = 9),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size =11, hjust = 0.5, face = "bold"),
        plot.margin = margin(0.1, 0.1, 0.1, 0.1, "cm"),
        legend.position = "none",
        strip.background=element_blank(),
        strip.background.x= element_blank(),
        strip.text.x = element_text(size = 1, color = "white"),
        panel.grid.major.y = element_line())+
  NoLegend()+  
  ylab("Enrichment")


# fine tune the co-enrichment plot
# ILC2s around Myeloid cells -----------------------------------------
interactions <- c("B cells & Plasma cells--NK cells/ILC1s")
unic <- "B cells & Plasma cells"
background_image <- backgroundlist[["B cells & Plasma cells"]]
interaction_celltypes <- NULL
for(samp in vr_list_names){
    cur_cell_proximities <- cell_proximities_list[[samp]]
    cur_cell_proximities <- cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,]
    sample <- unique(metadatax$FullInfo[metadatax$Sample==samp])
    if(nrow(cur_cell_proximities) > 0 & sample != "20210906_FOV3_D3"){
      interaction_celltypes <- rbind(interaction_celltypes,
                                     data.frame(cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,], 
                                                experiment = strsplit(sample, split = "_")[[1]][1], fov = strsplit(sample, split = "_")[[1]][2], condition = strsplit(sample, split = "_")[[1]][3]))
    }
  }
interaction_celltypes$p.adj <- ifelse(interaction_celltypes$enrichm > 0, interaction_celltypes$p.adj_higher, interaction_celltypes$p.adj_lower)
  # plot test results
  # sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("p=",round(interaction_celltypes$p.adj,3)), ""))
  sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("*"), ""))
    # print(sig_label)
plot_coenrichment_bpc <- ggplot(interaction_celltypes, aes(x = condition, y = enrichm, fill = condition)) +
  geom_bar(stat = "identity", position = position_dodge2(width=0.9, preserve = "single")) +
  facet_grid(.~condition, scales = "free_x") +
  geom_text(aes(label=sig_label), position=position_dodge2(width=0.9, preserve = "single"), angle = 90, hjust = -0.02, size = 4) +
  ylim(-1.6,3.6)+
  NoLegend()+
  theme_classic2()+
  scale_fill_manual(values = cols_fov, name = "") +
  ggtitle(gsub("LYVE1 CD90 ", "", interactions)) +
  theme(axis.text.x = element_text(#angle = 50,
                                   vjust = 1, size = 9, hjust = 0.5
                                   ),
        axis.text.y = element_text(hjust = 0.5, size = 9),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size =11, hjust = 0.5, face = "bold"),
        plot.margin = margin(0.1, 0.1, 0.1, 0.1, "cm"),
        legend.position = "none",
        strip.background=element_blank(),
        strip.background.x= element_blank(),
        strip.text.x = element_text(size = 1, color = "white"),
        panel.grid.major.y = element_line())+
  NoLegend()+  
  ylab("Enrichment")

coenrichment_fig_supp <- ggarrange(plot_coenrichment_nkilc1, plot_coenrichment_bpc, plot_coenrichment_bec, ncol = 3, nrow = 1)

coenrichment_fig_supp_ann <- annotate_figure(coenrichment_fig_supp, 
                left = text_grob("Coenrichment score", color = "black", rot = 90, 
                                 size = 11))

coenrichment_fig_supp_ann
```

# CIN analysis

```{r, fig.height=3, fig.width=9}

tissuearea <- "Lung"
input_dir <- here::here("data", "CIN_analysis", tissuearea, "/")


# List all radii calculated 
list_files_rad <- list.files(path=input_dir, 
                         pattern=NULL, all.files=FALSE,
                         full.names=FALSE)

list_files_rad
df_cin_lung = data.frame()

# define a dataframe to collect all data from one tissue area
df_radius <- data.frame()
for (defined_radius in list_files_rad) {
  print(defined_radius)
  list_files <- list.files(path=paste0(input_dir,
                                   defined_radius, 
                                   "/"), 
                       pattern=NULL, all.files=FALSE,
                       full.names=FALSE)
  # define a dataframe to collect the data for all datasets of one 
  df_cin = data.frame()
  for (element in list_files) {
    dir <- paste0(input_dir,
                  defined_radius, 
                  "/", element)
    df_sub <- read_csv(dir, 
      col_types = cols(...1 = col_skip()))
    df_sub$Filename <- paste0(defined_radius, 
                  "/", element)
    df_cin <- rbind(df_cin, df_sub)
  }

  df_cin$Radius <- defined_radius
  df_radius <- rbind(df_radius, df_cin)
}

# collect data from current tissue area to datafram df_all_cin
df_cin_lung <- df_radius
df_cin_lung$Radius <- gsub("_micm", " \u03BCm", df_cin_lung$Radius)
df_cin_lung$Experiment <- str_sub(df_cin_lung$Dataset,-8,-1)
df_cin_lung$FOV <- str_sub(df_cin_lung$Dataset,-10,-10)  
df_cin_lung$Treatment <- str_extract(df_cin_lung$Dataset, "[^_]+")
df_cin_lung$ExpID <- paste0(df_cin_lung$Experiment, "_", df_cin_lung$FOV)
colnames(df_cin_lung) <- gsub("Reference cell", "Reference", 
                                colnames(df_cin_lung))

colnames(df_cin_lung) <- gsub("NK cells & ILC1s", "NK cells/ILC1s", 
                              colnames(df_cin_lung))
df_cin_lung$Reference <- gsub("NK cells & ILC1s", "NK cells/ILC1s",
                              df_cin_lung$Reference)
head(df_cin_lung)
colnames(df_cin_lung)

```

```{r fig.height=6, fig.width=9, warning=FALSE}

# 1. Target cells, References, and Radii
target_cells <- c("Epithelia", "EMCN CD31 Blood vessels", "LYVE1 CD31 vessels", 
                  "LYVE1 CD90 Lymphatics", "Myeloid cells", "B cells & Plasma cells", 
                  "NK cells/ILC1s", "ILC3s", "T cytotox cells", "T helper cells", "ILC2s")

all_refs <- unique(df_cin_lung$Reference)
all_radii <- unique(df_cin_lung$Radius)

# Initialize the global master list 
master_plot_list <- list()

# 2. Loop through Radii and References
for (set_radius in all_radii) {
  for (ref_type in all_refs) {
    
    base_df <- df_cin_lung %>%
      filter(Reference == ref_type) %>%
      filter(Radius == set_radius) %>%
      mutate(Treatment = factor(Treatment, levels = c("CTRL", "D1", "D2", "D3")))
    
    if (nrow(base_df) == 0) next
    
    # CRITICAL: This temporary list MUST be outside the target_cells loop
    freq_plot_list <- list()
    
    # 3. Loop through the 11 target cell types
    for (tgt_cell in target_cells) {
      
      stats_df <- base_df %>%
        select(Treatment, FOV, Frequency = all_of(tgt_cell))
      
      global_max <- max(stats_df$Frequency, na.rm = TRUE)
      
      n_df <- stats_df %>%
        group_by(Treatment) %>%
        summarise(n = n(), .groups = "drop") %>%
        mutate(
          Label = paste0("n=", n),
          Y_pos = ifelse(is.finite(global_max), global_max * 1.05, 1) 
        )
      
      # --- STATISTICAL TESTING ---
      stat.test <- tryCatch({
        stats_df %>%
          wilcox_test(Frequency ~ Treatment, ref.group = "CTRL") %>%
          add_significance() %>% 
          add_xy_position(x = "Treatment")
      }, error = function(e) NULL)
      
      plot_title <- tgt_cell
      
      # --- BUILD THE PLOT ---
      p <- ggplot(stats_df, aes(x = Treatment, y = Frequency)) +
        geom_boxplot(outlier.colour = NA, fill = "white", alpha = 0.5) +
        geom_beeswarm(aes(color = Treatment), size = 1.5, cex = 2) + 
        # geom_text(data = n_df, aes(x = Treatment, y = Y_pos, label = Label), 
        #           size = 3.2, fontface = "italic", inherit.aes = FALSE) +
        ggtitle(gsub("LYVE1 CD90 |EMCN CD31 ", "", plot_title)) +
        scale_y_continuous(expand = expansion(mult = c(0, 0.45))) +
        theme_classic() + 
        theme(axis.text.x = element_text(size = 9#, face = "bold"
                                         ),
              axis.text.y = element_text(size = 9),
              axis.title.x = element_blank(), 
              axis.title.y = element_text(size = 9), 
              plot.title = element_text(size = 11, hjust = 0.5),
              legend.position = "none") +
        ylab(paste0("Frequency\nwithinr=", set_radius))+
        scale_color_manual(values = cols_treat) 
      
      # --- THE BULLETPROOF STATS FIX ---
      if (!is.null(stat.test) && nrow(stat.test) > 0) {
        # Dynamically detect if rstatix used multiple testing (p.adj.signif) 
        # or a single comparison (p.signif) for this specific cell type!
        sig_label <- ifelse("p.adj.signif" %in% colnames(stat.test), "p.adj.signif", "p.signif")
        
        p <- p + stat_pvalue_manual(stat.test, 
                                    label = sig_label, # Uses the dynamically detected column
                                    hide.ns = TRUE, 
                                    y.position = global_max * 1.1, 
                                    step.increase = 0.15, 
                                    size = 4)
      }
      
      # Save to the temporary grid list
      freq_plot_list[[plot_title]] <- p
      
      # Save to the master list
      plot_key <- paste(set_radius, ref_type, tgt_cell, sep = "_")
      master_plot_list[[plot_key]] <- p
    }
    
    # --- PRINT THE PATCHWORK GRID ---
    if (length(freq_plot_list) > 0) {
      final_cin_grid <- wrap_plots(freq_plot_list, ncol = 4) +
        plot_annotation(
          title = paste0("Spatial Neighborhood (", set_radius, ") - Reference: ", ref_type),
          subtitle = 'Points represent individual FOVs; Stats: Wilcoxon vs CTRL',
          theme = theme(
            plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
            plot.subtitle = element_text(size = 11, hjust = 0.5)
          )
        )
      
      message("Printing grid for: ", ref_type, " at ", set_radius)
      print(final_cin_grid)
    }
  }
}
```

### ILC2s

```{r fig.height=6, fig.width=6, warning=FALSE}
cin_ilc <- 
  ggarrange(
  master_plot_list[["10 μm_ILC2s_ILC2s"]], 
  master_plot_list[["10 μm_ILC2s_Myeloid cells"]], 
  master_plot_list[["10 μm_ILC2s_LYVE1 CD90 Lymphatics"]], 
  ncol = 3, nrow = 1)+
  plot_annotation(
          title = paste0("ILC2s (CIN = 10 µm)"),
          theme = theme(
            plot.title = element_text(size = 11, face = "bold", hjust = 0.5), 
            plot.margin = margin(0, 0, 0, 0, "cm")
          )
        )


cin_my <- ggarrange(
  master_plot_list[["10 μm_Myeloid cells_ILC2s"]], 
  master_plot_list[["10 μm_Myeloid cells_Myeloid cells"]], 
  master_plot_list[["10 μm_Myeloid cells_LYVE1 CD90 Lymphatics"]], 
  ncol = 3, nrow = 1)+
  plot_annotation(
          title = paste0("Myeloid cells (CIN = 10 µm)"),
          theme = theme(
            plot.title = element_text(size = 11, face = "bold", hjust = 0.5), 
            plot.margin = margin(0, 0, 0, 0, "cm")
          )
        )

cin_ly <- ggarrange(
  master_plot_list[["10 μm_LYVE1 CD90 Lymphatics_ILC2s"]], 
  master_plot_list[["10 μm_LYVE1 CD90 Lymphatics_Myeloid cells"]], 
  master_plot_list[["10 μm_LYVE1 CD90 Lymphatics_LYVE1 CD90 Lymphatics"]], 
  ncol = 3, nrow = 1)+
  plot_annotation(
          title = paste0("Lymphatics (CIN = 10 µm)"),
          theme = theme(
            plot.title = element_text(size = 11, face = "bold", hjust = 0.5), 
            plot.margin = margin(0, 0, 0, 0, "cm")
          )
        )

plot_cin <- ggarrange(cin_ilc, cin_my, cin_ly, 
                      ncol = 1, nrow = 3, labels = c("E", "", ""))+
  theme(plot.margin = margin(0, 0, 0, 0, "cm"))

plot_cin
```

### NK cells/ILC1s

```{r fig.height=6, fig.width=6, warning=FALSE}
cin_nkilc1 <- 
  ggarrange(
  master_plot_list[["10 μm_NK cells/ILC1s_NK cells/ILC1s"]], 
  master_plot_list[["10 μm_NK cells/ILC1s_EMCN CD31 Blood vessels"]], 
  master_plot_list[["10 μm_NK cells/ILC1s_B cells & Plasma cells"]], 
  ncol = 3, nrow = 1)+
  plot_annotation(
          title = paste0("NK cells/ILC1s (CIN = 10 µm)"),
          theme = theme(
            plot.title = element_text(size = 11, face = "bold", hjust = 0.5), 
            plot.margin = margin(0, 0, 0, 0, "cm")
          )
        )


cin_bec <- ggarrange(
  master_plot_list[["10 μm_EMCN CD31 Blood vessels_NK cells/ILC1s"]], 
  master_plot_list[["10 μm_EMCN CD31 Blood vessels_EMCN CD31 Blood vessels"]], 
  master_plot_list[["10 μm_EMCN CD31 Blood vessels_B cells & Plasma cells"]], 
  ncol = 3, nrow = 1)+
  plot_annotation(
          title = paste0("Blood vessels (CIN = 10 µm)"),
          theme = theme(
            plot.title = element_text(size = 11, face = "bold", hjust = 0.5), 
            plot.margin = margin(0, 0, 0, 0, "cm")
          )
        )

cin_bpc <- ggarrange(
  master_plot_list[["10 μm_B cells & Plasma cells_NK cells/ILC1s"]], 
  master_plot_list[["10 μm_B cells & Plasma cells_EMCN CD31 Blood vessels"]], 
  master_plot_list[["10 μm_B cells & Plasma cells_B cells & Plasma cells"]], 
  ncol = 3, nrow = 1)+
  plot_annotation(
          title = paste0("B cells & Plasma cells (CIN = 10 µm)"),
          theme = theme(
            plot.title = element_text(size = 11, face = "bold", hjust = 0.5), 
            plot.margin = margin(0, 0, 0, 0, "cm")
          )
        )

plot_cin_nkilc1 <- ggarrange(cin_nkilc1, cin_bec, cin_bpc, 
                      ncol = 1, nrow = 3, labels = c("D", "", ""))+
  theme(plot.margin = margin(0, 0, 0, 0, "cm"))

plot_cin_nkilc1
```

# Minimum distance

## Load data

```{r, fig.width=4, fig.height=6}
set.seed(8)
radius <- 10


df_dist_all <- read_csv(here::here("data", "SPIAT_min_dist_all_lung.csv"), 
    col_types = cols(...1 = col_skip()))

df_dist_all$RefType <- factor(df_dist_all$RefType, levels = celltypes)
df_dist_all$NearestType <- factor(df_dist_all$NearestType, levels = celltypes)

df_dist_all <- df_dist_all %>%
  mutate(Treatment = paste0("D", Treatment)) %>%
  mutate(Treatment = gsub("DCTRL", "CTRL", Treatment))

head(df_dist_all)

```

## Immune cells to lymphatics

### ILC2s

```{r, fig.width=6, fig.height=3}
# filter condition and reference cell types to compare ILC subtypes and 
# T cells
df_dist_ref <- df_dist_all %>%
  # filter(Treatment == condition) %>%
  filter(RefType == "ILC2s"
         |
           RefType == "NK cells/ILC1s"| RefType == "ILC3s"|
           RefType == "T helper cells"| RefType == "T cytotox cells"| RefType == "Myeloid cells"| RefType == "B cells & Plasma cells"
         )

unique(df_dist_ref$Treatment)

df_dist_ref$Treatment <- factor(df_dist_ref$Treatment, levels = c(
  "CTRL", "D1", "D2", "D3"
))
unique(df_dist_ref$RefType)


# LYMPHATICS ------------------------------------------------------------
celltype_of_interest <- "LYVE1 CD90 Lymphatics"
ypos <- 100


my_plot_list <- list()

celltype_of_interest <- "LYVE1 CD90 Lymphatics"

for (condition in c("CTRL", "D1", "D2", "D3")) {
  
  plot_data <- df_dist_ref %>%
    filter(Treatment == condition) %>%
    filter(NearestType == celltype_of_interest)
  
  # Test for statistical significance of ILC2s to the other cell types
  stat.test <- plot_data %>%
    dunn_test(Distance ~ RefType) %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance()
  stat.test <- stat.test %>% add_xy_position(x = "RefType")
  
  stat.test <- stat.test %>%
    filter(group1 == "ILC2s"|group2 == "ILC2s"
           )
  
  # Create lav´bels that depict mean value 
  Labs = plot_data %>% 
    group_by(RefType) %>%
    summarise(lab_text = paste0(round(median(Distance), 0), " µm"), lab_pos = quantile(Distance)[3]
                )
  
  # create plot
  plot <- ggplot(plot_data, aes(x=RefType , 
            y = Distance, 
            fill = "RefType")) +
    geom_boxplot(fill="white", outliers = FALSE)+
    geom_beeswarm(aes(color = RefType), 
                  alpha = 0.2, size = 0.1, cex = 0.2)+
    stat_pvalue_manual(stat.test,
                       size = 6,
                       hide.ns = TRUE, 
                       y.position = 250, 
                       step.increase = 0.1
                       )+
    scale_y_continuous(expand = c(0, 0), limits = c(0,400))+
    rotate_x_text(angle = 30)+
  ggtitle(condition)+
  scale_color_manual(values = ColorsCellType)+
  theme_classic2()+
  theme(
    axis.text.x = element_text(angle = 30, vjust = 1, hjust = 1, 
                                   # vjust = 1, hjust = 0.5, 
                                   size = 9),
    axis.title.x = element_blank(),
    axis.text.y = element_text(hjust = 0.5, size = 9),
    axis.title.y = element_text(size = 9),
          plot.title = element_text(size = 10, hjust = 0.5, 
                                    face = "bold"),
    plot.margin = margin(0, 0, 0.2, 0.2, "cm"),
    legend.title = element_text(size =11),
    legend.text = element_text(size =12))+
    NoLegend()+
    xlab("Cell types")+
    ylab("Minimal distance [µm]")+
    geom_text(aes(y = 150 , #lab_pos, 
                  label = lab_text, vjust = -0.5), angle = 30,
              data = Labs, size=2)
  plot
  
  assigned_name <- gsub(" ", "", paste0(celltype_of_interest, "_", condition))
  assign(assigned_name, plot )
  my_plot_list[[assigned_name]] <- plot
  

}

ggarrange(plotlist = my_plot_list[c(1:4)], nrow = 1, ncol = 4)

dist_lymph <- ggarrange(plotlist = my_plot_list[c(1, 4)], nrow = 1, ncol = 2, labels = c("C", "D"))

dist_lymph

```

```{r fig.height=6, fig.width=3, warning=FALSE}

dist_lymph <- ggarrange(plotlist = my_plot_list[c(1, 4)], nrow = 2, ncol = 1, labels = c("D"))+
  theme(plot.margin = margin(0, 1, 0, 0, "cm"))

dist_lymph

```

```{r fig.height=4, fig.width=9, warning=FALSE}

# --- 1. SETTINGS ---
celltype_of_interest <- "LYVE1 CD90 Lymphatics"
ref_cell_types <- unique(df_dist_ref$RefType)
treatment_levels <- c("CTRL", "D1", "D2", "D3")

master_dist_list <- list() 
grid_plot_list <- list()   

# --- 2. LOOP THROUGH REFERENCE CELL TYPES ---
for (ref_type in ref_cell_types) {
  
  # Step A: Filter data
  stats_df <- df_dist_ref %>%
    filter(RefType == ref_type) %>%
    filter(NearestType == celltype_of_interest) %>%
    mutate(Treatment = factor(Treatment, levels = treatment_levels))
  
  # Skip if there's no data for this specific pairing
  if (nrow(stats_df) < 5) next
  
  # Step B: Statistical Testing (vs CTRL only)
  stat.test <- tryCatch({
    stats_df %>%
      wilcox_test(Distance ~ Treatment, ref.group = "CTRL") %>%
      add_significance() %>%
      add_xy_position(x = "Treatment")
  }, error = function(e) NULL)
  
  # Step C: Calculate Median Distance for Labels
  median_df <- stats_df %>%
    group_by(Treatment) %>%
    summarise(med_dist = median(Distance, na.rm = TRUE), .groups = "drop") %>%
    mutate(
      Label = paste0(round(med_dist, 0), " \u00b5m"), 
      Y_pos = 250 # Median labels kept at 400
    )
  
  # Step D: Build the Plot
  plot_title <- paste0(ref_type) 
  
  p <- ggplot(stats_df, aes(x = Treatment, y = Distance)) +
    geom_boxplot(outlier.colour = NA, fill = "white", alpha = 0.5) +
    geom_beeswarm(aes(color = Treatment), size = 0.1, cex = 0.2, alpha = 0.2) + 
    
    # Median Labels at y=400
    geom_text(data = median_df, aes(x = Treatment, y = Y_pos,
                                    label = Label), 
              size = 3, angle = 30, color = "black", 
              inherit.aes = FALSE) +
    ggtitle(plot_title) +
    # Y-axis fixed from 0 to 1000
    scale_y_continuous(limits = c(0, 500), expand = c(0, 0)) + 
    theme_classic() + 
    theme(axis.text.x = element_text(size = 9),
          axis.text.y = element_text(size = 9),
          axis.title.x = element_blank(),
          axis.title.y = element_text(size = 9),
          plot.title = element_text(size = 10, hjust = 0.5, 
                                    face = "bold"),
          plot.margin = margin(0.25, 0.5, 0.25, 0.5, "cm"),
          legend.position = "none") +
    ylab("Distance [\u00b5m]") +
    scale_color_manual(values = cols_treat)

  # Step E: Add Stat Brackets
  if (!is.null(stat.test) && nrow(stat.test) > 0) {
    sig_col <- if ("p.adj.signif" %in% colnames(stat.test)) "p.adj.signif" else "p.signif"
    
    p <- p + stat_pvalue_manual(
      stat.test, 
      label = sig_col, 
      hide.ns = TRUE, 
      tip.length = 0.01,
      # --- FIXED: Significance brackets start at y = 300 ---
      y.position = 380, 
      step.increase = 0.1, 
      size = 3.5
    )
  }
  
  # Save to lists
  grid_plot_list[[ref_type]] <- p
  plot_key <- paste0(ref_type, "_to_Lymphatics")
  master_dist_list[[plot_key]] <- p
}

# --- 3. PRINT THE FINAL GRID ---
if (length(grid_plot_list) > 0) {
  final_grid_dist_immune <- wrap_plots(grid_plot_list, ncol = 4) +
    plot_annotation(
      title = paste0("Minimum Distances to: ", gsub("LYVE1 CD90 ", "", celltype_of_interest)),
      theme = theme(
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
    )
  
  print(final_grid_dist_immune)
}
```

### NK cells/ILC1s

```{r fig.height=6, fig.width=3, warning=FALSE}
# filter condition and reference cell types to compare ILC subtypes and 
# T cells
df_dist_ref <- df_dist_all %>%
  # filter(Treatment == condition) %>%
  filter(RefType == "ILC2s"
         |
           RefType == "NK cells/ILC1s"| RefType == "ILC3s"|
           RefType == "T helper cells"| RefType == "T cytotox cells"| RefType == "Myeloid cells"| RefType == "B cells & Plasma cells"
         )

unique(df_dist_ref$Treatment)

df_dist_ref$Treatment <- factor(df_dist_ref$Treatment, levels = c(
  "CTRL", "D1", "D2", "D3"
))
unique(df_dist_ref$RefType)


# Blood vessels ------------------------------------------------------------
celltype_of_interest <- "EMCN CD31 Blood vessels"
ypos <- 100


my_plot_list <- list()

for (condition in c("CTRL", "D1", "D2", "D3")) {
  
  plot_data <- df_dist_ref %>%
    filter(Treatment == condition) %>%
    filter(NearestType == celltype_of_interest)
  
  # Test for statistical significance of ILC2s to the other cell types
  stat.test <- plot_data %>%
    dunn_test(Distance ~ RefType) %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance()
  stat.test <- stat.test %>% add_xy_position(x = "RefType")
  
  stat.test <- stat.test %>%
    filter(group1 == "NK cells/ILC1s"|group2 == "NK cells/ILC1s"
           )
  
  # Create lav´bels that depict mean value 
  Labs = plot_data %>% 
    group_by(RefType) %>%
    summarise(lab_text = paste0(round(median(Distance), 0), " µm"), lab_pos = quantile(Distance)[3]
                )
  
  # create plot
  plot <- ggplot(plot_data, aes(x=RefType , 
            y = Distance, 
            fill = "RefType")) +
    geom_boxplot(fill="white", outliers = FALSE)+
    geom_beeswarm(aes(color = RefType), 
                  alpha = 0.2, size = 0.1, cex = 0.08)+
    stat_pvalue_manual(stat.test,
                       size = 6,
                       hide.ns = TRUE, 
                       y.position = 110, 
                       step.increase = 0.15
                       )+
    scale_y_continuous(expand = c(0, 0), limits = c(0,200))+
    rotate_x_text(angle = 30)+
  ggtitle(condition)+
  scale_color_manual(values = ColorsCellType)+
  theme_classic2()+
  theme(
    axis.text.x = element_text(angle = 30, vjust = 1, hjust = 1, 
                                   # vjust = 1, hjust = 0.5, 
                                   size = 9),
    axis.title.x = element_blank(),
    axis.text.y = element_text(hjust = 0.5, size = 9),
    axis.title.y = element_text(size = 9),
          plot.title = element_text(size = 10, hjust = 0.5, 
                                    face = "bold"),
    plot.margin = margin(0, 0, 0.2, 0.2, "cm"),
    legend.title = element_text(size =11),
    legend.text = element_text(size =12))+
    NoLegend()+
    xlab("Cell types")+
    ylab("Minimal distance [µm]")+
    geom_text(aes(y = 80 , #lab_pos, 
                  label = lab_text, vjust = -0.5), angle = 30,
              data = Labs, size=2)
  plot
  
  assigned_name <- gsub(" ", "", paste0(celltype_of_interest, "_", condition))
  assign(assigned_name, plot )
  my_plot_list[[assigned_name]] <- plot
  

}

# ggarrange(plotlist = my_plot_list[c(1:4)], nrow = 1, ncol = 4)

dist_bec <- ggarrange(plotlist = my_plot_list[c(1, 4)], nrow = 2, ncol = 1, labels = c("C"))+
  theme(plot.margin = margin(0, 1, 0, 0, "cm"))

dist_bec

```

```{r fig.height=4, fig.width=9, warning=FALSE}

# --- 1. SETTINGS ---
celltype_of_interest <- "EMCN CD31 Blood vessels"
ref_cell_types <- unique(df_dist_ref$RefType)
treatment_levels <- c("CTRL", "D1", "D2", "D3")

master_dist_list <- list() 
grid_plot_list <- list()   

# --- 2. LOOP THROUGH REFERENCE CELL TYPES ---
for (ref_type in ref_cell_types) {
  
  # Step A: Filter data
  stats_df <- df_dist_ref %>%
    filter(RefType == ref_type) %>%
    filter(NearestType == celltype_of_interest) %>%
    mutate(Treatment = factor(Treatment, levels = treatment_levels))
  
  # Skip if there's no data for this specific pairing
  if (nrow(stats_df) < 5) next
  
  # Step B: Statistical Testing (vs CTRL only)
  stat.test <- tryCatch({
    stats_df %>%
      wilcox_test(Distance ~ Treatment, ref.group = "CTRL") %>%
      add_significance() %>%
      add_xy_position(x = "Treatment")
  }, error = function(e) NULL)
  
  # Step C: Calculate Median Distance for Labels
  median_df <- stats_df %>%
    group_by(Treatment) %>%
    summarise(med_dist = median(Distance, na.rm = TRUE), .groups = "drop") %>%
    mutate(
      Label = paste0(round(med_dist, 0), " \u00b5m"), 
      Y_pos = 100 # Median labels kept at 400
    )
  
  # Step D: Build the Plot
  plot_title <- paste0(ref_type) 
  
  p <- ggplot(stats_df, aes(x = Treatment, y = Distance)) +
    geom_boxplot(outlier.colour = NA, fill = "white", alpha = 0.5) +
    geom_beeswarm(aes(color = Treatment), size = 0.1, cex = 0.1, alpha = 0.2) + 
    
    # Median Labels at y=400
    geom_text(data = median_df, aes(x = Treatment, y = Y_pos,
                                    label = Label), 
              size = 3, angle = 30, color = "black", 
              inherit.aes = FALSE) +
    ggtitle(plot_title) +
    # Y-axis fixed from 0 to 1000
    scale_y_continuous(limits = c(0, 220), expand = c(0, 0)) + 
    theme_classic() + 
    theme(axis.text.x = element_text(size = 9),
          axis.text.y = element_text(size = 9),
          axis.title.x = element_blank(),
          axis.title.y = element_text(size = 9),
          plot.title = element_text(size = 10, hjust = 0.5, 
                                    face = "bold"),
          plot.margin = margin(0.25, 0.5, 0.25, 0.5, "cm"),
          legend.position = "none") +
    ylab("Distance [\u00b5m]") +
    scale_color_manual(values = cols_treat)

  # Step E: Add Stat Brackets
  if (!is.null(stat.test) && nrow(stat.test) > 0) {
    sig_col <- if ("p.adj.signif" %in% colnames(stat.test)) "p.adj.signif" else "p.signif"
    
    p <- p + stat_pvalue_manual(
      stat.test, 
      label = sig_col, 
      hide.ns = TRUE, 
      tip.length = 0.01,
      # --- FIXED: Significance brackets start at y = 300 ---
      y.position = 150, 
      step.increase = 0.1, 
      size = 3.5
    )
  }
  
  # Save to lists
  grid_plot_list[[ref_type]] <- p
  plot_key <- paste0(ref_type, "_to_Lymphatics")
  master_dist_list[[plot_key]] <- p
}

# --- 3. PRINT THE FINAL GRID ---
if (length(grid_plot_list) > 0) {
  final_grid_dist_immune_nkilc1 <- wrap_plots(grid_plot_list, ncol = 4) +
    plot_annotation(
      title = paste0("Minimum Distances to: ", gsub("EMCN CD31 ", "", celltype_of_interest)),
      theme = theme(
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
    )
  
  print(final_grid_dist_immune_nkilc1)
}
```

# Visualization for figures

## Figure 7

```{r, fig.width=9, fig.height=10.2}
top_figure <- ggarrange(final_plot_ILC2s, img_plot_if_ilc2s, img_plot_if_ilc2s_roi, ncol = 1, nrow = 3, heights = c(3, 3.6, 3.6), labels = c("A", "B", "C"))+
  theme(plot.margin = margin(0, 0, 0.5, 0, "cm"))

top_figure
```

```{r, fig.width=9, fig.height=5.5}
bottom_figure <- ggarrange(dist_lymph, plot_cin, ncol = 2, widths = c(1, 2))

bottom_figure
```

```{r, fig.width=9, fig.height=16}
ggarrange(top_figure, bottom_figure, nrow = 2, ncol = 1, heights = c(10.2, 5.5))
```

## Supplementary figures

```{r, fig.width=9, fig.height=12}
ggarrange(coenrichment_fig_ann, "NONE", final_grid_dist_immune, "NONE",
          img_plot_if_ilc2s_my, 
          nrow = 5, ncol = 1, heights = c(3, 0.25, 4, 0.25, 4.5), 
          labels = c("A","", "B", "", "C"))
```

```{r, fig.width=9, fig.height=15}
ggarrange(final_plot_nkilc1, coenrichment_fig_supp_ann, "NONE",  final_grid_dist_immune_nkilc1, "NONE", plot_cin_nkilc1, nrow = 6, ncol = 1, heights = c(1.7, 1.2, 0.1, 2, 0.1, 3), labels = c("A", "B", "", "C"))
```

## Session Information

```{r}
save.image(paste0(output_dir, "/environment.RData"))
sessionInfo()
```
