---
title: "Figure 8: Quantification and spatial analysis of ILC subtypes in the SI"
author: "Sandy Kroh"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    number_sections: yes
    fig_caption: yes
    keep_md: yes
  pdf_document:
    toc: yes
params:
  Sample: Default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", 
                      fig.align='center', dpi = 600, 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Libraries

```{r}
# remove.packages("rlang")
# remove.packages("dplyr")
# install.packages("rlang")
# install.packages("dplyr")

library(SeuratObject)
library(dplyr)
library(rstatix)
library(rlang)

if (!requireNamespace("Giotto", quietly = TRUE))
  devtools::install_github("drieslab/Giotto@suite")
if (!requireNamespace("VoltRon", quietly = TRUE))
  devtools::install_github("Artur-man/VoltRon")
if (!requireNamespace("Seurat", quietly = TRUE))
  install.packages("Seurat")
library(Giotto)
library(Seurat)
library(VoltRon)
library(ggplot2)
library(ggpubr)
library(readr)
library(ggbeeswarm)
library(stringr)
```

## Parameters

```{r}
set.seed(123)

input_dir <- here::here("1_data_tidying", "Lung_SI_all_cells_all_ALs_files")

output_dir <- here::here("2_visualizations_for_figures", "Fig_8_quantification_spatial_analysis_SI_files")
dir.create(output_dir)




main_markers <- c(
  "EpCAM", "EMCN", "LYVE1", "PDPN", "PDGFRa", "CD8a", "CD4",
  "CD45", "CD3", "IRF4", "Kappa", "CD11c", "CD127", "GATA3eGFP", "RORgt"
)


immune_markers <- c(
 "CD3", "CD4", "CD8a", "Kappa", "IRF4", "CD11c",
  "CD127", "CD90", "EOMES", "GATA3eGFP", "RORgt", "Ki67",  "KLRG1", "NKp46", "CD117", "Areg", "CCR6", "CD44", "MHCII", "Sca1"
)

ilc_markers <- c(
  "CD3", "CD4", "CD8a",
  "CD127", "CD90", "EOMES", "GATA3eGFP", "RORgt", "KLRG1", "NKp46", "CD117", "CCR6", "MHCII", "Ki67", "Areg", "IRF4", "Sca1", "CD44"
)


cols_nat <- c("magenta", "cyan", "blue", "purple", "green", 
                       "red", "yellow", "olivedrab1", "slateblue1", 
                       "darkcyan", "gold","indianred1", "seagreen", "deeppink", 
                       "orange", "brown", "violet",
                       "deeppink4", "pink", 
                       "grey", "black", "lightgreen", 
                       "#FF0066",  
                       "lightblue", "#FFCC99", "#CC00FF", 
                       "blueviolet",  "goldenrod4", 
                       "navy", "olivedrab", "lightcyan", "seagreen2", "darkviolet", "lightpink", "slateblue4", "olivedrab2")

colfunc <- colorRampPalette(c("darkcyan", "green", "yellow", "magenta", "purple"))

cols_ilcs_lung <- c("darkcyan", "seagreen2", "deeppink4")
cols_ilcs_si <- c("slateblue", "seagreen2")

cols_treat <- c("darkcyan", "gold", "slateblue")

```

# Load data

## Proportions

```{r, fig.width=9, fig.height=6}

df_villi <- read_csv(paste0(input_dir, "/si_villi_proportions.csv"), 
    col_types = cols(...1 = col_skip()))
df_ilf <- read_csv(paste0(input_dir, "/si_ilf_proportions.csv"), 
    col_types = cols(...1 = col_skip()))
df_lung <- read_csv(paste0(input_dir, "/lung_proportions.csv"), 
    col_types = cols(...1 = col_skip()))

```

## Spatial data for VoltRon and Giotto

```{r, fig.width=9, fig.height=6}

# from import_Giotto.Rmd
gio_list <- readRDS(here::here("data", "Giotto_data_Villi.rds"))

# from import_VoltRon.Rmd
vr_list <- readRDS(here::here("data", "VoltRon_data_Villi.rds"))

# original data
metadatax <- read_csv(here::here("data", "SO_arcsinh_si_imputed_Villi.csv"))

metadatax <- metadatax %>% 
  filter(`Tissue area` == "Villi")

unique(metadatax$CellType)

vr_list_names <- unique(metadatax$Dataset)


cell_proximities_list <- list()
for(samp in vr_list_names){
  print(samp)
  cell_proximities_list[[samp]] <-cellProximityEnrichment(
    gobject = gio_list[[samp]],
    cluster_column = 'CellType',
    spatial_network_name = 'Delaunay_network',
    adjust_method = 'fdr',
    number_of_simulations = 1000)
  cell_proximities_list[[samp]] <- cell_proximities_list[[samp]]$enrichm_res
}

vr_merged <- merge(vr_list[[1]], vr_list[-1])
vrImageNames(vr_merged)
unique(vr_merged$CellType)

```

Calculate co-enrichment scores and plot them:

```{r, fig.height=4, fig.width=9}

# FOVs for representative overview images
fovs <- vr_list_names

set_alpha <- 0.35
set_ptsize <- 2
set_nrows <- 1
set_ncols <- 3
cell_shape <- 20


cols_con <- c("darkcyan", "gold", "deeppink", 
                "slateblue")

cols_fov <- c("darkcyan", "gold", "deeppink", 
                "slateblue")

ColorsCellType <-  list(
  #`NK cells/ILC1s/ILC3s` = "cyan", 
  `NK cells/ILC1s/ILC3s` = "magenta",
  #`ILC3s` = "magenta", 
  `EMCN CD31 Blood vessels` = "green")

uni_celltypes <- unique(vr_merged$CellType)
backgroundlist <- list("EpCAM","EpCAM", "PDPN", "CD31", "LYVE1","CD11c","B220","Kappa", "GATA3eGFP", "CD8a", "DAPI", "EOMES", "CD4", "CD8a")
names(backgroundlist) <- uni_celltypes
uni_celltypes <- uni_celltypes[!uni_celltypes %in% "NK cells/ILC1s/ILC3s"]

g_master_list <- list()
for(unic in uni_celltypes){
  
    ### selected cell groups ####
  selected_celltypes <- c("NK cells/ILC1s/ILC3s", unic)
  interactions <- unique(cell_proximities_list[[vr_list_names[30]]]$unified_int)
  interactions <- interactions[grepl("NK cells/ILC1s/ILC3s", interactions) & grepl(unic, interactions)]
  background_image <- backgroundlist[[unic]]
  if(length(interactions) > 0){
     ### get interaction results ####
    interaction_celltypes <- NULL
    for(samp in vr_list_names){
      cur_cell_proximities <- cell_proximities_list[[samp]]
      cur_cell_proximities <- cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,]
      sample <- unique(metadatax$Dataset[metadatax$Dataset==samp])
      if(nrow(cur_cell_proximities) > 0 & sample != "20210906_FOV3_D3"){
        interaction_celltypes <- rbind(interaction_celltypes,
                                       data.frame(cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,], 
                                                  experiment = strsplit(sample, split = "_")[[1]][1], fov = strsplit(sample, split = "_")[[1]][2], condition = strsplit(sample, split = "_")[[1]][3]))
      }
    }
    interaction_celltypes$p.adj <- ifelse(interaction_celltypes$enrichm > 0, interaction_celltypes$p.adj_higher, interaction_celltypes$p.adj_lower)


    
    sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("*"), ""))

    g_test <- ggplot(interaction_celltypes, aes(x = experiment, y = enrichm, fill = experiment)) +
      geom_bar(stat = "identity", position = position_dodge2(width=0.9, preserve = "single")) +
      facet_grid(.~experiment, scales = "free_x") +
      geom_text(aes(label=sig_label), position=position_dodge2(width=0.9, preserve = "single"), 
                angle = 90, hjust = -0.02, size = 4) +
      ylim(-2,3.5)+
      NoLegend()+
      theme_classic2()+
      scale_fill_manual(values = cols_fov, name = "") +
      theme(axis.text.x = element_text(#angle = 50,
                                       vjust = 1, size = 12, hjust = 0.5, face = "bold"
                                       ),
            axis.text.y = element_text(hjust = 0.5, size = 12),
            axis.title.x = element_blank(),
            axis.title.y = element_text(size = 12),
            plot.title = element_text(size =14, hjust = 0.5),
            plot.margin = margin(0, 0.5, 0.5, 0.5, "cm"),
            legend.position = "none",
            strip.background=element_blank(),
            strip.background.x= element_blank(),
            strip.text.x = element_text(size = 1, color = "white"),
            panel.grid.major.y = element_line())+
      NoLegend()+  
      ggtitle(gsub("LYVE1 CD90 ", "", interactions)) +
      ylab("Enrichment")
    
    
    g_master_list[[unic]] <- g_test+
        theme(plot.margin = margin(0, 0.5, 0, 0.5, "cm"))
  }
}

# ILC2s around ILC2s -----------------------------------------
interactions <- c("NK cells/ILC1s/ILC3s--NK cells/ILC1s/ILC3s")
unic <- "NK cells/ILC1s/ILC3s"
background_image <- backgroundlist[["NK cells/ILC1s/ILC3s"]]
interaction_celltypes <- NULL
for(samp in vr_list_names){
    cur_cell_proximities <- cell_proximities_list[[samp]]
    cur_cell_proximities <- cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,]
    sample <- unique(metadatax$Dataset[metadatax$Dataset==samp])
    if(nrow(cur_cell_proximities) > 0 & sample != "20210906_FOV3_D3"){
      interaction_celltypes <- rbind(interaction_celltypes,
                                     data.frame(cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,], 
                                                experiment = strsplit(sample, split = "_")[[1]][1], fov = strsplit(sample, split = "_")[[1]][2], condition = strsplit(sample, split = "_")[[1]][3]))
    }
  }
interaction_celltypes$p.adj <- ifelse(interaction_celltypes$enrichm > 0, interaction_celltypes$p.adj_higher, interaction_celltypes$p.adj_lower)
  # plot test results
  # sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("p=",round(interaction_celltypes$p.adj,3)), ""))
  sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("*"), ""))
    # print(sig_label)
g_test <- ggplot(interaction_celltypes, aes(x = experiment, y = enrichm, fill = experiment)) +
  geom_bar(stat = "identity", position = position_dodge2(width=0.9, preserve = "single")) +
  facet_grid(.~experiment, scales = "free_x") +
  geom_text(aes(label=sig_label), position=position_dodge2(width=0.9, preserve = "single"), 
            angle = 90, hjust = -0.02, size = 4) +
  ylim(-2,3.5)+
  NoLegend()+
  theme_classic2()+
  scale_fill_manual(values = cols_fov, name = "") +
  theme(axis.text.x = element_text(#angle = 50,
                                   vjust = 1, size = 12, hjust = 0.5, face = "bold"
                                   ),
        axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        plot.margin = margin(0, 0.5, 0.5, 0.5, "cm"),
        legend.position = "none",
        strip.background=element_blank(),
        strip.background.x= element_blank(),
        strip.text.x = element_text(size = 1, color = "white"),
        panel.grid.major.y = element_line())+
  NoLegend()+  
  ggtitle(interactions) +
  ylab("Enrichment")


g_master_list[[unic]] <- g_test+
    theme(plot.margin = margin(0, 0.5, 0, 0.5, "cm"))

ggarrange(plotlist = g_master_list)
```

# Visualization

## Co-enrichment of NK cells/ILC1s/ILC3s with epithelia II cells in SI villi

```{r}

# fine tune the co-enrichment plot
# ILC2s around ILC2s -----------------------------------------
interactions <- c("Epithelia II--NK cells/ILC1s/ILC3s")
unic <- "Epithelia II"
background_image <- backgroundlist[["Epithelia II"]]
interaction_celltypes <- NULL
for(samp in vr_list_names){
    cur_cell_proximities <- cell_proximities_list[[samp]]
    cur_cell_proximities <- cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,]
    sample <- unique(metadatax$Dataset[metadatax$Dataset==samp])
    if(nrow(cur_cell_proximities) > 0 & sample != "20210906_FOV3_D3"){
      interaction_celltypes <- rbind(interaction_celltypes,
                                     data.frame(cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,], 
                                                experiment = strsplit(sample, split = "_")[[1]][1], fov = strsplit(sample, split = "_")[[1]][2], condition = strsplit(sample, split = "_")[[1]][3]))
    }
  }
interaction_celltypes$p.adj <- ifelse(interaction_celltypes$enrichm > 0, interaction_celltypes$p.adj_higher, interaction_celltypes$p.adj_lower)
  # plot test results
  # sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("p=",round(interaction_celltypes$p.adj,3)), ""))
  sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("*"), ""))
    # print(sig_label)
plot_coenrichment <- ggplot(interaction_celltypes, aes(x = experiment, y = enrichm, fill = experiment)) +
  geom_bar(stat = "identity", position = position_dodge2(width=0.9, preserve = "single")) +
  facet_grid(.~experiment, scales = "free_x") +
  geom_text(aes(label=sig_label), position=position_dodge2(width=0.9, preserve = "single"), angle = 90, hjust = -0.02, size = 4) +
  ylim(-2,2.1)+
  NoLegend()+
  theme_classic2()+
  scale_fill_manual(values = cols_treat, name = "") +
  ggtitle(gsub("LYVE1 CD90 ", "", interactions)) +
  theme(axis.text.x = element_text(#angle = 50,
                                   vjust = 1, size = 12, hjust = 0.5, face = "bold"
                                   ),
        axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
        legend.position = "none",
        strip.background=element_blank(),
        strip.background.x= element_blank(),
        strip.text.x = element_text(size = 1, color = "white"),
        panel.grid.major.y = element_line())+
  NoLegend()+  
  ylab("Enrichment")

plot_coenrichment
```

## IF example image of re-distribution of NK cells/ILC1s/ILC3s after IL-33 application

```{r, fig.width=4.5, fig.height=4.5}

ColorsCellType <-  list(
  #`NK cells/ILC1s/ILC3s` = "cyan", 
  `NK cells/ILC1s/ILC3s` = "magenta",
  #`ILC3s` = "magenta", 
  `Epithelia II` = "green")


cell_type_of_interest <- c("NK cells/ILC1s/ILC3s", "Epithelia II")

set_ptsize <- 1
cell_shape <- 18
set_alpha <- 1

plot_ctrl <- vrSpatialPlot(vr_merged, assay = paste0("Assay", 6), 
                        group.by = "CellType", 
                        group.ids = cell_type_of_interest,
                        alpha = set_alpha, 
                        background = c("image_1", "EpCAM"), 
                        pt.size = set_ptsize, cell.shape = cell_shape)+
  guides(color = guide_legend(override.aes = list(size = 5)))+
  scale_color_manual(values = ColorsCellType)+
  scale_fill_manual(values = ColorsCellType)+
    theme(plot.title = element_text(size = 14), 
          legend.position = "bottom", 
          legend.text = element_text(size = 12), 
          plot.margin = margin(0, 0, 0, 0, "cm"), 
          legend.key = element_blank())+
  ggtitle("CTRL")

plot_ctrl


plot_d1 <- vrSpatialPlot(vr_merged, assay = paste0("Assay", 13), 
                        group.by = "CellType", 
                        group.ids = cell_type_of_interest,
                        alpha = set_alpha, 
                        background = c("image_1", "EpCAM"), 
                        pt.size = set_ptsize, cell.shape = cell_shape)+ 
  guides(color = guide_legend(override.aes = list(size = 5)))+
  scale_color_manual(values = ColorsCellType)+
  scale_fill_manual(values = ColorsCellType)+
    theme(plot.title = element_text(size = 14), 
          legend.position = "bottom", 
          legend.text = element_text(size = 12),
          plot.margin = margin(0, 0, 0, 0, "cm"), 
          legend.key = element_blank())+
  ggtitle("IL-33 Day 1")

plot_d1
```

## ILC count @CTRL in SI villi and SI ILF

```{r}

df_si <- rbind(df_villi, df_ilf)

# select only the relevent columns for comparison
df_si <- df_si %>%
  filter(Treatment == "CTRL") %>%
  mutate(Tissue.area = paste0("SI ", Tissue.area))%>%
  select(Dataset, Tissue.area, ILCs, `Immune cells`, TotalCellCountFOV)

# do the same fr the lung data
df_lung <- df_lung %>%
  filter(Treatment == "CTRL") %>%
  select(Dataset, `Tissue area`, ILCs, `Immune cells`, TotalCellCountFOV) %>%
  mutate(`Tissue.area` = `Tissue area`) %>%
  select(-`Tissue area`)

df_all <- rbind(df_si, df_lung)

df_all$Tissue.area <- factor(df_all$Tissue.area, levels = c("Lung", "SI Villi", "SI ILF"))

# filter for CTRL and convert to longer format
df <- df_all %>%
  select(Tissue.area, Dataset, ILCs) %>%
  mutate(`Tissue area` = Tissue.area, 
         Tissue.area = factor(Tissue.area, level = c("Villi", "ILF")))

# Testing for normal distribution
shapiro.test(df$ILCs)
moments::kurtosis(df$ILCs)
moments::jarque.test(df$ILCs)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`ILCs` ~ `Tissue area`)
res.kruskal
df %>% kruskal_effsize(`ILCs` ~ `Tissue area`)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`ILCs` ~ `Tissue.area`, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ `Tissue.area`, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Tissue.area")

plot <- ggplot(df, aes(x = `Tissue area`, y = ILCs, fill = "Tissue.area"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = `Tissue.area`), size = 2, cex = 3)+
  scale_color_manual(values = cols_ilcs_si)+
  theme_classic2()+
  theme(plot.margin=margin(0,0.5,1,1,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
        axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  # stat_pvalue_manual(pwc,
  #                      hide.ns = TRUE, size = 6,
  #                      step.increase = 0.1, y.position = 400) +
  ggtitle("ILCs")+
  xlab(NULL)+
  ylab("Count per FOV [#]")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,450))+
    # annotate(geom = 'text',
    #        x=1.5,
    #        y=280,
    #        label=Labels[1],
    #        #angle = 90,
    #        size = 10/.pt)+
  NoLegend()

plot

# df %>% group_by()%>% summarise(median = median(value, na.rm = TRUE))
```

## Count of ILCs \@CTRL conditions

### SI villi

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_villi %>%
  filter(Treatment == "CTRL") %>%
  select(Treatment, Dataset, `NK cells/ILC1s/ILC3s`, ILC2s) %>%
  tidyr::pivot_longer(cols = c(`NK cells/ILC1s/ILC3s`, ILC2s), names_to = "ILCtype") %>%
  mutate(ILCtype = factor(ILCtype, level = c(
    "NK cells/ILC1s/ILC3s", "ILC2s"
  )))

# Testing for normal distribution
shapiro.test(df$value)
moments::kurtosis(df$value)
moments::jarque.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILCtype)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILCtype)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILCtype, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILCtype, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILCtype")

plot_prop_villi <- ggplot(df, aes(x = ILCtype, y = value, fill = "ILCtype"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILCtype), size = 2, cex = 3)+
  scale_color_manual(values = cols_ilcs_si)+
  theme_classic2()+
  theme(plot.margin=margin(0,0.5,1,1,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
        axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 20) +
  ggtitle("SI villi")+
  xlab(NULL)+
  ylab("Count per FOV [#]")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,300))+
    annotate(geom = 'text',
           x=1.5,
           y=280,
           label=Labels[1],
           #angle = 90,
           size = 10/.pt)+
  NoLegend()

plot_prop_villi

df %>% group_by(ILCtype)%>% summarise(median = median(value, na.rm = TRUE))
```

### SI ILF

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_ilf %>%
  filter(Treatment == "CTRL") %>%
  select(Treatment, Dataset, `NK cells/ILC1s/ILC3s`, ILC2s) %>%
  tidyr::pivot_longer(cols = c(`NK cells/ILC1s/ILC3s`, ILC2s), names_to = "ILCtype") %>%
  mutate(ILCtype = factor(ILCtype, level = c(
    "NK cells/ILC1s/ILC3s", "ILC2s"
  )))

# Testing for normal distribution
shapiro.test(df$value)
moments::kurtosis(df$value)
moments::jarque.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILCtype)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILCtype)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILCtype, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILCtype, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILCtype")

plot_prop_ilf <- ggplot(df, aes(x = ILCtype, y = value, fill = "ILCtype"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILCtype), size = 2, cex = 3)+
  scale_color_manual(values = cols_ilcs_si)+
  theme_classic2()+
  theme(plot.margin=margin(0,0.5,1,1,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 20) +
  ggtitle("SI ILF")+
  xlab(NULL)+
  ylab("Count per FOV [#]")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,300))+
    annotate(geom = 'text',
           x=1.5,
           y=280,
           label=Labels[1],
           #angle = 90,
           size = 10/.pt)+
  NoLegend()

plot_prop_ilf

df %>% group_by(ILCtype)%>% summarise(median = median(value, na.rm = TRUE))
```

## Frequencies of ILCs within ILC compartment \@ CTRL conditions

### SI villi

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_villi %>%
  filter(Treatment == "CTRL") %>%
  select(Treatment, Dataset, `Prop_NK cells/ILC1s/ILC3s_perTotalILCsFOV`, Prop_ILC2s_perTotalILCsFOV) %>%
  tidyr::pivot_longer(cols = c(`Prop_NK cells/ILC1s/ILC3s_perTotalILCsFOV`, Prop_ILC2s_perTotalILCsFOV), names_to = "ILCtype") %>%
  mutate(ILCtype = gsub("Prop_|_perTotalILCsFOV", "", ILCtype), 
         ILCtype = factor(ILCtype, level = c(
    "NK cells/ILC1s/ILC3s", "ILC2s"
  )))

# Testing for normal distribution
shapiro.test(df$value)
moments::kurtosis(df$value)
moments::jarque.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILCtype)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILCtype)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILCtype, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILCtype, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILCtype")

plot_freq_villi <- ggplot(df, aes(x = ILCtype, y = value, fill = "ILCtype"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILCtype), size = 2, cex = 3)+
  scale_color_manual(values = cols_ilcs_lung)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
        axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 90) +
  xlab(NULL)+
  ylab("Frequency per FOV/ILCs [%]")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,105))+
  NoLegend()


plot_freq_villi

df %>% group_by(ILCtype)%>% summarise(median = median(value, na.rm = TRUE))
```

### SI ILF

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_ilf %>%
  filter(Treatment == "CTRL") %>%
  select(Treatment, Dataset, `Prop_NK cells/ILC1s/ILC3s_perTotalILCsFOV`, Prop_ILC2s_perTotalILCsFOV) %>%
  tidyr::pivot_longer(cols = c(`Prop_NK cells/ILC1s/ILC3s_perTotalILCsFOV`, Prop_ILC2s_perTotalILCsFOV), names_to = "ILCtype") %>%
  mutate(ILCtype = gsub("Prop_|_perTotalILCsFOV", "", ILCtype), 
         ILCtype = factor(ILCtype, level = c(
    "NK cells/ILC1s/ILC3s", "ILC2s"
  )))

# Testing for normal distribution
shapiro.test(df$value)
moments::kurtosis(df$value)
moments::jarque.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILCtype)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILCtype)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILCtype, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILCtype, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILCtype")

plot_freq_ilf <- ggplot(df, aes(x = ILCtype, y = value, fill = "ILCtype"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILCtype), size = 2, cex = 3)+
  scale_color_manual(values = cols_ilcs_lung)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
        axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 90) +
  xlab(NULL)+
  ylab("Frequency per FOV/ILCs [%]")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,105))+
  NoLegend()


plot_freq_ilf
```

## Frequency of ILCs within immune compartment \@ CTRL

### SI villi

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_villi %>%
  filter(Treatment == "CTRL") %>%
  select(Treatment, Dataset, `Prop_NK cells/ILC1s/ILC3s_perTotalImmuneFOV`, Prop_ILC2s_perTotalImmuneFOV) %>%
  tidyr::pivot_longer(cols = c(`Prop_NK cells/ILC1s/ILC3s_perTotalImmuneFOV`, Prop_ILC2s_perTotalImmuneFOV), names_to = "ILCtype") %>%
  mutate(ILCtype = gsub("Prop_|_perTotalImmuneFOV", "", ILCtype), 
         ILCtype = factor(ILCtype, level = c(
    "NK cells/ILC1s/ILC3s", "ILC2s"
  )))

# Testing for normal distribution
shapiro.test(df$value)
moments::kurtosis(df$value)
moments::jarque.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILCtype)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILCtype)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILCtype, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILCtype, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILCtype")

plot_freq_immune_villi <- ggplot(df, aes(x = ILCtype, y = value, fill = "ILCtype"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILCtype), size = 2, cex = 3)+
  scale_color_manual(values = cols_ilcs_lung)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1.5,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
        axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 11) +
  xlab(NULL)+
  ylab("Frequency per FOV/Immune [%]")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,14))+
  NoLegend()

plot_freq_immune_villi
```

### SI ILF

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_ilf %>%
  filter(Treatment == "CTRL") %>%
  select(Treatment, Dataset, `Prop_NK cells/ILC1s/ILC3s_perTotalImmuneFOV`, Prop_ILC2s_perTotalImmuneFOV) %>%
  tidyr::pivot_longer(cols = c(`Prop_NK cells/ILC1s/ILC3s_perTotalImmuneFOV`, Prop_ILC2s_perTotalImmuneFOV), names_to = "ILCtype") %>%
  mutate(ILCtype = gsub("Prop_|_perTotalImmuneFOV", "", ILCtype), 
         ILCtype = factor(ILCtype, level = c(
    "NK cells/ILC1s/ILC3s", "ILC2s"
  )))

# Testing for normal distribution
shapiro.test(df$value)
moments::kurtosis(df$value)
moments::jarque.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILCtype)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILCtype)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILCtype, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILCtype, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILCtype")

plot_freq_immune_ilf <- ggplot(df, aes(x = ILCtype, y = value, fill = "ILCtype"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILCtype), size = 2, cex = 3)+
  scale_color_manual(values = cols_ilcs_lung)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1.5,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
        axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 11) +
  xlab(NULL)+
  ylab("Frequency per FOV/Immune [%]")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,14))+
  NoLegend()

plot_freq_immune_ilf
```

## Total cell count IL-33

We will only analyze SI villi data as we only have n=1 for D1 and D3 of the SI ILF data.

```{r, fig.width=3, fig.height=3}

# filter for CTRL and convert to longer format
df <- df_villi %>%
  select(Treatment, Dataset, TotalCellCountFOV) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "1", "3"
  )))

# Testing for normal distribution
shapiro.test(df$TotalCellCountFOV)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`TotalCellCountFOV` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`TotalCellCountFOV` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`TotalCellCountFOV` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_count_all <- ggplot(df, aes(x = Treatment, y = TotalCellCountFOV, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(#angle = 45, 
                                   vjust = 1, size = 12, hjust = 0.5, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 6800) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("Total cells")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,7200))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=5000,
           label=Labels,
           #angle = 90,
           size = 10/.pt)

plot_count_all
```

## Total immune count IL-33

```{r, fig.width=3, fig.height=3}

# filter for CTRL and convert to longer format
df <- df_villi %>%
  select(Treatment, Dataset, `Immune cells`) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "1", "3"
  )))

# Testing for normal distribution
shapiro.test(df$`Immune cells`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`Immune cells` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`Immune cells` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`Immune cells` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_count_immune <- ggplot(df, aes(x = Treatment, y = `Immune cells`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(#angle = 45, 
                                   vjust = 1, size = 12, hjust = 0.5, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 1750) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("Immune cells")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,2300))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=2100,
           label=Labels,
           #angle = 90,
           size = 10/.pt)

plot_count_immune
```

## Total ILC count IL-33

```{r, fig.width=3, fig.height=3}

# filter for CTRL and convert to longer format
df <- df_villi %>%
  select(Treatment, Dataset, ILCs) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "1", "3"
  )))

# Testing for normal distribution
shapiro.test(df$ILCs)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`ILCs` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`ILCs` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`ILCs` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_count_ilc <- ggplot(df, aes(x = Treatment, y = ILCs, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(#angle = 45, 
                                   vjust = 1, size = 12, hjust = 0.5, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.08, y.position = 240) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("ILCs")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,300))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=280,
           label=Labels,
           #angle = 90,
           size = 10/.pt)

plot_count_ilc
```

## Total count ILC1s/NK cells

```{r, fig.width=3, fig.height=3}

# filter for CTRL and convert to longer format
df <- df_villi %>%
  select(Treatment, Dataset, `NK cells/ILC1s/ILC3s`) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "1", "3"
  )))

# Testing for normal distribution
shapiro.test(df$`NK cells/ILC1s/ILC3s`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`NK cells/ILC1s/ILC3s` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`NK cells/ILC1s/ILC3s` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`NK cells/ILC1s/ILC3s` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_count_ilc1 <- ggplot(df, aes(x = Treatment, y = `NK cells/ILC1s/ILC3s`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(#angle = 45, 
                                   vjust = 1, size = 12, hjust = 0.5, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 300) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("NK cells/ILC1s/ILC3s")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,400))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=390,
           label=Labels,
           #angle = 90,
           size = 10/.pt)

plot_count_ilc1
```

## Total count ILC2s

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_villi %>%
  select(Treatment, Dataset, `ILC2s`) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "1", "3"
  )))

# Testing for normal distribution
shapiro.test(df$`ILC2s`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`ILC2s` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`ILC2s` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`ILC2s` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_count_ilc2 <- ggplot(df, aes(x = Treatment, y = `ILC2s`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(#angle = 45, 
                                   vjust = 1, size = 12, hjust = 0.5, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 100) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("ILC2s")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,170))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=160,
           label=Labels,
           #angle = 90,
           size = 10/.pt)

plot_count_ilc2
```

## Freq of immune cells, stromal cells and epithelia within total cells

```{r, fig.width=9, fig.height=3}

# IMMUNE CELLS ---------------------------------------------------------------
df <- df_villi %>%
  select(Treatment, Dataset, `Prop_Immune cells_perTotalCountFOV`) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "1", "3"
  )), 
  `Immune cells` = `Prop_Immune cells_perTotalCountFOV`)

# Testing for normal distribution
shapiro.test(df$`Immune cells`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`Immune cells` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`Immune cells` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`Immune cells` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_immune <- ggplot(df, aes(x = Treatment, y = `Immune cells`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(#angle = 45, 
                                   vjust = 1, size = 12, hjust = 0.5, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.5, y.position = 40) +
  xlab(NULL)+
  ylab("Frequency/total cells per FOV [#]")+
  ggtitle("Immune cells")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,100))+
  NoLegend()


# Stromal cells ---------------------------------------------------------------
df <- df_villi %>%
  select(Treatment, Dataset, `Prop_Vessels & stroma_perTotalCountFOV`) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "1", "3"
  )), 
  `cells_of_interest` = `Prop_Vessels & stroma_perTotalCountFOV`)

# Testing for normal distribution
shapiro.test(df$`cells_of_interest`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`cells_of_interest` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`cells_of_interest` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`cells_of_interest` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_stroma <- ggplot(df, aes(x = Treatment, y = `cells_of_interest`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(#angle = 45, 
                                   vjust = 1, size = 12, hjust = 0.5, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.2, y.position = 80) +
  xlab(NULL)+
  ylab("Frequency/total cells per FOV [#]")+
  ggtitle("Vessels & stroma")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,100))+
  NoLegend()


# Epithelia ---------------------------------------------------------------
df <- df_villi %>%
  select(Treatment, Dataset, `Prop_Epithelia_perTotalCountFOV`) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "1", "3"
  )), 
  `cells_of_interest` = `Prop_Epithelia_perTotalCountFOV`)

# Testing for normal distribution
shapiro.test(df$`cells_of_interest`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`cells_of_interest` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`cells_of_interest` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`cells_of_interest` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_epithelia <- ggplot(df, aes(x = Treatment, y = `cells_of_interest`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(#angle = 45, 
                                   vjust = 1, size = 12, hjust = 0.5, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.15, y.position = 40) +
  xlab(NULL)+
  ylab("Frequency/total cells per FOV [#]")+
  ggtitle("Epithelia")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,100))+
  NoLegend()


ggarrange(plot_immune, plot_stroma, plot_epithelia, 
          ncol = 3, nrow = 1, 
          labels = c("A", "B", "C", "D", "E", "F", "G", "H", "I"))+
  theme(plot.margin = margin(0, 0.1, 0, 0, "cm"))
```

## Combine plots for figure

```{r, fig.width=9, fig.height=13}
ggarrange(plot_prop_villi, plot_prop_ilf, plot_freq_immune_villi, plot_freq_villi, 
          plot_freq_immune_ilf, plot_freq_ilf, 
          plot_count_all, plot_count_immune, plot_count_ilc, 
          plot_count_ilc1, plot_count_ilc2, "NONE", 
          ncol = 3, nrow = 4, heights = c(3.5, 3.5, 3, 3), 
          labels = c("A", "", "B", "C", "D", "E", "F", "G", "H", "I"))+
  theme(plot.margin = margin(0, 0.1, 0, 0, "cm"))
```

```{r, fig.width=9, fig.height=3}
ggarrange(plot_freq_ilc1, plot_freq_ilc2, 
          ncol = 3, nrow = 1, 
          labels = c("A", "B", "C", "D", "E", "F", "G", "H", "I"))+
  theme(plot.margin = margin(0, 0.1, 0, 0, "cm"))
```

## Session Information

```{r}
save.image(paste0(output_dir, "/environment.RData"))
sessionInfo()
```
