---
title: "Figure 5: Quantification lung"
author: "Sandy Kroh"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    fig_caption: yes
    keep_md: yes
  pdf_document:
    toc: yes
params:
  Sample: Default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", 
                      fig.align='center', dpi = 600, 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Libraries

```{r}
library(Seurat)
library(SeuratObject)
library(dplyr)
library(ggplot2)
library(rstatix)
library(ggbeeswarm)
library(readr)
library(ggpubr)

```

## Parameters

```{r}
set.seed(123)

input_dir <- here::here("1_data_tidying", "Lung_SI_all_cells_all_ALs_files")

output_dir <- here::here("2_visualizations_for_figures", "Fig_5_quantification_lung_files")
dir.create(output_dir)




main_markers <- c(
  "EpCAM", "EMCN", "LYVE1", "PDPN", "PDGFRa", "CD8a", "CD4",
  "CD45", "CD3", "IRF4", "Kappa", "CD11c", "CD127", "GATA3eGFP", "RORgt"
)


immune_markers <- c(
 "CD3", "CD4", "CD8a", "Kappa", "IRF4", "CD11c",
  "CD127", "CD90", "EOMES", "GATA3eGFP", "RORgt", "Ki67",  "KLRG1", "NKp46", "CD117", "Areg", "CCR6", "CD44", "MHCII", "Sca1"
)

ilc_markers <- c(
  "CD3", "CD4", "CD8a",
  "CD127", "CD90", "EOMES", "GATA3eGFP", "RORgt", "KLRG1", "NKp46", "CD117", "CCR6", "MHCII", "Ki67", "Areg", "IRF4", "Sca1", "CD44"
)


cols_nat <- c("magenta", "cyan", "blue", "purple", "green", 
                       "red", "yellow", "olivedrab1", "slateblue1", 
                       "darkcyan", "gold","indianred1", "seagreen", "deeppink", 
                       "orange", "brown", "violet",
                       "deeppink4", "pink", 
                       "grey", "black", "lightgreen", 
                       "#FF0066",  
                       "lightblue", "#FFCC99", "#CC00FF", 
                       "blueviolet",  "goldenrod4", 
                       "navy", "olivedrab", "lightcyan", "seagreen2", "darkviolet", "lightpink", "slateblue4", "olivedrab2")

colfunc <- colorRampPalette(c("darkcyan", "green", "yellow", "magenta", "purple"))

cols_ilcs_lung <- c("darkcyan", "seagreen2", "deeppink4")

cols_treat <- c("darkcyan", "gold", "deeppink", "slateblue")

```

# Load data

```{r, fig.width=9, fig.height=6}

df_lung <- read_csv(paste0(input_dir, "/lung_proportions.csv"), 
    col_types = cols(...1 = col_skip()))

```

# Visualization

## 4A - Count of ILCs \@CTRL conditions

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  filter(Treatment == "CTRL") %>%
  select(Treatment, Dataset, `NK cells/ILC1s`, ILC2s, ILC3s) %>%
  tidyr::pivot_longer(cols = c(`NK cells/ILC1s`, ILC2s, ILC3s), names_to = "ILCtype") %>%
  mutate(ILCtype = factor(ILCtype, level = c(
    "NK cells/ILC1s", "ILC2s", "ILC3s"
  )))

# Testing for normal distribution
shapiro.test(df$value)
moments::kurtosis(df$value)
moments::jarque.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILCtype)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILCtype)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILCtype, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILCtype, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILCtype")

plot_prop <- ggplot(df, aes(x = ILCtype, y = value, fill = "ILCtype"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILCtype), size = 2, cex = 3)+
  scale_color_manual(values = cols_ilcs_lung)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 20) +
  xlab(NULL)+
  ylab("Count per FOV [#]")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,24))+
    # annotate(geom = 'text',
    #        x="ILC2s",
    #        y=28,
    #        label=Labels[1], 
    #        #angle = 90, 
    #        size = 10/.pt)+
  NoLegend()

plot_prop

```

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  filter(Treatment == "3") %>%
  select(Treatment, Dataset, `NK cells/ILC1s`, ILC2s, ILC3s) %>%
  tidyr::pivot_longer(cols = c(`NK cells/ILC1s`, ILC2s, ILC3s), names_to = "ILCtype") %>%
  mutate(ILCtype = factor(ILCtype, level = c(
    "NK cells/ILC1s", "ILC2s", "ILC3s"
  )))

# Testing for normal distribution
shapiro.test(df$value)
moments::kurtosis(df$value)
moments::jarque.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILCtype)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILCtype)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILCtype, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILCtype, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILCtype")

plot_d3 <- ggplot(df, aes(x = ILCtype, y = value, fill = "ILCtype"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILCtype), size = 2, cex = 3)+
  scale_color_manual(values = cols_ilcs_lung)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 180) +
  xlab(NULL)+
  ylab("Count per FOV [#]")+
  ggtitle("ILC subtypes D3")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,250))+
    # annotate(geom = 'text',
    #        x="ILC2s",
    #        y=28,
    #        label=Labels[1], 
    #        #angle = 90, 
    #        size = 10/.pt)+
  NoLegend()

plot_d3

```

## 4C - Frequencies of ILCs within ILC compartment \@ CTRL conditions

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  filter(Treatment == "CTRL") %>%
  select(Treatment, Dataset, `Prop_NK cells/ILC1s_perTotalILCsFOV`, Prop_ILC2s_perTotalILCsFOV, Prop_ILC3s_perTotalILCsFOV) %>%
  tidyr::pivot_longer(cols = c(`Prop_NK cells/ILC1s_perTotalILCsFOV`, Prop_ILC2s_perTotalILCsFOV, Prop_ILC3s_perTotalILCsFOV), names_to = "ILCtype") %>%
  mutate(ILCtype = gsub("Prop_|_perTotalILCsFOV", "", ILCtype), 
         ILCtype = factor(ILCtype, level = c(
    "NK cells/ILC1s", "ILC2s", "ILC3s"
  )))

# Testing for normal distribution
shapiro.test(df$value)
moments::kurtosis(df$value)
moments::jarque.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILCtype)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILCtype)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILCtype, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILCtype, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILCtype")

plot_freq <- ggplot(df, aes(x = ILCtype, y = value, fill = "ILCtype"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILCtype), size = 2, cex = 3)+
  scale_color_manual(values = cols_ilcs_lung)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 90) +
  xlab(NULL)+
  ylab("Frequency per FOV/ILCs [%]")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,105))+
  NoLegend()


plot_freq
```

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  filter(Treatment == "3") %>%
  select(Treatment, Dataset, `Prop_NK cells/ILC1s_perTotalILCsFOV`, Prop_ILC2s_perTotalILCsFOV, Prop_ILC3s_perTotalILCsFOV) %>%
  tidyr::pivot_longer(cols = c(`Prop_NK cells/ILC1s_perTotalILCsFOV`, Prop_ILC2s_perTotalILCsFOV, Prop_ILC3s_perTotalILCsFOV), names_to = "ILCtype") %>%
  mutate(ILCtype = gsub("Prop_|_perTotalILCsFOV", "", ILCtype), 
         ILCtype = factor(ILCtype, level = c(
    "NK cells/ILC1s", "ILC2s", "ILC3s"
  )))

# Testing for normal distribution
shapiro.test(df$value)
moments::kurtosis(df$value)
moments::jarque.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILCtype)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILCtype)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILCtype, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILCtype, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILCtype")

plot_freq_d3 <- ggplot(df, aes(x = ILCtype, y = value, fill = "ILCtype"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILCtype), size = 2, cex = 3)+
  scale_color_manual(values = cols_ilcs_lung)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.subtitle = element_text(hjust = 0.5, size = 11), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 90) +
  xlab(NULL)+
  ylab("Frequency per FOV/ILCs [%]")+
  ggtitle("D3", subtitle = "ILC compartment")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,119))+
  NoLegend()


plot_freq_d3
```

## 4B - Frequency of ILCs within immune compartment \@ CTRL

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  filter(Treatment == "CTRL") %>%
  select(Treatment, Dataset, `Prop_NK cells/ILC1s_perTotalImmuneFOV`, Prop_ILC2s_perTotalImmuneFOV, Prop_ILC3s_perTotalImmuneFOV) %>%
  tidyr::pivot_longer(cols = c(`Prop_NK cells/ILC1s_perTotalImmuneFOV`, Prop_ILC2s_perTotalImmuneFOV, Prop_ILC3s_perTotalImmuneFOV), names_to = "ILCtype") %>%
  mutate(ILCtype = gsub("Prop_|_perTotalImmuneFOV", "", ILCtype), 
         ILCtype = factor(ILCtype, level = c(
    "NK cells/ILC1s", "ILC2s", "ILC3s"
  )))

# Testing for normal distribution
shapiro.test(df$value)
moments::kurtosis(df$value)
moments::jarque.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILCtype)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILCtype)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILCtype, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILCtype, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILCtype")

plot_freq_immune <- ggplot(df, aes(x = ILCtype, y = value, fill = "ILCtype"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILCtype), size = 2, cex = 3)+
  scale_color_manual(values = cols_ilcs_lung)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 11) +
  xlab(NULL)+
  ylab("Frequency per FOV/Immune [%]")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,14))+
  NoLegend()

plot_freq_immune
```

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  filter(Treatment == "3") %>%
  select(Treatment, Dataset, `Prop_NK cells/ILC1s_perTotalImmuneFOV`, Prop_ILC2s_perTotalImmuneFOV, Prop_ILC3s_perTotalImmuneFOV) %>%
  tidyr::pivot_longer(cols = c(`Prop_NK cells/ILC1s_perTotalImmuneFOV`, Prop_ILC2s_perTotalImmuneFOV, Prop_ILC3s_perTotalImmuneFOV), names_to = "ILCtype") %>%
  mutate(ILCtype = gsub("Prop_|_perTotalImmuneFOV", "", ILCtype), 
         ILCtype = factor(ILCtype, level = c(
    "NK cells/ILC1s", "ILC2s", "ILC3s"
  )))

# Testing for normal distribution
shapiro.test(df$value)
moments::kurtosis(df$value)
moments::jarque.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILCtype)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILCtype)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILCtype, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILCtype, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILCtype")

plot_freq_immune_d3 <- ggplot(df, aes(x = ILCtype, y = value, fill = "ILCtype"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILCtype), size = 2, cex = 3)+
  scale_color_manual(values = cols_ilcs_lung)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.subtitle = element_text(hjust = 0.5, size = 11), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 15) +
  xlab(NULL)+
  ylab("Frequency per FOV/Immune [%]")+
  ggtitle("D3", subtitle = "Immune compartment")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,20))+
  NoLegend()

plot_freq_immune_d3
```

## Total cell count IL-33

```{r, fig.width=3, fig.height=3}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  select(Treatment, Dataset, TotalCellCountFOV) %>%
  mutate(Treatment = paste0("D", Treatment)) %>%
  mutate(Treatment = gsub("DCTRL", "CTRL", Treatment)) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "D1", "D2", "D3"
  )))

# Testing for normal distribution
shapiro.test(df$TotalCellCountFOV)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`TotalCellCountFOV` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`TotalCellCountFOV` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`TotalCellCountFOV` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_count_all <- ggplot(df, aes(x = Treatment, y = TotalCellCountFOV, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 5200) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("Total cells")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,6800))+
  NoLegend()
# +
#     annotate(geom = 'text',
#            x="ILC2s",
#            y=28,
#            label=Labels[1], 
#            #angle = 90, 
#            size = 10/.pt)

plot_count_all
```

## Total immune count IL-33

```{r, fig.width=3, fig.height=3}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  select(Treatment, Dataset, `Immune cells`) %>%
  mutate(Treatment = paste0("D", Treatment)) %>%
  mutate(Treatment = gsub("DCTRL", "CTRL", Treatment)) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "D1", "D2", "D3"
  )))

# Testing for normal distribution
shapiro.test(df$`Immune cells`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`Immune cells` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`Immune cells` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`Immune cells` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_count_immune <- ggplot(df, aes(x = Treatment, y = `Immune cells`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 1750) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("Immune cells")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,2300))+
  NoLegend()
# +
#     annotate(geom = 'text',
#            x="ILC2s",
#            y=28,
#            label=Labels[1], 
#            #angle = 90, 
#            size = 10/.pt)

plot_count_immune
```

## Total ILC count IL-33

```{r, fig.width=3, fig.height=3}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  select(Treatment, Dataset, ILCs) %>%
  mutate(Treatment = paste0("D", Treatment)) %>%
  mutate(Treatment = gsub("DCTRL", "CTRL", Treatment)) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "D1", "D2", "D3"
  )))

# Testing for normal distribution
shapiro.test(df$ILCs)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`ILCs` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`ILCs` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`ILCs` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_count_ilc <- ggplot(df, aes(x = Treatment, y = ILCs, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 180) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("ILCs")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,250))+
  NoLegend()
# +
#     annotate(geom = 'text',
#            x="ILC2s",
#            y=28,
#            label=Labels[1], 
#            #angle = 90, 
#            size = 10/.pt)

plot_count_ilc
```

## Total count ILC1s/NK cells

```{r, fig.width=3, fig.height=3}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  select(Treatment, Dataset, `NK cells/ILC1s`) %>%
  mutate(Treatment = paste0("D", Treatment)) %>%
  mutate(Treatment = gsub("DCTRL", "CTRL", Treatment)) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "D1", "D2", "D3"
  )))

# Testing for normal distribution
shapiro.test(df$`NK cells/ILC1s`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`NK cells/ILC1s` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`NK cells/ILC1s` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`NK cells/ILC1s` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_count_ilc1 <- ggplot(df, aes(x = Treatment, y = `NK cells/ILC1s`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 65) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("NK cells/ILC1s")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,78))+
  NoLegend()
# +
#     annotate(geom = 'text',
#            x="ILC2s",
#            y=28,
#            label=Labels[1], 
#            #angle = 90, 
#            size = 10/.pt)

plot_count_ilc1
```

## Total count ILC2s

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  select(Treatment, Dataset, `ILC2s`) %>%
  mutate(Treatment = paste0("D", Treatment)) %>%
  mutate(Treatment = gsub("DCTRL", "CTRL", Treatment)) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "D1", "D2", "D3"
  )))

# Testing for normal distribution
shapiro.test(df$`ILC2s`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`ILC2s` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`ILC2s` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`ILC2s` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_count_ilc2 <- ggplot(df, aes(x = Treatment, y = `ILC2s`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 180) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("ILC2s")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,250))+
  NoLegend()
# +
#     annotate(geom = 'text',
#            x="ILC2s",
#            y=28,
#            label=Labels[1], 
#            #angle = 90, 
#            size = 10/.pt)

plot_count_ilc2
```

## Total count ILC3s

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  select(Treatment, Dataset, `ILC3s`) %>%
  mutate(Treatment = paste0("D", Treatment)) %>%
  mutate(Treatment = gsub("DCTRL", "CTRL", Treatment)) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "D1", "D2", "D3"
  )))

# Testing for normal distribution
shapiro.test(df$`ILC3s`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`ILC3s` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`ILC3s` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`ILC3s` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_count_ilc3 <- ggplot(df, aes(x = Treatment, y = `ILC3s`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 11) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("ILC3s")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,14))+
  NoLegend()
# +
#     annotate(geom = 'text',
#            x="ILC2s",
#            y=28,
#            label=Labels[1], 
#            #angle = 90, 
#            size = 10/.pt)

plot_count_ilc3
```

## Freq ILC1s of ILC compartment

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  select(Treatment, Dataset, `Prop_NK cells/ILC1s_perTotalILCsFOV`) %>%
  mutate(Treatment = paste0("D", Treatment)) %>%
  mutate(Treatment = gsub("DCTRL", "CTRL", Treatment)) %>%
  mutate(
    Treatment = factor(Treatment, level =c(
    "CTRL", "D1", "D2", "D3"
  )),
  `NK cells/ILC1s` = `Prop_NK cells/ILC1s_perTotalILCsFOV`)

# Testing for normal distribution
shapiro.test(df$`NK cells/ILC1s`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`NK cells/ILC1s` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`NK cells/ILC1s` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`NK cells/ILC1s` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_freq_ilc1 <- ggplot(df, aes(x = Treatment, y = `NK cells/ILC1s`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 14) +
  xlab(NULL)+
  ylab("Frequency per FOV [%]")+
  ggtitle("NK cells/ILC1s")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,85))+
  NoLegend()
# +
#     annotate(geom = 'text',
#            x="ILC2s",
#            y=28,
#            label=Labels[1], 
#            #angle = 90, 
#            size = 10/.pt)

plot_freq_ilc1
```

## Freq ILC2s of ILC compartment

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  select(Treatment, Dataset, Prop_ILC2s_perTotalILCsFOV) %>%
  mutate(Treatment = paste0("D", Treatment)) %>%
  mutate(Treatment = gsub("DCTRL", "CTRL", Treatment)) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "D1", "D2", "D3"
  )),
  ILC2s = Prop_ILC2s_perTotalILCsFOV)

# Testing for normal distribution
shapiro.test(df$`ILC2s`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`ILC2s` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`ILC2s` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`ILC2s` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_freq_ilc2 <- ggplot(df, aes(x = Treatment, y = `ILC2s`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 14) +
  xlab(NULL)+
  ylab("Frequency per FOV [%]")+
  ggtitle("ILC2s")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,90))+
  NoLegend()
# +
#     annotate(geom = 'text',
#            x="ILC2s",
#            y=28,
#            label=Labels[1], 
#            #angle = 90, 
#            size = 10/.pt)

plot_freq_ilc2
```

## Freq ILC3s of ILC compartment

```{r, fig.width=3, fig.height=4}

# filter for CTRL and convert to longer format
df <- df_lung %>%
  select(Treatment, Dataset, Prop_ILC3s_perTotalILCsFOV) %>%
  mutate(Treatment = paste0("D", Treatment)) %>%
  mutate(Treatment = gsub("DCTRL", "CTRL", Treatment)) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "D1", "D2", "D3"
  )),
  ILC3s = Prop_ILC3s_perTotalILCsFOV)

# Testing for normal distribution
shapiro.test(df$`ILC3s`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`ILC3s` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`ILC3s` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`ILC3s` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_freq_ilc3 <- ggplot(df, aes(x = Treatment, y = `ILC3s`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 14) +
  xlab(NULL)+
  ylab("Frequency per FOV [%]")+
  ggtitle("ILC3s")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,16))+
  NoLegend()
# +
#     annotate(geom = 'text',
#            x="ILC2s",
#            y=28,
#            label=Labels[1], 
#            #angle = 90, 
#            size = 10/.pt)

plot_freq_ilc3
```

## Freq of immune cells, Endothelia & stroma and epithelia within total cells

```{r, fig.width=9, fig.height=3}

# IMMUNE CELLS ---------------------------------------------------------------
df <- df_lung %>%
  select(Treatment, Dataset, `Prop_Immune cells_perTotalCountFOV`) %>%
  mutate(Treatment = paste0("D", Treatment)) %>%
  mutate(Treatment = gsub("DCTRL", "CTRL", Treatment)) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "D1", "D2", "D3"
  )),
  `Immune cells` = `Prop_Immune cells_perTotalCountFOV`)

# Testing for normal distribution
shapiro.test(df$`Immune cells`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`Immune cells` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`Immune cells` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`Immune cells` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_immune <- ggplot(df, aes(x = Treatment, y = `Immune cells`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.5, y.position = 40) +
  xlab(NULL)+
  ylab("Frequency/total cells per FOV [#]")+
  ggtitle("Immune cells")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,100))+
  NoLegend()


# Endothelia & stroma ---------------------------------------------------------------
df <- df_lung %>%
  select(Treatment, Dataset, `Prop_Endothelia & stroma_perTotalCountFOV`) %>%
  mutate(Treatment = paste0("D", Treatment)) %>%
  mutate(Treatment = gsub("DCTRL", "CTRL", Treatment)) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "D1", "D2", "D3"
  )),
  `cells_of_interest` = `Prop_Endothelia & stroma_perTotalCountFOV`)

# Testing for normal distribution
shapiro.test(df$`cells_of_interest`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`cells_of_interest` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`cells_of_interest` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`cells_of_interest` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_stroma <- ggplot(df, aes(x = Treatment, y = `cells_of_interest`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.2, y.position = 80) +
  xlab(NULL)+
  ylab("Frequency/total cells per FOV [#]")+
  ggtitle("Endothelia & stroma")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,100))+
  NoLegend()


# Epithelia ---------------------------------------------------------------
df <- df_lung %>%
  select(Treatment, Dataset, `Prop_Epithelia_perTotalCountFOV`) %>%
  mutate(Treatment = paste0("D", Treatment)) %>%
  mutate(Treatment = gsub("DCTRL", "CTRL", Treatment)) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "D1", "D2", "D3"
  )),
  `cells_of_interest` = `Prop_Epithelia_perTotalCountFOV`)

# Testing for normal distribution
shapiro.test(df$`cells_of_interest`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`cells_of_interest` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`cells_of_interest` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`cells_of_interest` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_epithelia <- ggplot(df, aes(x = Treatment, y = `cells_of_interest`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic() +
  theme(
      axis.text.x = element_text(angle = 45, size = 10, face = "bold", hjust = 1),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
      plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
      legend.position = "none"
    ) +
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.15, y.position = 40) +
  xlab(NULL)+
  ylab("Frequency/total cells per FOV [#]")+
  ggtitle("Epithelia")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,100))+
  NoLegend()


ggarrange(plot_immune, plot_stroma, plot_epithelia, 
          ncol = 3, nrow = 1, 
          labels = c("A", "B", "C", "D", "E", "F", "G", "H", "I"))+
  theme(plot.margin = margin(0, 0.1, 0, 0, "cm"))
```

## Differential marker expression of ILCs across conditions

```{r}

SO.lung <- readRDS(paste0(here::here("1_data_tidying", "Lung_SI_all_cells_all_ALs_files"), "/lung_all_cells_all_ALs.rds"))
dim(SO.lung)

SO.lung$AL1 <- gsub("Vessels", "Stromal cells", SO.lung$AL1)

# 1. Extract the current expression matrix from the active assay
# Note: Using slot = "counts". If your MELC data is only stored in "data", change this.
current_matrix <- GetAssayData(SO.lung, assay = "MELC", slot = "counts")

# 2. Extract TBET from metadata and format it as a 1-row matrix
tbet_values <- matrix(SO.lung$TBET, nrow = 1)
rownames(tbet_values) <- "TBET"
colnames(tbet_values) <- colnames(SO.lung) # Ensure cell names match perfectly

# 3. Bind the TBET row to the bottom of the expression matrix
new_matrix <- rbind(current_matrix, tbet_values)

# 4. Overwrite the existing MELC assay with the updated matrix
SO.lung[["MELC"]] <- CreateAssayObject(counts = new_matrix)

# 5. Clean up: Remove TBET from the metadata so it doesn't cause confusion later
SO.lung$TBET <- NULL

# Verify the fix! TBET should now appear at the bottom of this list:
rownames(SO.lung)
```

```{r, fig.height=4, fig.width=3}
library(Seurat)
library(ComplexHeatmap)
library(ggplotify) # Required for as.ggplot()

set.seed(8)

# --- 1. Subsetting ---
SO.sub <- subset(SO.lung, subset = AL3 %in% c("ILC2s"))
SO.sub$AL3 <- droplevels(as.factor(SO.sub$AL3))

# --- 2. Direct Pseudo-bulk Aggregation ---
ilc_markers <- c("KLRG1", #"CD127", "CD90",
                 "ICOS", 
                 "MHCII", "CD44", "Ki67")

# Calculate the mean directly from Seurat
avg_exp <- AverageExpression(SO.sub, features = ilc_markers, slot = "counts", group.by = "Treatment")
mat_annotated <- avg_exp[[DefaultAssay(SO.sub)]]
# --- 3. Manually Scale (Z-score) the Matrix ---
mat_scaled <- t(scale(t(mat_annotated)))
colnames(mat_scaled) <- gsub("g", "D", colnames(mat_scaled))

# Clean up math errors and cap outliers
mat_scaled[is.na(mat_scaled)] <- 0
mat_scaled[mat_scaled > 2] <- 2
mat_scaled[mat_scaled < -2] <- -2

# --- 4. Plot Heatmap using ComplexHeatmap ---
plot_heat_ilc2s <- ComplexHeatmap::pheatmap(
  mat = mat_scaled, 
  scale = "none", 
  clustering_method = "ward.D2",
  color = colorRampPalette(c("#648FFF", "white", "#FFB000"))(101), 
  breaks = seq(-2, 2, length.out = 102), 
  display_numbers = round(mat_scaled, 2), 
  number_color = "black",
  treeheight_col = 10,
  treeheight_row = 20, 
  name = "Z-Score" 
)

# --- 5. Convert to ggplot object for arrangement ---
# Using grid.grabExpr(draw()) ensures that all the internal sizing, 
# legends, and dendrograms from ComplexHeatmap are captured perfectly.
gg_heat_ilc2s <- as.ggplot(grid::grid.grabExpr(ComplexHeatmap::draw(plot_heat_ilc2s, heatmap_legend_side = "bottom")))+
  ggtitle("ILC2s\nacross conditions\n")+
  theme(
        legend.title = element_text(size = 9, face = "bold"),
        legend.text = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
        plot.margin = margin(0.1, 0.25, 0, 0.25, "cm"),
        legend.position = "bottom"
      )

# 'gg_heat' is now a standard ggplot object! 
# You can now combine it using patchwork, e.g.:
# combined_plot <- spatial_scatter_plot + gg_heat
print(gg_heat_ilc2s)
```

```{r, fig.height=4, fig.width=3}
set.seed(8)

# --- 1. Subsetting ---
SO.sub <- subset(SO.lung, subset = AL3 %in% c("NK cells/ILC1s"))
SO.sub$AL3 <- droplevels(as.factor(SO.sub$AL3))

# --- 2. Direct Pseudo-bulk Aggregation ---
ilc_markers <- c("KLRG1", "CD127", "CD90",
                 "ICOS", "NKp46", "NK11",
                 "MHCII", "CD44", "Ki67")

# Calculate the mean directly from Seurat
avg_exp <- AverageExpression(SO.sub, features = ilc_markers, slot = "counts", group.by = "Treatment")
mat_annotated <- avg_exp[[DefaultAssay(SO.sub)]]
# --- 3. Manually Scale (Z-score) the Matrix ---
mat_scaled <- t(scale(t(mat_annotated)))
colnames(mat_scaled) <- gsub("g", "D", colnames(mat_scaled))

# Clean up math errors and cap outliers
mat_scaled[is.na(mat_scaled)] <- 0
mat_scaled[mat_scaled > 2] <- 2
mat_scaled[mat_scaled < -2] <- -2

# --- 4. Plot Heatmap using ComplexHeatmap ---
plot_heat_ilc1s <- ComplexHeatmap::pheatmap(
  mat = mat_scaled, 
  scale = "none", 
  clustering_method = "ward.D2",
  color = colorRampPalette(c("#648FFF", "white", "#FFB000"))(101), 
  breaks = seq(-2, 2, length.out = 102), 
  display_numbers = round(mat_scaled, 2), 
  number_color = "black",
  treeheight_col = 10,
  treeheight_row = 20, 
  name = "Z-Score" 
)

# --- 5. Convert to ggplot object for arrangement ---
# Using grid.grabExpr(draw()) ensures that all the internal sizing, 
# legends, and dendrograms from ComplexHeatmap are captured perfectly.
gg_heat_ilc1s <- as.ggplot(grid::grid.grabExpr(ComplexHeatmap::draw(plot_heat_ilc1s, heatmap_legend_side = "bottom")))+
  ggtitle("NK cells/ILC1s\nacross conditions\n")+
  theme(
        legend.title = element_text(size = 9, face = "bold"),
        legend.text = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
        plot.margin = margin(0.1, 0.25, 0, 0.25, "cm"),
        legend.position = "bottom"
      )

# 'gg_heat' is now a standard ggplot object! 
# You can now combine it using patchwork, e.g.:
# combined_plot <- spatial_scatter_plot + gg_heat
print(gg_heat_ilc1s)


# --- 1. Subsetting ---
SO.sub <- subset(SO.lung, subset = AL3 %in% c("ILC3s"))
SO.sub$AL3 <- droplevels(as.factor(SO.sub$AL3))

# --- 2. Direct Pseudo-bulk Aggregation ---
ilc_markers <- c("KLRG1", "CD127", "CD90", "CD3", "CD4",
                 "ICOS", "NKp46", "NK11",
                 "MHCII", "CD44", "Ki67")

# Calculate the mean directly from Seurat
avg_exp <- AverageExpression(SO.sub, features = ilc_markers, slot = "counts", group.by = "Treatment")
mat_annotated <- avg_exp[[DefaultAssay(SO.sub)]]
# --- 3. Manually Scale (Z-score) the Matrix ---
mat_scaled <- t(scale(t(mat_annotated)))
colnames(mat_scaled) <- gsub("g", "D", colnames(mat_scaled))

# Clean up math errors and cap outliers
mat_scaled[is.na(mat_scaled)] <- 0
mat_scaled[mat_scaled > 2] <- 2
mat_scaled[mat_scaled < -2] <- -2

# --- 4. Plot Heatmap using ComplexHeatmap ---
plot_heat_ilc3s <- ComplexHeatmap::pheatmap(
  mat = mat_scaled, 
  scale = "none", 
  clustering_method = "ward.D2",
  color = colorRampPalette(c("#648FFF", "white", "#FFB000"))(101), 
  breaks = seq(-2, 2, length.out = 102), 
  display_numbers = round(mat_scaled, 2), 
  number_color = "black",
  treeheight_col = 10,
  treeheight_row = 20, 
  name = "Z-Score" 
)

# --- 5. Convert to ggplot object for arrangement ---
# Using grid.grabExpr(draw()) ensures that all the internal sizing, 
# legends, and dendrograms from ComplexHeatmap are captured perfectly.
gg_heat_ilc3s <- as.ggplot(grid::grid.grabExpr(ComplexHeatmap::draw(plot_heat_ilc3s, heatmap_legend_side = "bottom")))+
  ggtitle("ILC3s\nacross conditions\n")+
  theme(
        legend.title = element_text(size = 9, face = "bold"),
        legend.text = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
        plot.margin = margin(0.1, 0.25, 0, 0.25, "cm"),
        legend.position = "bottom"
      )

# 'gg_heat' is now a standard ggplot object! 
# You can now combine it using patchwork, e.g.:
# combined_plot <- spatial_scatter_plot + gg_heat
print(gg_heat_ilc3s)


# --- 1. Subsetting ---
SO.sub <- subset(SO.lung, subset = AL3 %in% c("ILC3s"))
SO.sub$AL3 <- droplevels(as.factor(SO.sub$AL3))

# --- 2. Direct Pseudo-bulk Aggregation ---
ilc_markers <- c("ICOS", "CD11c", "CD68", "SiglecF", "CD4",
                 "MHCII", "CD44", "Ki67")

# Calculate the mean directly from Seurat
avg_exp <- AverageExpression(SO.sub, features = ilc_markers, slot = "counts", group.by = "Treatment")
mat_annotated <- avg_exp[[DefaultAssay(SO.sub)]]
# --- 3. Manually Scale (Z-score) the Matrix ---
mat_scaled <- t(scale(t(mat_annotated)))
colnames(mat_scaled) <- gsub("g", "D", colnames(mat_scaled))

# Clean up math errors and cap outliers
mat_scaled[is.na(mat_scaled)] <- 0
mat_scaled[mat_scaled > 2] <- 2
mat_scaled[mat_scaled < -2] <- -2

# --- 4. Plot Heatmap using ComplexHeatmap ---
plot_heat_my <- ComplexHeatmap::pheatmap(
  mat = mat_scaled, 
  scale = "none", 
  clustering_method = "ward.D2",
  color = colorRampPalette(c("#648FFF", "white", "#FFB000"))(101), 
  breaks = seq(-2, 2, length.out = 102), 
  display_numbers = round(mat_scaled, 2), 
  number_color = "black",
  treeheight_col = 10,
  treeheight_row = 20, 
  name = "Z-Score" 
)

# --- 5. Convert to ggplot object for arrangement ---
# Using grid.grabExpr(draw()) ensures that all the internal sizing, 
# legends, and dendrograms from ComplexHeatmap are captured perfectly.
gg_heat_my <- as.ggplot(grid::grid.grabExpr(ComplexHeatmap::draw(plot_heat_my, heatmap_legend_side = "bottom")))+
  ggtitle("Myeloid cells\nacross conditions\n")+
  theme(
        legend.title = element_text(size = 9, face = "bold"),
        legend.text = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
        plot.margin = margin(0.1, 0.25, 0, 0.25, "cm"),
        legend.position = "bottom"
      )

# 'gg_heat' is now a standard ggplot object! 
# You can now combine it using patchwork, e.g.:
# combined_plot <- spatial_scatter_plot + gg_heat
print(gg_heat_my)
```

```{r, fig.height=4, fig.width=9}
set.seed(8)

# --- 1. Subsetting ---
SO.sub <- subset(SO.lung, subset = AL3 %in% c("NK cells/ILC1s", "ILC2s", "ILC3s"))
SO.sub$AL3 <- droplevels(as.factor(SO.sub$AL3))

SO.sub <- subset(SO.sub, subset = Treatment %in% c("CTRL"))
SO.sub$Treatment <- droplevels(as.factor(SO.sub$Treatment))

# --- 2. Direct Pseudo-bulk Aggregation ---
ilc_markers <- c("KLRG1",   
                 "ICOS", 
                 "MHCII", "CD44", "Ki67", "NKp46")

# Calculate the mean directly from Seurat
avg_exp <- AverageExpression(SO.sub, features = ilc_markers, slot = "counts", group.by = "AL3")
mat_annotated <- avg_exp[[DefaultAssay(SO.sub)]]
# --- 3. Manually Scale (Z-score) the Matrix ---
mat_scaled <- t(scale(t(mat_annotated)))
colnames(mat_scaled) <- gsub("g", "D", colnames(mat_scaled))

# Clean up math errors and cap outliers
mat_scaled[is.na(mat_scaled)] <- 0
mat_scaled[mat_scaled > 2] <- 2
mat_scaled[mat_scaled < -2] <- -2

# --- 4. Plot Heatmap using ComplexHeatmap ---
plot_heat_ctrl <- ComplexHeatmap::pheatmap(
  mat = mat_scaled, 
  scale = "none", 
  clustering_method = "ward.D2",
  color = colorRampPalette(c("#648FFF", "white", "#FFB000"))(101), 
  breaks = seq(-2, 2, length.out = 102), 
  display_numbers = round(mat_scaled, 2), 
  number_color = "black",
  treeheight_col = 10,
  treeheight_row = 20, 
  name = "Z-Score" 
)

# --- 5. Convert to ggplot object for arrangement ---
# Using grid.grabExpr(draw()) ensures that all the internal sizing, 
# legends, and dendrograms from ComplexHeatmap are captured perfectly.
gg_heat_ctrl <- as.ggplot(grid::grid.grabExpr(ComplexHeatmap::draw(plot_heat_ctrl, heatmap_legend_side = "right")))+
  ggtitle("ILC subtypes\n@ CTRL\n")+
  theme(
        legend.title = element_text(size = 9, face = "bold"),
        legend.text = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
        plot.margin = margin(0.1, 0.5, 0, 0.5, "cm"),
        legend.position = "bottom"
      )

# 'gg_heat' is now a standard ggplot object! 
# You can now combine it using patchwork, e.g.:
# combined_plot <- spatial_scatter_plot + gg_heat


# --- 1. Subsetting ---
SO.sub <- subset(SO.lung, subset = AL3 %in% c("NK cells/ILC1s", "ILC2s", "ILC3s"))
SO.sub$AL3 <- droplevels(as.factor(SO.sub$AL3))

SO.sub <- subset(SO.sub, subset = Treatment %in% c("3"))
SO.sub$Treatment <- droplevels(as.factor(SO.sub$Treatment))

# Calculate the mean directly from Seurat
avg_exp <- AverageExpression(SO.sub, features = ilc_markers, slot = "counts", group.by = "AL3")
mat_annotated <- avg_exp[[DefaultAssay(SO.sub)]]
# --- 3. Manually Scale (Z-score) the Matrix ---
mat_scaled <- t(scale(t(mat_annotated)))
colnames(mat_scaled) <- gsub("g", "D", colnames(mat_scaled))

# Clean up math errors and cap outliers
mat_scaled[is.na(mat_scaled)] <- 0
mat_scaled[mat_scaled > 2] <- 2
mat_scaled[mat_scaled < -2] <- -2

# --- 4. Plot Heatmap using ComplexHeatmap ---
plot_heat_d3 <- ComplexHeatmap::pheatmap(
  mat = mat_scaled, 
  scale = "none", 
  clustering_method = "ward.D2",
  color = colorRampPalette(c("#648FFF", "white", "#FFB000"))(101), 
  breaks = seq(-2, 2, length.out = 102), 
  display_numbers = round(mat_scaled, 2), 
  number_color = "black",
  treeheight_col = 10,
  treeheight_row = 20, 
  name = "Z-Score" 
)

# --- 5. Convert to ggplot object for arrangement ---
# Using grid.grabExpr(draw()) ensures that all the internal sizing, 
# legends, and dendrograms from ComplexHeatmap are captured perfectly.
gg_heat_d3 <- as.ggplot(grid::grid.grabExpr(ComplexHeatmap::draw(plot_heat_d3, heatmap_legend_side = "right")))+
  ggtitle("ILC subtypes\n@ D3\n")+
  theme(
        legend.title = element_text(size = 9, face = "bold"),
        legend.text = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), 
        plot.margin = margin(0.1, 0.5, 0, 0.5, "cm"),
        legend.position = "bottom"
      )

# 'gg_heat' is now a standard ggplot object! 
# You can now combine it using patchwork, e.g.:
# combined_plot <- spatial_scatter_plot + gg_heat

ggarrange(gg_heat_ctrl, gg_heat_d3, ncol = 2)
```

## MHCII expression by ILC2s

```{r, fig.height=5, fig.width=7}
library(Seurat)
library(ggplot2)
library(dplyr)
library(rstatix)
library(ggpubr)

# --- 1. Subset for ILC2s ---
SO.ilc2 <- subset(SO.lung, subset = AL3 == "ILC2s")
SO.ilc2$Treatment <- paste0("D", SO.ilc2$Treatment)
SO.ilc2$Treatment <- gsub("DCTRL", "CTRL", SO.ilc2$Treatment)
SO.ilc2$Treatment <- factor(SO.ilc2$Treatment, levels = c("CTRL", "D1", "D2", "D3"))
MHCII_threshold <- 2.3

# --- 2. Ridge Plot with Threshold Line ---
# Fetch MHCII expression and metadata
plot_data <- FetchData(SO.ilc2, vars = c("MHCII", "Treatment"))

ggplot(plot_data, aes(x = MHCII, y = Treatment, fill = Treatment)) +
  ggridges::geom_density_ridges(scale = 1.2, alpha = 0.7) +
  geom_vline(xintercept = MHCII_threshold, linetype = "dashed", color = "red", linewidth = 1) +
  scale_fill_manual(values = cols_treat) +
  theme_classic() +
  labs(title = "MHCII Expression in ILC2s",
       subtitle = paste0("Red line indicates threshold (> ", MHCII_threshold, ")"),
       x = "Expression Level") +
  theme(legend.position = "none")
```

```{r}

```

## Combine plots for figure

```{r, fig.width=9, fig.height=9}
top_figure <- ggarrange(plot_prop, plot_freq_immune, plot_freq, 
          ncol = 3, nrow = 1, 
          labels = c("AUTO"), 
          label.x = 0.1)+
  theme(plot.margin = margin(0, 0, 0.2, 0, "cm"))


middle_figure <- ggarrange(
          plot_count_all, plot_count_immune, plot_count_ilc,
          plot_count_ilc2, 
          ncol = 2, nrow = 2, 
          labels = c("D", "E", "F",
                     "G"),
          label.x = 0.1)+
  theme(plot.margin = margin(0, 0, 0, 0, "cm"))



final_figure <- ggarrange(top_figure, middle_figure, 
          ncol = 1, nrow = 2, heights = c(3, 4.5))+
  theme(plot.margin = margin(0, 0.1, 0, 0, "cm"))

final_figure

annotate_figure(final_figure,
               bottom = text_grob("N: CTRL/D1/D2 = 9 FOVs, D3 = 8 FOVs; Kruskal-Wallis-test & Dunn’s test; Significance levels of adjusted p-values: **** = 1e-04, *** = 0.001, ** = 0.01, * = 0.05", 
                                  color = "black", 
                                  # hjust = -1,
                                  face = "italic", size = 8))
```

```{r, fig.width=9, fig.height=9}
top_figure <- ggarrange(plot_prop, plot_freq_immune, plot_freq, 
          ncol = 3, nrow = 1, 
          labels = c("AUTO"), 
          label.x = 0.1)+
  theme(plot.margin = margin(0, 0, 0.2, 0, "cm"))


middle_figure_right <- ggarrange(
          plot_count_all, plot_count_immune, plot_count_ilc,
          plot_count_ilc2, 
          ncol = 2, nrow = 2, 
          labels = c("D", "E", "F",
                     "G"),
          label.x = 0.1)+
  theme(plot.margin = margin(0, 0, 0, 0, "cm"))

middle_figure <- ggarrange(
          middle_figure_right,
          gg_heat_ilc2s, 
          ncol = 2, nrow = 1, 
          widths = c(2, 1),
          labels = c("", "H"),
          label.x = 0.1)+
  theme(plot.margin = margin(0, 0, 0, 0, "cm"))


final_figure <- ggarrange(top_figure, middle_figure, 
          ncol = 1, nrow = 2, heights = c(3, 4.5))+
  theme(plot.margin = margin(0, 0.1, 0, 0, "cm"))

final_figure

annotate_figure(final_figure,
               bottom = text_grob("N: CTRL/D1/D2 = 9 FOVs, D3 = 8 FOVs; Kruskal-Wallis-test & Dunn’s test; Significance levels of adjusted p-values: **** = 1e-04, *** = 0.001, ** = 0.01, * = 0.05", 
                                  color = "black", 
                                  # hjust = -1,
                                  face = "italic", size = 8))
```

```{r, fig.width=9, fig.height=7}
suppl_figure <- ggarrange(plot_count_ilc1, plot_count_ilc3, plot_freq_immune_d3, plot_freq_d3,
          ncol = 2, nrow = 2, 
          heights = c(3, 4),
          labels = c("AUTO"),
          label.x = 0.1)+
  theme(plot.margin = margin(0, 0, 0, 0, "cm"))

suppl_figure

annotate_figure(suppl_figure,
               bottom = text_grob("N: CTRL/D1/D2 = 9 FOVs, D3 = 8 FOVs; Kruskal-Wallis-test & Dunn’s test; Significance levels of adjusted p-values: **** = 1e-04, *** = 0.001, ** = 0.01, * = 0.05",
                                  color = "black",
                                  # hjust = -1,
                                  face = "italic", size = 8))
```

```{r, fig.width=9, fig.height=3}
ggarrange(plot_freq_ilc1, plot_freq_ilc2, plot_freq_ilc3, 
          ncol = 3, nrow = 1, 
          labels = c("A", "B", "C", "D", "E", "F", "G", "H", "I"))+
  theme(plot.margin = margin(0, 0.1, 0, 0, "cm"))
```

## Session Information

```{r}
save.image(paste0(output_dir, "/environment.RData"))
sessionInfo()
```
