---
title: "Figure 8: Quantification and spatial analysis of ILC subtypes in the SI"
author: "Sandy Kroh"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    number_sections: yes
    fig_caption: yes
    keep_md: yes
  pdf_document:
    toc: yes
params:
  Sample: Default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", 
                      fig.align='center', dpi = 600, 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Libraries

```{r}
# remove.packages("rlang")
# remove.packages("dplyr")
# install.packages("rlang")
# install.packages("dplyr")

library(SeuratObject)
library(dplyr)
library(rstatix)
library(rlang)

if (!requireNamespace("Giotto", quietly = TRUE))
  devtools::install_github("drieslab/Giotto@suite")
if (!requireNamespace("VoltRon", quietly = TRUE))
  devtools::install_github("Artur-man/VoltRon")
if (!requireNamespace("Seurat", quietly = TRUE))
  install.packages("Seurat")
library(Giotto)
library(Seurat)
library(VoltRon)
library(ggplot2)
library(ggpubr)
library(readr)
library(ggbeeswarm)
library(stringr)
```

## Parameters

```{r}
set.seed(123)

input_dir <- here::here("1_data_tidying", "Lung_SI_all_cells_all_ALs_files")

output_dir <- here::here("2_visualizations_for_figures", "Fig_8_quantification_spatial_analysis_SI_files")
dir.create(output_dir)




main_markers <- c(
  "EpCAM", "EMCN", "LYVE1", "PDPN", "PDGFRa", "CD8a", "CD4",
  "CD45", "CD3", "IRF4", "Kappa", "CD11c", "CD127", "GATA3eGFP", "RORgt"
)


immune_markers <- c(
 "CD3", "CD4", "CD8a", "Kappa", "IRF4", "CD11c",
  "CD127", "CD90", "EOMES", "GATA3eGFP", "RORgt", "Ki67",  "KLRG1", "NKp46", "CD117", "Areg", "CCR6", "CD44", "MHCII", "Sca1"
)

ilc_markers <- c(
  "CD3", "CD4", "CD8a",
  "CD127", "CD90", "EOMES", "GATA3eGFP", "RORgt", "KLRG1", "NKp46", "CD117", "CCR6", "MHCII", "Ki67", "Areg", "IRF4", "Sca1", "CD44"
)


cols_nat <- c("magenta", "cyan", "blue", "purple", "green", 
                       "red", "yellow", "olivedrab1", "slateblue1", 
                       "darkcyan", "gold","indianred1", "seagreen", "deeppink", 
                       "orange", "brown", "violet",
                       "deeppink4", "pink", 
                       "grey", "black", "lightgreen", 
                       "#FF0066",  
                       "lightblue", "#FFCC99", "#CC00FF", 
                       "blueviolet",  "goldenrod4", 
                       "navy", "olivedrab", "lightcyan", "seagreen2", "darkviolet", "lightpink", "slateblue4", "olivedrab2")

colfunc <- colorRampPalette(c("darkcyan", "green", "yellow", "magenta", "purple"))

cols_ilcs_lung <- c("darkcyan", "seagreen2", "deeppink4")
cols_ilcs_si <- c("slateblue", "seagreen2")

cols_treat <- c("darkcyan", "gold", "slateblue")

cols_organs <- c("indianred2", "seagreen", "navy")
cols_organs_si <- c("seagreen", "navy")

```

# Load data

## Proportions

```{r, fig.width=9, fig.height=6}

df_villi <- read_csv(paste0(input_dir, "/si_villi_proportions.csv"), 
    col_types = cols(...1 = col_skip()))

df_ilf <- read_csv(paste0(input_dir, "/si_ilf_proportions.csv"), 
    col_types = cols(...1 = col_skip()))

# the FOV CTRL_FOV1_20210810 is falsely labeled as villi but is actually an ILF, so it must be changed in the metadata
df_correct <- df_villi %>%
  filter(Dataset == "CTRL_FOV1_20210810") %>%
  mutate(Tissue.area = "ILF")

df_ilf <- rbind(df_ilf, df_correct)

df_villi <- df_villi %>%
  filter(Dataset != "CTRL_FOV1_20210810")


df_lung <- read_csv(paste0(input_dir, "/lung_proportions.csv"), 
    col_types = cols(...1 = col_skip()))

```

## Spatial data for VoltRon and Giotto

```{r, fig.width=9, fig.height=6}

# from import_Giotto.Rmd
gio_list <- readRDS(here::here("data", "Giotto_data_Villi.rds"))

# from import_VoltRon.Rmd
vr_list <- readRDS(here::here("data", "VoltRon_data_Villi.rds"))

# original data
metadatax <- read_csv(here::here("data", "SO_arcsinh_si_imputed_Villi.csv"))

metadatax <- metadatax %>% 
  filter(`Tissue area` == "Villi")

unique(metadatax$CellType)

vr_list_names <- unique(metadatax$Dataset)


cell_proximities_list <- list()
for(samp in vr_list_names){
  print(samp)
  cell_proximities_list[[samp]] <-cellProximityEnrichment(
    gobject = gio_list[[samp]],
    cluster_column = 'CellType',
    spatial_network_name = 'Delaunay_network',
    adjust_method = 'fdr',
    number_of_simulations = 1000)
  cell_proximities_list[[samp]] <- cell_proximities_list[[samp]]$enrichm_res
}

vr_merged <- merge(vr_list[[1]], vr_list[-1])
vrImageNames(vr_merged)
unique(vr_merged$CellType)

```

# Visualization

Define theme for plotting of boxplots and coenrichment plots:

```{r}

# BOXPLOTS --------------------------------------------------------------------------------
theme_boxplots <-   function() {
  theme(plot.margin=margin(0,0.5,1,0.8,"cm"),
        axis.text.x = element_text(#angle = 45, 
                                   vjust = 1, size = 10, hjust = 0.5, face = "bold"),
        axis.text.y = element_text(hjust = 0.5, size = 10),
        axis.title.x = element_text(size = 11),
        axis.title.y = element_text(size = 10),
        plot.title = element_text(size =12, hjust = 0.5),
        legend.title = element_text(size =10),
        legend.text = element_text(size =10)
  )
  }

annot_text_size <- 8/.pt

# COENRICHMENT PLOTS ----------------------------------------------------------------------
theme_coenrich <-   function() {
  theme(axis.text.x = element_text(#angle = 50,
                                   vjust = 1, size = 10, hjust = 0.5, face = "bold"
                                   ),
        axis.text.y = element_text(hjust = 0.5, size = 10),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        plot.title = element_text(size =12, hjust = 0.5),
        plot.margin = margin(0, 1, 0.5, 1, "cm"),
        legend.position = "none",
        strip.background=element_blank(),
        strip.background.x= element_blank(),
        strip.text.x = element_text(size = 1, color = "white"),
        panel.grid.major.y = element_line()
        )
}

```

## Fig. 8A-C - Total cell, immune cell, and ILC count in SI villi at conditions

```{r, fig.width=3, fig.height=6}

# filter for CTRL and convert to longer format
df <- df_villi %>%
  select(Treatment, Dataset, TotalCellCountFOV, `Immune cells`, ILCs, ILC2s, `NK cells/ILC1s/ILC3s`) %>%
  mutate(Treatment = factor(Treatment, level =c(
    "CTRL", "1", "3"
  )))

# TOTAL CELL COUNT -------------------------------------------------------------------
# Testing for normal distribution
shapiro.test(df$TotalCellCountFOV)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`TotalCellCountFOV` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`TotalCellCountFOV` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`TotalCellCountFOV` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_count_all <- ggplot(df, aes(x = Treatment, y = TotalCellCountFOV, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme_boxplots()+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 2500) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("Total cells")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,3000))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=2900,
           label=Labels,
           #angle = 90,
           size = annot_text_size)




# IMMUNE -------------------------------------------------------------------
# Testing for normal distribution
shapiro.test(df$`Immune cells`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`Immune cells` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`Immune cells` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`Immune cells` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_count_immune <- ggplot(df, aes(x = Treatment, y = `Immune cells`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme_boxplots()+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 1500) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("Immune cells")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,1800))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=1700,
           label=Labels,
           #angle = 90,
           size = annot_text_size)




# ILCs -------------------------------------------------------------------
# Testing for normal distribution
shapiro.test(df$ILCs)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`ILCs` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`ILCs` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`ILCs` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_count_ilc <- ggplot(df, aes(x = Treatment, y = ILCs, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme_boxplots()+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 250) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("ILCs")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,290))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=280,
           label=Labels,
           #angle = 90,
           size = annot_text_size)

# ILC2s -------------------------------------------------------------------
# Testing for normal distribution
shapiro.test(df$ILC2s)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`ILC2s` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`ILC2s` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`ILC2s` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

ggplot(df, aes(x = Treatment, y = ILC2s, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme_boxplots()+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 250) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("ILC2s")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,290))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=280,
           label=Labels,
           #angle = 90,
           size = annot_text_size)

# Nk cells/ILC1s/ILC3s -------------------------------------------------------
# Testing for normal distribution
shapiro.test(df$`NK cells/ILC1s/ILC3s`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`NK cells/ILC1s/ILC3s` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`NK cells/ILC1s/ILC3s` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`NK cells/ILC1s/ILC3s` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

ggplot(df, aes(x = Treatment, y = `NK cells/ILC1s/ILC3s`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme_boxplots()+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 250) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("NK cells/ILC1s/ILC3s")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,290))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=280,
           label=Labels,
           #angle = 90,
           size = annot_text_size)


plot_ac <- ggarrange(plot_count_all, plot_count_immune, plot_count_ilc, ncol = 1, nrow = 3, labels = "AUTO")

plot_ac
```

## Fig. 8D-E - Immune and ILC frequency in SI villi

```{r, fig.width=3, fig.height=5.5}

# filter for CTRL and convert to longer format
df <- df_villi %>%
  select(Treatment, Dataset, TotalCellCountFOV, `Immune cells`, ILCs) %>%
  mutate(Prop_immune = `Immune cells`/TotalCellCountFOV*100, 
         Prop_ilc = ILCs/`Immune cells`*100,
         Treatment = factor(Treatment, level =c(
           "CTRL", "1", "3"
  )))

# TOTAL CELL COUNT -------------------------------------------------------------------
# Testing for normal distribution
shapiro.test(df$Prop_immune)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`Prop_immune` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`Prop_immune` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`Prop_immune` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_freq_immune <- ggplot(df, aes(x = Treatment, y = Prop_immune, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme_boxplots()+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 85) +
  xlab(NULL)+
  ylab("Frequency cells/total cells FOV [%]")+
  ggtitle("Immune cells")+
  scale_y_continuous(expand = c(0, 0), limits = c(40,100))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=95,
           label=Labels,
           #angle = 90,
           size = annot_text_size)




# IMMUNE -------------------------------------------------------------------
# Testing for normal distribution
shapiro.test(df$`Prop_ilc`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`Prop_ilc` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`Prop_ilc` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`Prop_ilc` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_freq_ilc <- ggplot(df, aes(x = Treatment, y = `Prop_ilc`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme_boxplots()+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 35) +
  xlab(NULL)+
  ylab("Frequency cells/immune cells FOV [%]")+
  ggtitle("ILCs")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,43))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=40,
           label=Labels,
           #angle = 90,
           size = annot_text_size)


plot_de <- ggarrange(plot_freq_immune, plot_freq_ilc, ncol = 1, nrow = 2, labels = c("D", "E"))
plot_de

```

## Fig. 8F-G - Frequency of ILC subtypes in SI villi at different conditions

```{r, fig.width=3, fig.height=5.5}

# filter for CTRL and convert to longer format
df <- df_villi %>%
  select(Treatment, Dataset, `Immune cells`, `NK cells/ILC1s/ILC3s`, ILC2s) %>%
  mutate(Prop_nk =  `NK cells/ILC1s/ILC3s`/`Immune cells`*100, 
         Prop_ilc = ILC2s/`Immune cells`*100,
         Treatment = factor(Treatment, level =c(
           "CTRL", "1", "3"
  )))

# TOTAL CELL COUNT -------------------------------------------------------------------
# Testing for normal distribution
shapiro.test(df$Prop_nk)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`Prop_nk` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`Prop_nk` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`Prop_nk` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_freq_immune <- ggplot(df, aes(x = Treatment, y = Prop_nk, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme_boxplots()+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 17) +
  xlab(NULL)+
  ylab("Frequency cells/immune cells FOV [%]")+
  ggtitle("NK cells/ILC1s/ILC3s")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,23))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=22,
           label=Labels,
           #angle = 90,
           size = annot_text_size)




# IMMUNE -------------------------------------------------------------------
# Testing for normal distribution
shapiro.test(df$`Prop_ilc`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`Prop_ilc` ~ Treatment)
res.kruskal
df %>% kruskal_effsize(`Prop_ilc` ~ Treatment)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`Prop_ilc` ~ Treatment, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Treatment, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Treatment")

plot_freq_ilc <- ggplot(df, aes(x = Treatment, y = `Prop_ilc`, fill = "Treatment"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Treatment), size = 2, cex = 3)+
  scale_color_manual(values = cols_treat)+
  theme_classic2()+
  theme_boxplots()+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 15) +
  xlab(NULL)+
  ylab("Frequency cells/immune cells FOV [%]")+
  ggtitle("ILC2s")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,23))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=22,
           label=Labels,
           #angle = 90,
           size = annot_text_size)


plot_fg <- ggarrange(plot_freq_immune, plot_freq_ilc, ncol = 1, nrow = 2, labels = c("F", "G"))
plot_fg

```

```{r, fig.width=9, fig.height=5.5}
plot_dg <- ggarrange(plot_de, plot_fg, ncol = 2, nrow = 1)

plot_ag <- ggarrange(plot_ac, plot_dg, ncol = 2, nrow = 1, widths = c(1, 2))

plot_ag
```

## Fig. 8H - Co-enrichment ILC2s localize with ILC2s and myeloid cells in the SI villi

```{r}

# fine tune the co-enrichment plot
# ILC2s around ILC2s -----------------------------------------
interactions <- c("ILC2s--ILC2s")
unic <- "ILC2s"
interaction_celltypes <- NULL
for(samp in vr_list_names){
    cur_cell_proximities <- cell_proximities_list[[samp]]
    cur_cell_proximities <- cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,]
    sample <- unique(metadatax$Dataset[metadatax$Dataset==samp])
    if(nrow(cur_cell_proximities) > 0 & sample != "20210906_FOV3_D3"){
      interaction_celltypes <- rbind(interaction_celltypes,
                                     data.frame(cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,], 
                                                experiment = strsplit(sample, split = "_")[[1]][1], fov = strsplit(sample, split = "_")[[1]][2], condition = strsplit(sample, split = "_")[[1]][3]))
    }
  }
interaction_celltypes$p.adj <- ifelse(interaction_celltypes$enrichm > 0, interaction_celltypes$p.adj_higher, interaction_celltypes$p.adj_lower)
  # plot test results
  # sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("p=",round(interaction_celltypes$p.adj,3)), ""))
  sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("*"), ""))
    # print(sig_label)
plot_coenrichment_ilc2s <- ggplot(interaction_celltypes, aes(x = experiment, y = enrichm, fill = experiment)) +
  geom_bar(stat = "identity", position = position_dodge2(width=0.9, preserve = "single")) +
  facet_grid(.~experiment, scales = "free_x") +
  geom_text(aes(label=sig_label), position=position_dodge2(width=0.9, preserve = "single"), angle = 90, hjust = -0.02, size = 4) +
  ylim(-0.4,2.5)+
  NoLegend()+
  theme_classic2()+
  scale_fill_manual(values = cols_treat, name = "") +
  ggtitle(gsub("LYVE1 CD90 ", "", interactions)) +
  theme_coenrich()+
  NoLegend()+  
  ylab("Enrichment")

plot_coenrichment_ilc2s
```

```{r}

# fine tune the co-enrichment plot
# ILC2s around Myeloid cells -----------------------------------------
interactions <- c("ILC2s--Myeloid cells")
unic <- "ILC2s"
interaction_celltypes <- NULL
for(samp in vr_list_names){
    cur_cell_proximities <- cell_proximities_list[[samp]]
    cur_cell_proximities <- cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,]
    sample <- unique(metadatax$Dataset[metadatax$Dataset==samp])
    if(nrow(cur_cell_proximities) > 0 & sample != "20210906_FOV3_D3"){
      interaction_celltypes <- rbind(interaction_celltypes,
                                     data.frame(cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,], 
                                                experiment = strsplit(sample, split = "_")[[1]][1], fov = strsplit(sample, split = "_")[[1]][2], condition = strsplit(sample, split = "_")[[1]][3]))
    }
  }
interaction_celltypes$p.adj <- ifelse(interaction_celltypes$enrichm > 0, interaction_celltypes$p.adj_higher, interaction_celltypes$p.adj_lower)
  # plot test results
  # sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("p=",round(interaction_celltypes$p.adj,3)), ""))
  sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("*"), ""))
    # print(sig_label)
plot_coenrichment_my <- ggplot(interaction_celltypes, aes(x = experiment, y = enrichm, fill = experiment)) +
  geom_bar(stat = "identity", position = position_dodge2(width=0.9, preserve = "single")) +
  facet_grid(.~experiment, scales = "free_x") +
  geom_text(aes(label=sig_label), position=position_dodge2(width=0.9, preserve = "single"), angle = 90, hjust = -0.02, size = 4) +
  ylim(-0.1,1.2)+
  NoLegend()+
  theme_classic2()+
  scale_fill_manual(values = cols_treat, name = "") +
  ggtitle(gsub("LYVE1 CD90 ", "", interactions)) +
  theme_coenrich()+
  NoLegend()+  
  ylab("Enrichment")

plot_coenrichment_my
```

## Fig. 8J - IF overlays ILC2s localize with ILC2s and myeloid cells in the SI villi

IF overlay with ILC2s and myeloid cells on top:

```{r eval=FALSE, fig.height=4.5, fig.width=4.5, include=FALSE}

set_ptsize <- 1
cell_shape <- 5
set_alpha <- 1

# define cell type of interested that should be plotted on the overlay
celltype_of_interest <- c("")

ColorsCellTypeSingle <-  list(
  `Myeloid cells` = "green",
  `ILC2s` = "orange")


# overlay 1 ------------------------------------------------------------
# define markers for the Overlay
# CYAN
marker1 <- "EpCAM"
# MAGENTA
marker2 <- "Sytox"
# YELLOW
marker3 <- "CD45"

name_channel_key <- paste0(marker1, "-c_", marker2, "-m_", marker3, "-y_")
vr_merged <- combineChannels(vr_merged,
                             channels = c(marker1, marker2, marker3),
                             colors = c("cyan", "magenta", "yellow"),
                             channel_key = name_channel_key)


plot <- vrSpatialPlot(vr_merged, assay = paste0("Assay", 18), #2
                        group.by = "CellType", 
                        group.ids = celltype_of_interest,
                        alpha = set_alpha, 
                        background = c("image_1", name_channel_key), 
                        pt.size = set_ptsize, cell.shape = cell_shape)+
  guides(color = guide_legend(override.aes = list(size = 5)))+
  scale_color_manual(values = ColorsCellTypeSingle)+
  scale_fill_manual(values = ColorsCellTypeSingle)+
  theme_void()+ NoLegend()+ ggtitle(NULL)+
  theme(plot.title = element_blank(), 
        text = element_text(size = 12))



plot_if_1 <- plot +
  annotate("text", x=200, y=900, label= marker1,
           col="cyan", size=5, parse=TRUE) +
  annotate("text", x=520, y=900, label= marker2,
           col="magenta", size=5, parse=TRUE) +
  annotate("text", x=800, y=900, label= marker3,
           col="yellow", size=5, parse=TRUE)+
  annotate("segment", x = 680, xend = 985, y = 45, yend = 45, size = 1.6, 
  colour = "white")

plot_if_1
```

```{r, fig.width=4.5, fig.height=4.5}

set_ptsize <- 1.5
cell_shape <- 18
set_alpha <- 0.8

# define cell type of interested that should be plotted on the overlay
celltype_of_interest <- c("Myeloid cells", "ILC2s")

ColorsCellTypeSingle <-  list(
  `Myeloid cells` = "magenta",
  `ILC2s` = "yellow")


# overlay 1 ------------------------------------------------------------
# define markers for the Overlay
# CYAN
marker1 <- "EpCAM"
# MAGENTA
marker2 <- "Sytox"
# YELLOW
marker3 <- "DAPI"

name_channel_key <- paste0(marker1, "-c_", marker2, "-m_", marker3, "-y_")
vr_merged <- combineChannels(vr_merged,
                             channels = c(marker1, marker2, marker3),
                             colors = c("white", "magenta", "yellow"),
                             channel_key = name_channel_key)


plot <- vrSpatialPlot(vr_merged, assay = paste0("Assay", 19), #2
                        group.by = "CellType", 
                        group.ids = celltype_of_interest,
                        alpha = set_alpha, 
                        background = c("image_1", name_channel_key), 
                        pt.size = set_ptsize, cell.shape = cell_shape)+
  guides(color = guide_legend(override.aes = list(size = 5)))+
  scale_color_manual(values = ColorsCellTypeSingle)+
  scale_fill_manual(values = ColorsCellTypeSingle)+
  theme_void()+ NoLegend()+ ggtitle(NULL)+
  theme(plot.title = element_blank(), 
        text = element_text(size = 12))



plot_if_2 <- plot +
  annotate("text", x=150, y=950, label= marker1,
           col="white", size=5, parse=TRUE) +
  annotate("segment", x = 680, xend = 985, y = 45, yend = 45, size = 1.6, 
  colour = "white")

plot_if_2
```

Combine subplots:

```{r, fig.width=9, fig.height=3.5}

plot_coenrichments <- ggarrange(plot_coenrichment_my, plot_coenrichment_ilc2s, ncol = 1, nrow = 2, labels = c("H"))+
  theme(plot.margin = margin(0, 0.5, 0, 0, "cm"))

plot_si_ilc2s <- ggarrange(plot_coenrichments, plot_if_2, ncol = 2, nrow = 1, widths = c(1.5, 1), labels = c("", "I"), label.x = -0.05)

plot_si_ilc2s
```

## Fig. 8K - Co-enrichment of NK cells/ILC1s/ILC3s with epithelia II cells in SI villi

```{r}
# fine tune the co-enrichment plot
# ILC2s around ILC2s -----------------------------------------
interactions <- c("Epithelia II--NK cells/ILC1s/ILC3s")
unic <- "Epithelia II"
interaction_celltypes <- NULL
for(samp in vr_list_names){
    cur_cell_proximities <- cell_proximities_list[[samp]]
    cur_cell_proximities <- cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,]
    sample <- unique(metadatax$Dataset[metadatax$Dataset==samp])
    if(nrow(cur_cell_proximities) > 0 & sample != "20210906_FOV3_D3"){
      interaction_celltypes <- rbind(interaction_celltypes,
                                     data.frame(cur_cell_proximities[cur_cell_proximities$unified_int %in% interactions,], 
                                                experiment = strsplit(sample, split = "_")[[1]][1], fov = strsplit(sample, split = "_")[[1]][2], condition = strsplit(sample, split = "_")[[1]][3]))
    }
  }
interaction_celltypes$p.adj <- ifelse(interaction_celltypes$enrichm > 0, interaction_celltypes$p.adj_higher, interaction_celltypes$p.adj_lower)
  # plot test results
  # sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("p=",round(interaction_celltypes$p.adj,3)), ""))
  sig_label <- as.character(ifelse(interaction_celltypes$p.adj < 0.1, paste0("*"), ""))
    # print(sig_label)
plot_coenrichment <- ggplot(interaction_celltypes, aes(x = experiment, y = enrichm, fill = experiment)) +
  geom_bar(stat = "identity", position = position_dodge2(width=0.9, preserve = "single")) +
  facet_grid(.~experiment, scales = "free_x") +
  geom_text(aes(label=sig_label), position=position_dodge2(width=0.9, preserve = "single"), angle = 90, hjust = -0.02, size = 4) +
  ylim(-1.5,1)+
  NoLegend()+
  theme_classic2()+
  scale_fill_manual(values = cols_treat, name = "") +
  ggtitle(gsub("LYVE1 CD90 ", "", interactions)) +
  theme_coenrich()+
  NoLegend()+  
  ylab("Enrichment")

plot_coenrichment
```

## Fig. 8L - IF example image of re-distribution of NK cells/ILC1s/ILC3s after IL-33 application

```{r, fig.width=4.5, fig.height=4.5}

ColorsCellType <-  list(
  #`NK cells/ILC1s/ILC3s` = "cyan", 
  `NK cells/ILC1s/ILC3s` = "magenta",
  #`ILC3s` = "magenta", 
  `Epithelia II` = "green")


cell_type_of_interest <- c("NK cells/ILC1s/ILC3s", "Epithelia II")

set_ptsize <- 1
cell_shape <- 18
set_alpha <- 1

plot <- vrSpatialPlot(vr_merged, assay = paste0("Assay", 6), 
                        group.by = "CellType", 
                        group.ids = cell_type_of_interest,
                        alpha = set_alpha, 
                        background = c("image_1", "EpCAM"), 
                        pt.size = set_ptsize, cell.shape = cell_shape)+
  guides(color = guide_legend(override.aes = list(size = 5)))+
  scale_color_manual(values = ColorsCellType)+
  scale_fill_manual(values = ColorsCellType)+
    theme(plot.title = element_text(size = 12), 
          legend.position = "bottom", 
          legend.text = element_text(size = 10),
          legend.title = element_text(size = 10),
          plot.margin = margin(0.5, 0, 0, 0, "cm"), 
          legend.key = element_blank())+
  ggtitle("CTRL")

plot_ctrl <- plot +
  annotate("text", x=200, y=920, label= marker1,
           col="white", size=5, parse=TRUE) +
  annotate("segment", x = 680, xend = 985, y = 45, yend = 45, size = 1.6, 
  colour = "white")


plot_ctrl


plot <- vrSpatialPlot(vr_merged, assay = paste0("Assay", 13), 
                        group.by = "CellType", 
                        group.ids = cell_type_of_interest,
                        alpha = set_alpha, 
                        background = c("image_1", "EpCAM"), 
                        pt.size = set_ptsize, cell.shape = cell_shape)+ 
  guides(color = guide_legend(override.aes = list(size = 5)))+
  scale_color_manual(values = ColorsCellType)+
  scale_fill_manual(values = ColorsCellType)+
    theme(plot.title = element_text(size = 12), 
          legend.position = "bottom", 
          legend.text = element_text(size = 10),
          legend.title = element_text(size = 10),
          plot.margin = margin(0.5, 0, 0, 0, "cm"), 
          legend.key = element_blank())+
  ggtitle("IL-33 Day 1")

plot_d1 <- plot +
  annotate("text", x=200, y=920, label= marker1,
           col="white", size=5, parse=TRUE) +
  annotate("segment", x = 680, xend = 985, y = 45, yend = 45, size = 1.6, 
  colour = "white")

plot_d1
```

Combine Spatial plots:

```{r, fig.width=9, fig.height=6.8}
plot_if_nk <- ggarrange(plot_ctrl, plot_d1, common.legend = TRUE, ncol = 2, nrow = 1, legend = "bottom", labels = c("L", "M"))

plot_nk <- ggarrange( plot_coenrichment, plot_if_nk, ncol = 2, nrow = 1, widths = c(1.5, 2), labels = c("K"))

plot_hm <- ggarrange(plot_si_ilc2s, plot_nk, nrow = 2, ncol = 1, heights = c(3.5, 3.3))

plot_hm
```

## Combine figure

```{r, fig.width=9, fig.height=12.3}
ggarrange(plot_ag, plot_hm, ncol = 1, nrow = 2, heights = c(5.5, 6.8))
```

## Session Information

```{r}
save.image(paste0(output_dir, "/environment.RData"))
sessionInfo()
```
