---
title: "Supplementary Figure 10: Quantification of ILC subtypes in the SI"
author: "Sandy Kroh"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    number_sections: yes
    fig_caption: yes
    keep_md: yes
  pdf_document:
    toc: yes
params:
  Sample: Default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", 
                      fig.align='center', dpi = 600, 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Libraries

```{r}
# remove.packages("rlang")
# remove.packages("dplyr")
# install.packages("rlang")
# install.packages("dplyr")

library(SeuratObject)
library(dplyr)
library(rstatix)
library(rlang)

if (!requireNamespace("Giotto", quietly = TRUE))
  devtools::install_github("drieslab/Giotto@suite")
if (!requireNamespace("VoltRon", quietly = TRUE))
  devtools::install_github("Artur-man/VoltRon")
if (!requireNamespace("Seurat", quietly = TRUE))
  install.packages("Seurat")
library(Giotto)
library(Seurat)
library(VoltRon)
library(ggplot2)
library(ggpubr)
library(readr)
library(ggbeeswarm)
library(stringr)
```

## Parameters

```{r}
set.seed(123)

input_dir <- here::here("1_data_tidying", "Lung_SI_all_cells_all_ALs_files")

output_dir <- here::here("2_visualizations_for_figures", "Supp_Fig_10_quantification_spatial_analysis_SI_files")
dir.create(output_dir)




main_markers <- c(
  "EpCAM", "EMCN", "LYVE1", "PDPN", "PDGFRa", "CD8a", "CD4",
  "CD45", "CD3", "IRF4", "Kappa", "CD11c", "CD127", "GATA3eGFP", "RORgt"
)


immune_markers <- c(
 "CD3", "CD4", "CD8a", "Kappa", "IRF4", "CD11c",
  "CD127", "CD90", "EOMES", "GATA3eGFP", "RORgt", "Ki67",  "KLRG1", "NKp46", "CD117", "Areg", "CCR6", "CD44", "MHCII", "Sca1"
)

ilc_markers <- c(
  "CD3", "CD4", "CD8a",
  "CD127", "CD90", "EOMES", "GATA3eGFP", "RORgt", "KLRG1", "NKp46", "CD117", "CCR6", "MHCII", "Ki67", "Areg", "IRF4", "Sca1", "CD44"
)


cols_nat <- c("magenta", "cyan", "blue", "purple", "green", 
                       "red", "yellow", "olivedrab1", "slateblue1", 
                       "darkcyan", "gold","indianred1", "seagreen", "deeppink", 
                       "orange", "brown", "violet",
                       "deeppink4", "pink", 
                       "grey", "black", "lightgreen", 
                       "#FF0066",  
                       "lightblue", "#FFCC99", "#CC00FF", 
                       "blueviolet",  "goldenrod4", 
                       "navy", "olivedrab", "lightcyan", "seagreen2", "darkviolet", "lightpink", "slateblue4", "olivedrab2")

colfunc <- colorRampPalette(c("darkcyan", "green", "yellow", "magenta", "purple"))

cols_ilcs_lung <- c("darkcyan", "seagreen2", "deeppink4")
cols_ilcs_si <- c("slateblue", "seagreen2")

cols_treat <- c("darkcyan", "gold", "slateblue")

cols_organs <- c("indianred2", "seagreen", "navy")
cols_organs_si <- c("seagreen", "navy")

```

# Load data

## Proportions

```{r, fig.width=9, fig.height=6}

df_villi <- read_csv(paste0(input_dir, "/si_villi_proportions.csv"), 
    col_types = cols(...1 = col_skip()))

df_ilf <- read_csv(paste0(input_dir, "/si_ilf_proportions.csv"), 
    col_types = cols(...1 = col_skip()))

# the FOV CTRL_FOV1_20210810 is falsely labeled as villi but is actually an ILF, so it must be changed in the metadata
df_correct <- df_villi %>%
  filter(Dataset == "CTRL_FOV1_20210810") %>%
  mutate(Tissue.area = "ILF")

df_ilf <- rbind(df_ilf, df_correct)

df_villi <- df_villi %>%
  filter(Dataset != "CTRL_FOV1_20210810")


df_lung <- read_csv(paste0(input_dir, "/lung_proportions.csv"), 
    col_types = cols(...1 = col_skip()))

```

# Visualization

## Suppl. Figure 9A - Total cell count between lung, Si villi and SI ILF at steady state

```{r, fig.width=3, fig.height=3}

df_villi_sub <- df_villi %>%
  mutate(Tissue.area = gsub("Villi", "villi", Tissue.area), 
         Tissue.area = paste0("SI ", Tissue.area)) %>%
  select(Dataset, Treatment, TotalCellCountFOV, `Immune cells`, ILCs, Tissue.area)

df_lung_sub <- df_lung %>%
  mutate(Tissue.area = `Tissue area`) %>%
  select(Dataset, Treatment, TotalCellCountFOV, `Immune cells`, ILCs, Tissue.area)

df_ilf_sub <- df_ilf %>%
  mutate(Tissue.area = paste0("SI ", Tissue.area)) %>%
  select(Dataset, Treatment, TotalCellCountFOV, `Immune cells`, ILCs, Tissue.area)

df <- rbind(df_villi_sub, df_lung_sub, df_ilf_sub)

# filter for CTRL and convert to longer format
df <- df %>%
  filter(Treatment == "CTRL") %>%
  select(Tissue.area, Dataset, TotalCellCountFOV) %>%
  mutate(Tissue.area = factor(Tissue.area, level =c(
    "Lung", "SI villi", "SI ILF"
  )))

# Testing for normal distribution
shapiro.test(df$TotalCellCountFOV)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`TotalCellCountFOV` ~ Tissue.area)
res.kruskal
df %>% kruskal_effsize(`TotalCellCountFOV` ~ Tissue.area)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`TotalCellCountFOV` ~ Tissue.area, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Tissue.area, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Tissue.area")

plot_count_all <- ggplot(df, aes(x = Tissue.area, y = TotalCellCountFOV, fill = "Tissue.area"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Tissue.area), size = 2, cex = 3)+
  scale_color_manual(values = cols_organs)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 6000) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("Total cells")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,8000))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=7700,
           label=Labels,
           #angle = 90,
           size = 10/.pt)

plot_count_all

```

## Suppl. Figure 9B - Total ILC count between lung, SI villi and SI ILF

```{r, fig.width=3, fig.height=3}

df_villi_sub <- df_villi %>%
  mutate(Tissue.area = gsub("Villi", "villi", Tissue.area), 
         Tissue.area = paste0("SI ", Tissue.area)) %>%
  select(Dataset, Treatment, TotalCellCountFOV, `Immune cells`, ILCs, Tissue.area)

df_lung_sub <- df_lung %>%
  mutate(Tissue.area = `Tissue area`) %>%
  select(Dataset, Treatment, TotalCellCountFOV, `Immune cells`, ILCs, Tissue.area)

df_ilf_sub <- df_ilf %>%
  mutate(Tissue.area = paste0("SI ", Tissue.area)) %>%
  select(Dataset, Treatment, TotalCellCountFOV, `Immune cells`, ILCs, Tissue.area)

df <- rbind(df_villi_sub, df_lung_sub, df_ilf_sub)

# filter for CTRL and convert to longer format
df <- df %>%
  filter(Treatment == "CTRL") %>%
  select(Tissue.area, Dataset, ILCs) %>%
  mutate(Tissue.area = factor(Tissue.area, level =c(
    "Lung", "SI villi", "SI ILF"
  )))

# Testing for normal distribution
shapiro.test(df$ILCs)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`ILCs` ~ Tissue.area)
res.kruskal
df %>% kruskal_effsize(`ILCs` ~ Tissue.area)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`ILCs` ~ Tissue.area, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Tissue.area, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Tissue.area")

plot_count_ilc <- ggplot(df, aes(x = Tissue.area, y = ILCs, fill = "Tissue.area"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Tissue.area), size = 2, cex = 3)+
  scale_color_manual(values = cols_organs)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 400) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("ILCs")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,650))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=600,
           label=Labels,
           #angle = 90,
           size = 10/.pt)

plot_count_ilc

```

## Suppl. Figure 9C - Total immune cells in lung, SI villi, and SI ILF

```{r, fig.width=3, fig.height=3}

df_villi_sub <- df_villi %>%
  mutate(Tissue.area = gsub("Villi", "villi", Tissue.area), 
         Tissue.area = paste0("SI ", Tissue.area)) %>%
  select(Dataset, Treatment, TotalCellCountFOV, `Immune cells`, ILCs, Tissue.area)

df_lung_sub <- df_lung %>%
  mutate(Tissue.area = `Tissue area`) %>%
  select(Dataset, Treatment, TotalCellCountFOV, `Immune cells`, ILCs, Tissue.area)

df_ilf_sub <- df_ilf %>%
  mutate(Tissue.area = paste0("SI ", Tissue.area)) %>%
  select(Dataset, Treatment, TotalCellCountFOV, `Immune cells`, ILCs, Tissue.area)

df <- rbind(df_villi_sub, df_lung_sub, df_ilf_sub)

# filter for CTRL and convert to longer format
df <- df %>%
  filter(Treatment == "CTRL") %>%
  select(Tissue.area, Dataset, `Immune cells`) %>%
  mutate(Tissue.area = factor(Tissue.area, level =c(
    "Lung", "SI villi", "SI ILF"
  )))

# Testing for normal distribution
shapiro.test(df$`Immune cells`)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`Immune cells` ~ Tissue.area)
res.kruskal
df %>% kruskal_effsize(`Immune cells` ~ Tissue.area)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`Immune cells` ~ Tissue.area, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Tissue.area, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Tissue.area")

plot_count_immune <- ggplot(df, aes(x = Tissue.area, y = `Immune cells`, fill = "Tissue.area"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Tissue.area), size = 2, cex = 3)+
  scale_color_manual(values = cols_organs)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 5000) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("Immune cells")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,6900))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=6500,
           label=Labels,
           #angle = 90,
           size = 10/.pt)

plot_count_immune

```

## Suppl. Figure 9D - Frequency of ILCs in lung, SI villi adn SI ILF

```{r, fig.width=3, fig.height=3}

df_villi_sub <- df_villi %>%
  mutate(Tissue.area = gsub("Villi", "villi", Tissue.area), 
         Tissue.area = paste0("SI ", Tissue.area)) %>%
  select(Dataset, Treatment, TotalCellCountFOV, `Immune cells`, ILCs, Tissue.area)

df_lung_sub <- df_lung %>%
  mutate(Tissue.area = `Tissue area`) %>%
  select(Dataset, Treatment, TotalCellCountFOV, `Immune cells`, ILCs, Tissue.area)

df_ilf_sub <- df_ilf %>%
  mutate(Tissue.area = paste0("SI ", Tissue.area)) %>%
  select(Dataset, Treatment, TotalCellCountFOV, `Immune cells`, ILCs, Tissue.area)

df <- rbind(df_villi_sub, df_lung_sub, df_ilf_sub)

# filter for CTRL and convert to longer format
df <- df %>%
  filter(Treatment == "CTRL") %>%
  mutate(ILC_freq = ILCs/`Immune cells`*100) %>%
  select(Tissue.area, Dataset, ILC_freq) %>%
  mutate(Tissue.area = factor(Tissue.area, level =c(
    "Lung", "SI villi", "SI ILF"
  )))

# Testing for normal distribution
shapiro.test(df$ILC_freq)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`ILC_freq` ~ Tissue.area)
res.kruskal
df %>% kruskal_effsize(`ILC_freq` ~ Tissue.area)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`ILC_freq` ~ Tissue.area, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ Tissue.area, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Tissue.area")

plot_freq_ilc <- ggplot(df, aes(x = Tissue.area, y = ILC_freq, fill = "Tissue.area"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = Tissue.area), size = 2, cex = 3)+
  scale_color_manual(values = cols_organs)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 15) +
  xlab(NULL)+
  ylab("Frequency/\nimmune cells per FOV [%]")+
  ggtitle("ILCs")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,22))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1, 2, 3),
           y=20,
           label=Labels,
           #angle = 90,
           size = 10/.pt)

plot_freq_ilc

```

Combine plots:

```{r, fig.width=9, fig.height=6}
ggarrange(plot_count_all, plot_count_immune, plot_count_ilc, plot_freq_ilc, ncol = 3, nrow = 2, labels = "AUTO")
```

## Suppl. Figure 9E - ILC subtypes in SI villi and SI ILF at steady state

```{r, fig.width=9, fig.height=6}

df_villi_sub <- df_villi %>%
  mutate(Tissue.area = gsub("Villi", "villi", Tissue.area), 
         Tissue.area = paste0("SI ", Tissue.area)) %>%
  select(Dataset, Treatment, TotalCellCountFOV, `Immune cells`, ILCs, `NK cells/ILC1s/ILC3s`, ILC2s, Tissue.area) %>%
  #change from wide to long format
  tidyr::pivot_longer(names_to = "ILC_type", cols = c(`NK cells/ILC1s/ILC3s`, ILC2s))

df_ilf_sub <- df_ilf %>%
  mutate(Tissue.area = paste0("SI ", Tissue.area)) %>%
  select(Dataset, Treatment, TotalCellCountFOV, `Immune cells`, ILCs, `NK cells/ILC1s/ILC3s`, ILC2s, Tissue.area) %>%
  #change from wide to long format
  tidyr::pivot_longer(names_to = "ILC_type", cols = c(`NK cells/ILC1s/ILC3s`, ILC2s))

df_long <- rbind(df_villi_sub, df_ilf_sub)

# filter for CTRL and convert to longer format
df_long <- df_long %>%
  filter(Treatment == "CTRL") %>%
  select(Tissue.area, Dataset, ILC_type, value) %>%
  mutate(Tissue.area = factor(Tissue.area, level =c(
    "Lung", "SI villi", "SI ILF"
  )), 
  ILC_type = factor(ILC_type, level =c(
    "NK cells/ILC1s/ILC3s", "ILC2s"
  )))

# VILLI -----------------------------------------------------------------
df <- df_long %>%
  filter(Tissue.area == "SI villi")

# Testing for normal distribution
shapiro.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILC_type)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILC_type)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILC_type, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILC_type, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILC_type")

plot_villi_ilcs <- ggplot(df, aes(x = ILC_type, y = value, fill = "Tissue.area"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILC_type), size = 2, cex = 5)+
  scale_color_manual(values = cols_ilcs_si)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 150) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("SI villi")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,310))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1.5),
           y=290,
           label=Labels[1],
           #angle = 90,
           size = 10/.pt)

# ILF -----------------------------------------------------------------
df <- df_long %>%
  filter(Tissue.area == "SI ILF")

# Testing for normal distribution
shapiro.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILC_type)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILC_type)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILC_type, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILC_type, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILC_type")

plot_ilf_ilcs <- ggplot(df, aes(x = ILC_type, y = value, fill = "Tissue.area"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILC_type), size = 2, cex = 5)+
  scale_color_manual(values = cols_ilcs_si)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 150) +
  xlab(NULL)+
  ylab("Total count per FOV [#]")+
  ggtitle("SI ILF")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,310))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1.5),
           y=290,
           label=Labels[1],
           #angle = 90,
           size = 10/.pt)


```

## Suppl. Figure 9F - ILC proportions in the 3 tissue areas/organs

```{r, fig.width=9, fig.height=3}

# get all three organs/ tissue area data into long format
df_villi_sub <- df_villi %>%
  mutate(Tissue.area = gsub("Villi", "villi", Tissue.area), 
         Tissue.area = paste0("SI ", Tissue.area)) %>%
  select(Dataset, Treatment, ILCs, `NK cells/ILC1s/ILC3s`, ILC2s, Tissue.area) %>%
  #change from wide to long format
  tidyr::pivot_longer(names_to = "ILC_type", cols = c(`NK cells/ILC1s/ILC3s`, ILC2s)) %>%
  mutate(value = value/ILCs*100)


df_ilf_sub <- df_ilf %>%
  mutate(Tissue.area = gsub("Villi", "villi", Tissue.area), 
         Tissue.area = paste0("SI ", Tissue.area)) %>%
  select(Dataset, Treatment, ILCs, `NK cells/ILC1s/ILC3s`, ILC2s, Tissue.area) %>%
  #change from wide to long format
  tidyr::pivot_longer(names_to = "ILC_type", cols = c(`NK cells/ILC1s/ILC3s`, ILC2s)) %>%
  mutate(value = value/ILCs*100)


df_lung_sub <- df_lung %>%
  mutate(Tissue.area = `Tissue area`) %>%
  select(Dataset, Treatment, ILCs, `NK cells/ILC1s`, ILC2s, ILC3s, Tissue.area) %>%
  #change from wide to long format
  tidyr::pivot_longer(names_to = "ILC_type", cols = c(`NK cells/ILC1s`, ILC2s, ILC3s)) %>%
  mutate(value = value/ILCs*100)



df_long <- rbind(df_villi_sub, df_ilf_sub, df_lung_sub)

# filter for CTRL and convert to longer format
df_long <- df_long %>%
  filter(Treatment == "CTRL") %>%
  select(Tissue.area, Dataset, ILC_type, value) %>%
  mutate(Tissue.area = factor(Tissue.area, level =c(
    "Lung", "SI villi", "SI ILF"
  )), 
  ILC_type = factor(ILC_type, level =c(
    "NK cells/ILC1s",
    "NK cells/ILC1s/ILC3s", "ILC2s", "ILC3s"
  )))

# LUNG ----------------------------------------------------------------------------
df <- df_long %>%
  filter(Tissue.area == "Lung") %>%
  mutate(
    ILC_type = factor(ILC_type, level =c(
    "NK cells/ILC1s",
    "ILC2s", "ILC3s"
  )))

# Testing for normal distribution
shapiro.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILC_type)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILC_type)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILC_type, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILC_type, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILC_type")

plot_lung_ilcs <- ggplot(df, aes(x = ILC_type, y = value, fill = "Tissue.area"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILC_type), size = 2, cex = 5)+
  scale_color_manual(values = cols_ilcs_lung)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1.6,1,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.08, y.position = 85) +
  xlab(NULL)+
  ylab("Frequency/\nimmune cells FOV [%]")+
  ggtitle("Lung")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,100))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1),
           y=95,
           label=Labels[1],
           #angle = 90,
           size = 10/.pt)

# SI VILLI ----------------------------------------------------------------------------
df <- df_long %>%
  filter(Tissue.area == "SI villi")

# Testing for normal distribution
shapiro.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILC_type)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILC_type)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILC_type, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILC_type, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILC_type")

plot_villi_ilcs <- ggplot(df, aes(x = ILC_type, y = value, fill = "Tissue.area"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILC_type), size = 2, cex = 5)+
  scale_color_manual(values = cols_ilcs_si)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 50) +
  xlab(NULL)+
  ylab("Frequency/\nimmune cells FOV [%]")+
  ggtitle("SI villi")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,100))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1.5),
           y=95,
           label=Labels[2],
           #angle = 90,
           size = 10/.pt)

# SI ILF ----------------------------------------------------------------------------
df <- df_long %>%
  filter(Tissue.area == "SI ILF")

# Testing for normal distribution
shapiro.test(df$value)

# Kruskal-Wallis-test to check for significance between tested groups and effect size
res.kruskal <- df %>% kruskal_test(`value` ~ ILC_type)
res.kruskal
df %>% kruskal_effsize(`value` ~ ILC_type)

# Pairwise comparisons using Dunn's test
pwc <- df %>% 
  dunn_test(`value` ~ ILC_type, p.adjust.method = "bonferroni") 
pwc

# add N to plot
tab <- data.frame(xtabs(~ ILC_type, data = df))
head(tab)
# Add cell number per cluster to cluster labels
Labels = paste0("n = ", tab$Freq, "")



# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "ILC_type")

plot_ilf_ilcs <- ggplot(df, aes(x = ILC_type, y = value, fill = "Tissue.area"))+
  geom_boxplot(fill="white")+
  geom_beeswarm(aes(color = ILC_type), size = 2, cex = 5)+
  scale_color_manual(values = cols_ilcs_si)+
  theme_classic2()+
  theme(plot.margin=margin(1,0.5,1,1,"cm"),
        axis.text.x = element_text(angle = 30, 
                                   vjust = 1, size = 12, hjust = 1, face = "bold"),
         axis.text.y = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size =14, hjust = 0.5),
        legend.title = element_text(size =14),
        legend.text = element_text(size =12))+
  stat_pvalue_manual(pwc,
                       hide.ns = TRUE, size = 6,
                       step.increase = 0.1, y.position = 50) +
  xlab(NULL)+
  ylab("Frequency/\nimmune cells FOV [%]")+
  ggtitle("SI ILF")+
  scale_y_continuous(expand = c(0, 0), limits = c(0,100))+
  NoLegend()+
    annotate(geom = 'text',
           x=c(1.5),
           y=95,
           label=Labels[2],
           #angle = 90,
           size = 10/.pt)

ggarrange(plot_lung_ilcs, plot_villi_ilcs, plot_ilf_ilcs, ncol = 3, nrow = 1, labels = c("E", "F", "G"))


```

## Combine plots

```{r, fig.width=9, fig.height=6}
ggarrange(plot_count_all, plot_count_immune, plot_count_ilc, plot_freq_ilc, plot_villi_ilcs, plot_ilf_ilcs, ncol = 3, nrow = 2, labels = "AUTO")
```

## Session Information

```{r}
save.image(paste0(output_dir, "/environment.RData"))
sessionInfo()
```
